---
title: "gdf to geotiff"
author: "Michelle DePrenger-Levin"
date: "December 24, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rgdal)
library(raster)
```

```{r}
adfs <- list.dirs(path = "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado",full.names = TRUE, recursive = TRUE)


adfs[1:(length(adfs)/2)]
length(unique(adfs))
# hdrs <- lapply(36:46, function(x){
#   out <- grep(paste("n",x,sep=""),adfs)
#   outgrdn <- grep("grdn",out)
# })
length(adfs[grep("grdn",adfs)])

adfs[grep("grdn",adfs)][1]


adf2tif <- lapply(adfs[grep("grdn",adfs)], function(x) raster(paste(x,"/hdr.adf",sep="")))


adfs[[1]]

plot(adf2tif[[1]])
lapply(1:length(adf2tif), function(r){
  writeRaster(adf2tif[[r]],paste(adfs[[1]],"/",
                                 gsub("^.*/", "", adfs[grep("grdn",adfs)][r]),
                                 '.tif',sep=""),options=c('TFW=YES'), overwrite=TRUE)

})
```

Missed 
       n38w108    
       n37w104   
       n36w111
```{r}
missinghdrs <- c("n38w108","n37w104","n36w111")

adfs[grep(paste(missinghdrs, collapse ="|"), adfs, value=TRUE) ]

getem <- lapply(lapply(missinghdrs, function(x) adfs[grep(x,adfs)]), 
                          function(x) x[grep("_13", x)])

adf2tif2 <- lapply(getem,
                   function(x) raster(paste(x[[1]],"/hdr.adf",sep="")))


lapply(1:length(adf2tif2), function(r){
  writeRaster(adf2tif2[[r]],paste(adfs[[1]],"/",
                                 gsub("^.*/", "", getem[grep("grdn",getem)][r]),
                                 '.tif',sep=""),options=c('TFW=YES'), overwrite=TRUE)

})

writeRaster(adf2tif2[[3]], paste(adfs[[1]],"/",
                                 gsub("^.*/", "", getem[grep("grdn",getem)][[3]]),
                                 '.tif',sep="")[[1]],options=c('TFW=YES'), overwrite=TRUE)
```


```{r}

r <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/aspect25int/hdr.adf")

writeRaster(r, "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/aspect25int.tif", options=c('TFW=YES'))


r <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/coplus25int/hdr.adf")

writeRaster(r, "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/coplus25int.tif", options=c('TFW=YES'))

#Slopedeg
r <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/slopedeg25int/hdr.adf")

writeRaster(r, "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/slopedeg25int.tif", options=c('TFW=YES'))


#rugged
r <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/rugged25int/hdr.adf")

writeRaster(r, "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/rugged25int.tif", options=c('TFW=YES'))

```

Turn US ecoregions level 4 into a raster for VisTrails    
```{r}
library(raster)
library(rgdal)

p <- shapefile("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/Ecogregions of the US/us_eco_l4.shp")
proj4string(p)
# [1] "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs +ellps=GRS80 +towgs84=0,0,0"
# in meters, can do 10 meter
p

ext <- floor(extent(p))
rout <- raster(ext, res = 1000)
r <- rasterize(p, rout, field=1)
plot(r)
# 
# library(doParallel)
# library(foreach)
# 
# usecores <- detectCores() -1
# cl <- makeCluster(usecores)
# registerDoParallel(cl)

path <- "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/"

writeRaster(r, filename = paste0(path,"EmptyRaster.tif"))


# foreach(i = )


```
Note that you are rasterizing very small polygons to large raster cells. A polygon is considered 'inside' if it covers the center of a cell (i.e. assuming a case where polygons cover multiple cells). So for these data you would need to use a much higher resolution (and then perhaps aggregate the results). Alternatively you could rasterize polygon centroids.   

