---
title: "R Notebook"
output: html_notebook
---

I think the outputs that don't have _EOR or _herb are the ones that used Herb to train and EOR to test. can see what the AUC patterns were for those. 
Making new herb train and EOR test ones with VisTrails and same size EOR numbers
```{r}
rm(list=ls())

library(ggplot2)
library(rgeos)
library(raster)

# pathstart <- "Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/RMaxEnt/Arabis crandallii_EOR/"
# 
# resultpath <- list.files(path = pathstart, pattern = "maxentResults.csv") # , full.names = TRUE)[[1]]

# I don't think I want these, went through one by one but had issues. 
    # maxouts <- list.dirs(path = "Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/RMaxEnt",full.names = TRUE, recursive = TRUE)
    # 
    # maxoutsubset <- maxouts[!grepl("plots", maxouts)]
    # maxoutsubset <- maxoutsubset[!grepl("_EOR", maxoutsubset)]
    # maxoutsubset <- maxoutsubset[!grepl("_herb", maxoutsubset)]
    # 
    # sapply(strsplit(maxoutsubset,"/"), function(x){
    #   x[7]
    # })

# Where is the information about EOR size and distances to match to the names? 

      # AUCresults <- do.call(rbind,lapply(maxoutsubset, function(x){
      #   read.csv(paste(x,"/maxentResults.csv", sep=""))
      # }))
      # 
      # names(AUCresults)[grep("AUC",names(AUCresults))]
      # 
      # jpeg("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Images/AUCtraintest1.jpg",
      #      width=200, height=175,units='mm', res=300)
      # hist(AUCresults$Training.AUC, breaks = 10,
      #      main="",
      #      xlab="AUC")
      # dev.off()
```

Get all the files that I'm making with VisTrails, then extract the information from the sets I want. Named similar runs the same way. 
```{r}
maxouts2 <- list.dirs(path = "Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/VisTrails/EORvHERB_2019/EORvHERB_2019",full.names = TRUE, recursive = TRUE)


#EORsamesize matches _Sp with the same number
# These all have k-fold of 5 so keeping out 20% each time to test but for some, too few points, so should compare to subsampling with replacement/bootstrapping 
EORsame <- maxouts2[grepl("_EORsamesize", maxouts2)]
speciesEOR <- unique(unlist(lapply(regmatches(EORsame, gregexpr("[[:digit:]]+", EORsame)),"[[", 3)))

# Must be an easier way but only want the first folder, don't want to dig in any deeper
EORsamesub <- EORsame[!grepl("Merged", EORsame)]
EORsamesub <- EORsamesub[!grepl("plots", EORsamesub)]
EORsamesub <- EORsamesub[!grepl("cvSplit", EORsamesub)]
EORsamesub <- EORsamesub[!grepl("Expanded", EORsamesub)]
EORsamesub <- EORsamesub[!grepl("responseCurves", EORsamesub)]

# The 9th layer is the folder for that Maxent output
unique(sapply(strsplit(EORsamesub,"/"), function(x){
  x[9]
}))

# Which species number is each? 
unique(sapply(strsplit(EORsame,"Sp"), function(x){
  as.numeric(gsub("([0-9]+).*$", "\\1", x[2]))
}))

# Split at the Sp, take that second string that starts with the species number and extract the first number
sapply(strsplit(EORsamesub, "Sp"), function(x){
  as.numeric(gsub("([0-9]+).*$", "\\1", x[2]))
})

# Where is the information about EOR size and distances to match to the names? 
AUCresults2 <- do.call(rbind,lapply(EORsamesub, function(x){
  resultfile <- read.csv(paste(x,"/maxentResults.csv", sep=""))
  resultfile$Species <- as.numeric(resultfile$Species)
  y <- strsplit(x,"Sp")
  resultfile$Species <- as.numeric(gsub("([0-9]+).*$", "\\1", y[[1]][2]))
  resultfile
}))

hist(AUCresults2$Training.AUC)

jpeg("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Images/AUC_EOR.jpg",
     width=200, height=175,units='mm', res=300)
hist(AUCresults2$Training.AUC, breaks = 10,
     main="",
     xlab="AUC")
dev.off()
```

Now get these rasters and see where the EOR differs from the Herb
```{r}
# Use the resampled one instead
# coElev <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/coplus50int.tif")
# coAspect <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/aspect50int.tif")
# coSlope <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/COplus_slope50.tif")
# coRugged <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/COplus_ruggedInt50.tif")

x<-EORsamesub[[1]]
raster1 <- raster(paste(x,"/Maxent_prob_map.tif", sep=""))

# resample(coAspect, raster1, method="bilinear", filename = "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/AspectResampled.tif")
# 
# resample(coSlope, raster1, method="bilinear", filename = "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/SlopeResampled.tif")
# 
# resample(coRugged, raster1, method="bilinear", filename = "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/RuggedResampled.tif")

# Get raster1 to match coElev, no go coElev to raster1
# resample(coElev, raster1, method="bilinear", filename = "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/ElevationResampled.tif")

# resample(coElev, raster1, method="bilinear", filename = "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/ElevationResampled.tif")

coElev_res <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/ElevationResampled.tif")

coAspect_res <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/AspectResampled.tif")

coSlope_res <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/SlopeResampled.tif")

coRugged_res <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/RuggedResampled.tif")

# Function to calculate the linear model of one raster to another
# 1 would be the intercept I guess, while y are the values for y (from the resulting probability map)
# Don't quite get it
          # X <- cbind(1, y)
          # invXtX <- solve(t(X) %*% X) %*% t(X)
          # quickfun <- function(i) (invXtX %*% i)
          # # raster::calc calculates values for a new raster from another
          # m <- calc(s, quickfun) 
          # names(m) <- c('intercept', 'slope')
# r <- calc(s, fun)
# y <- values(raster1)
# X1 <- values(coElev)

# Could also ignore the not likely but that cuts off lots... maybe just the really near 0?
raster1sm <- raster1
raster1sm[raster1sm<0.05] <- NA

# But this is basically just what maxent does, right? but with a prior that is flat or 'maximizes entropy'?
s <- stack(raster1, coElev_res, coAspect_res, coSlope_res, coRugged_res)
# Make it a little smaller
v <- sampleRegular(s, 100000)  
v <- data.frame(na.omit(v))
variables <- names(v)
m <- lm(v[,1] ~ v[,2] + v[,3] + v[,4] + v[,5])
summary(m)
plot(v[,5],v[,1],
     xlab = names(v)[5],
     ylab = names(v)[1])

# What do I really want to see? Maybe AIC if the model is different for Herb vs. EOR? That would be similar to loadings fo ra PCA? 
library(ks)
library(MASS)
?kde2d
plot(v[,1],v[,2])
contour(kde2d(v[,1], v[,2], lims = c(min(v[,1]), max(v[,1]),
                                     min(v[,2]), max(v[,2]))), 
        levels= c(min(v[,1]), max(v[,1]),
                                     min(v[,2]), max(v[,2])))

# X <- cbind(1, values(s))
# X <- cbind(1, as.matrix(v))
# invXtX <- solve(t(X) %*% X) %*% t(X)
# quickfun <- function(i) (invXtX %*% i)
# m <- quickfun(as.matrix(v))
# summary(m)

EORrasters <- lapply(EORsamesub, function(x){
  raster1 <- raster(paste(x,"/Maxent_prob_map.tif", sep=""))
  s <- stack(raster1, coElev_res, coAspect_res, coSlope_res, coRugged_res)
  v <- sampleRegular(s, 100000)  
  v <- data.frame(na.omit(v))
  variables <- names(v)
  m <- lm(v[,1] ~ v[,2])
  summary(m)
  plot(v[,2],v[,1])
  abline(m)
})

```



Now the herbarium specimens!
```{r}
# Compare to the Herb results
Herbsame <- maxouts2[grepl("Maxent_Sp", maxouts2)]
speciesHerb <- unique(unlist(lapply(regmatches(EORsame, gregexpr("[[:digit:]]+", EORsame)),"[[", 3)))
match(speciesHerb,speciesEOR)

# Must be an easier way but only want the first folder, don't want to dig in any deeper
Herbsamesub <- Herbsame[!grepl("Merged", Herbsame)]
Herbsamesub <- Herbsamesub[!grepl("plots", Herbsamesub)]
Herbsamesub <- Herbsamesub[!grepl("cvSplit", Herbsamesub)]
Herbsamesub <- Herbsamesub[!grepl("Expanded", Herbsamesub)]
Herbsamesub <- Herbsamesub[!grepl("responseCurves", Herbsamesub)]

# read.csv(paste(Herbsamesub[[5]],"/maxentResults.csv", sep=""))

AUClist <- lapply(Herbsamesub, function(x){
  resultfile <- read.csv(paste(x,"/maxentResults.csv", sep=""))
  resultfile$Species <- as.numeric(resultfile$Species)
  y <- strsplit(x,"Sp")
  resultfile$Species <- as.numeric(gsub("([0-9]+).*$", "\\1", y[[1]][2]))
  resultfile
})


# Go back and see if can run those without threshold and such so they're the same. 
# don't want 5, 7, 10
AUCresults3 <- do.call(rbind,lapply(c(1:4,6,8:9,11:14), function(x){
  AUClist[[x]]
}))


# AUCresults3 <- do.call(rbind,lapply(Herbsamesub, function(x){
#   read.csv(paste(x,"/maxentResults.csv", sep=""))
# }))

hist(AUCresults3$Training.AUC)

jpeg("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Images/AUC_Herb.jpg",
     width=200, height=175,units='mm', res=300)
hist(AUCresults3$Training.AUC, breaks = 10,
     main="",
     xlab="AUC")
dev.off()
# TestEOR_ only want one per species, manually go in and delete extras that had mishaps

```

If i can get more, add 'em up
```{r}


```

# Rounds of VisTrails with elevation: coplus50int; aspect: aspect50int; ruggedness: COplus_ruggedInt50; annual temp Bio1 and annual precip Bio12 

Name the EOR ones as EORSp1... and compare   
<https://gis.stackexchange.com/questions/29118/how-to-find-the-average-raster-value-of-an-area-defined-by-a-shapefile-using-r> 
```{r}
dirs.vistrails <- list.dirs(path = "Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/VisTrails/EORvHERB_2019/EORvHERB_2019",full.names = TRUE, recursive = TRUE) 

dirs.vistrails[grep("_Sp",dirs.vistrails)]


```


```{r}
load("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/pca_data.Rda")

# Instead, load:
load("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/distXsp_v3.Rda")


linkspeciesnumber <- do.call(rbind,lapply(1:length(pca_data), function(x){
  data.frame(SpNum = x, Species = unique(pca_data[[x]]$GNAME[pca_data[[x]]$pa==3]),
             Dist = pca_data[[x]]$dist[pca_data[[x]]$pa==2], # pa==2 are herb
             SpeciesArea = unique(pca_data[[x]]$area[pca_data[[x]]$pa==1]))
}))

# I don't think the area numbers are right!
for(i in 1:46){
  print(table(pca_data[[i]]$area, pca_data[[i]]$pa))
}


# the EOR AUC results from VisTrails, samesize runs from ForVisTrails.Rmd where I either thinned to match number of Herbarium records or where I bootstrapped and jittered to get matching points. Used a standalone Dan Warren thinning process to get equally spaced out but covering the most area (of information) for points. 
EORmaxentdata <- merge(linkspeciesnumber, AUCresults2, by.x = "SpNum", by.y = "Species")

avgDist <- aggregate(EORmaxentdata$Dist, list(EORmaxentdata$SpNum), mean) 
sdDist <- aggregate(EORmaxentdata$Dist, list(EORmaxentdata$SpNum), sd) 
EORsummarymaxent <- merge(unique(linkspeciesnumber[,-3]), AUCresults2, by.x = "SpNum", by.y = "Species")

summary(lm(Training.AUC ~ Dist,data = EORmaxentdata))

ggplot(EORmaxentdata, aes(SpeciesArea, Training.AUC))+
  geom_point()+
  stat_smooth(method="lm")+
  theme_bw()

ggplot(EORmaxentdata, aes(Dist, Training.AUC))+
  geom_point()+
  stat_smooth(method="lm")+
  theme_bw()


```


Want to add vegetation types, layer too. should work now that I've got all the other layers a better size

Make maxent commands
for output prediction map, use ENMeval   
```{r}


```


PCA for how different
```{r}



```
