---
title: "R Notebook"
output: html_notebook
---

Ignore blocks with 'morethan12' and skip to atleast12

I think the outputs that don't have _EOR or _herb are the ones that used Herb to train and EOR to test. can see what the AUC patterns were for those. 
Making new herb train and EOR test ones with VisTrails and same size EOR numbers
```{r}
rm(list=ls())

library(ggplot2)
library(rgeos)
library(raster)
library(foreach)
library(parallel)
library(doParallel)

library(maptools)
library(dismo)
library(Rmisc)
library(rgdal)
library(ENMeval)
library(Taxonstand)

load("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/coloradosps.g1g2_168.Rda")
load("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/g1g2namesall68.Rda")
load("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/distXsp_noEORgrid.Rda")
load("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/distXsp_noEORgrid68.Rda")
load("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/distXsp_noEORgrid1_68.Rda")
l1eor <- readShapePoly("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/L1shp")
#Each polygon might be made of multiple disconnected polygons. Need to separate and label
l1G1G2 <- disaggregate(l1eor[l1eor$GRANK %in% c("G1","G2"),])
l1G1G2$PolyID <- do.call(rbind, lapply(split(l1G1G2,l1G1G2$OBJECTID),
                                       function(x){
              x$POlyID <- paste(x$EO_ID, LETTERS[seq(from=1, to= nrow(x))],sep="")
              x
              }))

l1G1G2and <- disaggregate(l1eor[l1eor$GRANK %in% c("G1","G2","G1G2"),])
l1G1G2and$PolyID <- do.call(rbind, lapply(split(l1G1G2and,l1G1G2and$OBJECTID),
                                       function(x){
              x$POlyID <- paste(x$EO_ID, LETTERS[seq(from=1, to= nrow(x))],sep="")
              x
              }))
namesg1g2 <- table(as.character(l1G1G2$GNAME))
length(namesg1g2) #60 species
namesg1g2and <- table(as.character(l1G1G2and$GNAME))
length(namesg1g2and) # 68
#What? are some in twice? 
namesg1g2and <- c(namesg1g2and[match(names(namesg1g2),names(namesg1g2and))],
                  namesg1g2and[setdiff(names(namesg1g2and),names(namesg1g2))])

# g1g2names <- TPL(names(namesg1g2))
load("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/g1g2names.Rda")
g1g2names61_68 <- TPL(names(namesg1g2and[61:68]))


notnull <- which(!unlist(lapply(distXsp_noEORgrid, is.null)))
notnull61668 <- which(!unlist(lapply(distXsp_noEORgrid68, is.null)))

distXspall <- do.call(rbind, Map(cbind, distXsp_noEORgrid[notnull], 
                     SpNum = notnull))
distXsp6168 <- do.call(rbind, Map(cbind, distXsp_noEORgrid68[notnull61668],
                                  SpNum = notnull61668+60))

atleast12 <- c()
for(l in notnull){
  atleast12[match(l,notnull)] <- if(nrow(distXsp_noEORgrid[[l]])>11){ 
    l } else {
      NA
    }
}
atleast12 <- atleast12[!is.na(atleast12)] 
rm(l)

#Each polygon might be made of multiple disconnected polygons. Need to separate and label
length(namesg1g2) #60 species
length(namesg1g2and) # 68
g1g2namesall68 <- TPL(names(namesg1g2and))
g1g2namesall68$AcceptedName <- paste(g1g2namesall68$New.Genus,g1g2namesall68$New.Species)


# atleast1268 <- c()
# for(l in notnull61668){
#   atleast1268[match(l,notnull61668)] <- if(nrow(distXsp_noEORgrid68[[l]])>11){ 
#     l } else {
#       NA
#     }
# }
# atleast1268 <- atleast1268[!is.na(atleast1268)] 
# rm(l)

# February 26, 2019 added _bioclim where I've resampled to the bioclim raster cell size (less than 1km)
# coplus50int.tif
coElev_res <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/ElevationResampled_bioclim.tif")
# aspect50int.tif
coAspect_res <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/AspectResampled_bioclim.tif")

# COplus_ruggedInt50.tif
coRugged_res <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/RuggedResampled_bioclim.tif")

# plot(coRugged_res)
# bio1 <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/BioClim/20190122_WorldClimv2/wc2.0_bio_30s_01.tif")
# bio12 <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/BioClim/20190122_WorldClimv2/wc2.0_bio_30s_12.tif")
# 
# bio1_res <- resample(bio1, coRugged_res, method="bilinear",
#                      filename = "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/Bio1Resampled_bioclim.tif", overwrite=TRUE)
# 
# bio12_res <- resample(bio12, coRugged_res, method="bilinear",
                       # filename = "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/Bio12Resampled_bioclim.tif", overwrite=TRUE)


bio1_res <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/Bio1Resampled_bioclim.tif")
bio12_res <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/Bio12Resampled_bioclim.tif")
gc()


rasterstack <- stack(coElev_res,coAspect_res,coRugged_res,bio1_res,bio12_res)
rm(bio1_res,bio12_res,coAspect_res,coElev_res,coRugged_res)
gc()

stitchtogether <- function(whichones, pathstart, patternmatch, rasternames){
  lapply(whichones, function(i){
  gc()
  lapply(1:10, function(k){
    resultpath <- list.files(path = pathstart, 
                             pattern = paste(patternmatch,i,"kfold",k,"_",sep=""), 
                             full.names=TRUE)
    rastout <- lapply(resultpath, function(x){
      raster(x)
      })
    rastout$filename <- paste(pathstart,"ProbTiffSp",i,rasternames,k,".tif", sep="")
    rastout$overwrite <- TRUE
    m <- do.call(merge, rastout)
  })
})
}



#Need to averge and SD each of the 10

library(cluster)
library(snow)

forhistofaverage <- function(whichones = atleast12, pathstart, patternmatch){
  stacktoaverage <- lapply(whichones, function(i){
    rasterstoavg <- list.files(path = pathstart, 
                               pattern = paste("ProbTiffSp",
                                               i,
                                               patternmatch,sep=""), 
                               full.names=TRUE)
    ras <- stack(lapply(rasterstoavg, function(y){
      raster(y)
    }))
    beginCluster(10)
    ras.mean <- clusterR(ras, calc, args=list(mean, na.rm=T))
    writeRaster(ras.mean, paste(pathstart,
                                "AvgTiffSp",i,patternmatch,g1g2namesall68$AcceptedName[i],
                                ".tif", sep=""),overwrite=TRUE)
    gc()
    endCluster()
    ras.mean
  })
  stacktoaverage
}

forhistofsd <- function(whichones = atleast12, pathstart, patternmatch){
  stacktoaverage <- lapply(whichones, function(i){
    rasterstoavg <- list.files(path = pathstart, 
                               pattern = paste("ProbTiffSp",
                                               i,
                                               patternmatch,sep=""), 
                               full.names=TRUE)
    ras <- stack(lapply(rasterstoavg, function(y){
      raster(y)
    }))
    beginCluster(10)
    ras.mean <- clusterR(ras, calc, args=list(sd, na.rm=T))
    writeRaster(ras.mean, paste(pathstart, "SDTiffSp",i,patternmatch,
                              g1g2namesall68$AcceptedName[i],
                              ".tif", sep=""),overwrite=TRUE)
    gc()
    endCluster()
    ras.mean
  })
  stacktoaverage
}


load("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/distXsp_noEORgrid.Rda")
distXspall <- do.call(rbind, distXsp_noEORgrid)
#The over 400 km is clearly wrong; so are any above 300km is out of the state (if point is in the middle) so need to cutoff below 300,000
```

```{r}
# Now just putting in a data.frame that needs to be made into a spatialpointsdataframe, from herbarium so x y are longitude and latitude
maxentrun <- function(whichones, spatialpointsdataframe_herb, numberofReps = 10, 
                      maxentarguments = FALSE, predictorvariables = rasterstack, 
                      pathstart, filenames, kfoldnum = 4, 
                      error = FALSE, distdistribution = NULL){
        for(x in whichones){
            pointsspdf <- SpatialPointsDataFrame(coords = spatialpointsdataframe_herb[[x]][,c("decimalLongitude","decimalLatitude")],
                                                 data = spatialpointsdataframe_herb[[x]],
                                                 proj4string = CRS("+proj=utm +zone=13 ellps=NAD83 +ellps=WGS84"))
          if(error == TRUE){
               # for each point I will draw a circle of size (drawn from the distribution of error seen distXspall$Dist) and then pick a random point along the circle.
              errorpointsout <- do.call(rbind,lapply(1:nrow(pointsspdf), function(r){
                errordist <- sample(distdistribution, 1)
                if(errordist>0){
                  erroraround <- gBuffer(pointsspdf[r,], width=errordist)
                  newpoint <- erroraround@polygons[[1]]@Polygons[[1]]@coords
                  out <- newpoint[sample(1:nrow(newpoint),1),]
                } else {
                  out <- pointsspdf@coords[r,]
                  }
                out
                }))
                
                df <- SpatialPointsDataFrame(coords = errorpointsout,
                                             data = pointsspdf@data,
                                             proj4string = CRS("+proj=utm +zone=13 ellps=NAD83 +ellps=WGS84"))
                circlesout <- circles(df, d = 5000) #Should be 5km around
                polygns <- polygons(circlesout)
                bgpnts <- spsample(polygns, 300, "stratified") # one single random location in each 'cell' 
                
                convertxy <- spTransform(df, CRS("+proj=longlat +datum=WGS84"))
                proj4string(bgpnts) <- CRS("+proj=utm +zone=13 ellps=NAD83 +ellps=WGS84")
                bgpnts <- spTransform(bgpnts, CRS("+init=epsg:4326") )
          } else {
                convertxy <- spTransform(pointsspdf, CRS("+proj=longlat +datum=WGS84"))
                circlesout <- circles(pointsspdf, d = 5000)
                polygns <- polygons(circlesout)
                bgpnts <- spsample(polygns, 300, "stratified")
                proj4string(bgpnts) <- CRS("+proj=utm +zone=13 ellps=NAD83 +ellps=WGS84")
                bgpnts <- spTransform(bgpnts, CRS("+init=epsg:4326") )
            
          }
                
          for(rep in 1:numberofReps){
                convertxy$kfold <- kfold(convertxy, k=kfoldnum) # to have 75:25%
                if(maxentarguments == TRUE){
                  xm <- maxent(x = predictorvariables,p = convertxy[convertxy$kfold!=1,], a = bgpnts@coords,
                           args=c("noautofeature","noproduct","nothreshold"))
                } else {
                  xm <- maxent(x = predictorvariables,p = convertxy[convertxy$kfold!=1,], a = bgpnts@coords)
                }
                
                write.csv(data.frame(convertxy@coords,convertxy@data),
                          paste(pathstart,"presenceHerb",filenames,"Sp",x,"kfold",rep,".csv", sep=""))
                save(xm, file= paste(pathstart,"maxentHerb",filenames,"Sp",x,"kfold",rep,".Rda", sep=""))        
        
                gc()
                #register parallel computing backend
                cl = parallel::makeCluster(ncores)
                doParallel::registerDoParallel(cl,ncores)
                #compute indices for data splitting
                rows = 1:nrow(predictorvariables)
                split = sort(rows%%ncores)+1
                outname = paste(pathstart,"PredictHerb",filenames,"Sp", x,"kfold",rep, sep="")
                #perform the prediction on subsets of the predictor dataset
                foreach(i=unique(split), .combine=c)%dopar%{
                  rows_sub = rows[split==i]
                  sub = raster::crop(predictorvariables,raster::extent(predictorvariables, min(rows_sub), max(rows_sub), 
                                                                1, ncol(predictorvariables)))
                  raster::predict(sub, xm, filename=paste(outname, i, sep="_"), overwrite=TRUE)
                }
        
                e <- evaluate(convertxy[convertxy$kfold==1,], bgpnts, xm, predictorvariables)
                save(e, file= paste(pathstart,"evaluateHerb",filenames,"Sp",x,"kfold",rep,".Rda", sep=""))
        
                rm(xm)
                gc()
                stopCluster(cl)
          }
                e
        }
}

```


```{r}
# x<-atleast12[3]
for(x in atleast12168[21]){
            pointsspdf <- SpatialPointsDataFrame(coords = spatialpointsdataframe_herb[[x]][,c("decimalLongitude","decimalLatitude")],
                                                 data = spatialpointsdataframe_herb[[x]],
                                                 proj4string = CRS("+proj=utm +zone=13 ellps=NAD83 +ellps=WGS84"))
          # if(error == TRUE){
               # for each point I will draw a circle of size (drawn from the distribution of error seen distXspall$Dist) and then pick a random point along the circle.
              errorpointsout <- do.call(rbind,lapply(1:nrow(pointsspdf), function(r){
                errordist <- sample(distdistribution, 1)
                if(errordist>0){
                  erroraround <- gBuffer(pointsspdf[r,], width=errordist)
                  newpoint <- erroraround@polygons[[1]]@Polygons[[1]]@coords
                  out <- newpoint[sample(1:nrow(newpoint),1),]
                } else {
                  out <- pointsspdf@coords[r,]
                  }
                out
                }
                ))
                
                df <- SpatialPointsDataFrame(coords = errorpointsout,
                                             data = pointsspdf@data,
                                             proj4string = CRS("+proj=utm +zone=13 ellps=NAD83 +ellps=WGS84"))
                circlesout <- circles(df, d = 5000) #Should be 5km around
                polygns <- polygons(circlesout)
                bgpnts <- spsample(polygns, 300, "stratified") # one single random location in each 'cell' 
                
                convertxy <- spTransform(df, CRS("+proj=longlat +datum=WGS84"))
                proj4string(bgpnts) <- CRS("+proj=utm +zone=13 ellps=NAD83 +ellps=WGS84")
                bgpnts <- spTransform(bgpnts, CRS("+init=epsg:4326") )
          # } else {
          #       convertxy <- spTransform(pointsspdf, CRS("+proj=longlat +datum=WGS84"))
          #       # bgpnts <- background_points[[match(x,notnull)]]
          #       circlesout <- circles(pointsspdf, d = 5000)
          #       polygns <- polygons(circlesout)
          #       bgpnts <- spsample(polygns, 300, "stratified")
          #       proj4string(bgpnts) <- CRS("+proj=utm +zone=13 ellps=NAD83 +ellps=WGS84")
          #       bgpnts <- spTransform(bgpnts, CRS("+init=epsg:4326") )
          #   
          # }
                
          # for(rep in 1:4){
                convertxy$kfold <- kfold(convertxy, k=4) # to have 75:25%
                # if(maxentarguments == TRUE){
                #   xm <- maxent(x = predictorvariables,p = convertxy[convertxy$kfold!=1,], a = bgpnts@coords,
                #            args=c("noautofeature","noproduct","nothreshold"))
                # } else {
                  # xm <- maxent(x = predictorvariables,p = convertxy[convertxy$kfold!=1,], a = bgpnts@coords)
                # }                
          plot(colocounties) # , xlim=c(139982.4, 773240.8), ylim=c(408691.88, 4761362.70))
          points(bgpnts, pch=16, col="purple", cex=0.5)
          points(spTransform(pointsspdf, CRS("+init=epsg:4326")), pch=21, col="orange")

          # }
}

```

# More functions 

```{r}

habitatSpecificity <- function(whichones, pathstart, whichmaxentpoints, replicates){
  habsp <- do.call(rbind,lapply(whichones, function(x){
  out <- do.call(rbind,lapply(1:replicates, function(rep){
    load(paste(pathstart,whichmaxentpoints,"Sp", x,"kfold",rep, ".Rda",sep=""))
    outinner <- data.frame(SpeciesNum = x, kfold = rep, xm@presence)
    outinner
    }))
  out
  }))
  habsp
  
  prPCA <- princomp(habsp[,-c(1:2)])
  allsphabsp <- data.frame(habsp[,1:2],prPCA$scores)

  # for all species
  habspecificity <- do.call(rbind,lapply(atleast12, function(x){
    # get and average across the replicates
    ellipseout <- do.call(rbind,lapply(1:10, function(rep){
      p <- ggplot(allsphabsp[allsphabsp$SpeciesNum==x&allsphabsp$kfold==rep,], 
                  aes(Comp.1,Comp.2))+
        geom_point()+
        stat_ellipse(segments=201) #default is to draw 51 line segments to make the ellipse
      # get ellipse coordinates
      pb <- ggplot_build(p)
      table(pb$data[[2]]$group)
      
      el <- pb$data[[2]][c("x","y")]
      
      # Center of ellipse
      ctr <- MASS::cov.trob(el)$center
      
      # Distance to center from each point on ellipse
      dist2center <- sqrt(rowSums((t(t(el)-ctr))^2))
      
      # Area of ellipse from semi-major and semi-minor axes which are largest and smallest of dist to center
      habitatspecificity <- pi*min(dist2center)*max(dist2center)
      hsout <- data.frame(SpeciesNum = x, kfold = rep, habspec = habitatspecificity)
      hsout
      }))
    ellipseout
    }))
  habspecificity
}
```




# Maxent    
Get the years of the herbarium specimens and the years of the EORs mapped from distXsp_noEORgrid from "Herb to EORs_v4"
```{r}
# All are fine, should work. 
EORpolys <- lapply(1:68, function(i){
  polys <- l1G1G2and[l1G1G2and$GNAME %in% c(g1g2namesall68$AcceptedName[i],
                                        g1g2namesall68$Taxon[i]),]
  polys
})

notnull1_68 <- which(!unlist(lapply(distXsp_noEORgrid1_68, is.null)))
length(notnull1_68) # 52 have data

whichnull <- lapply(1:68, function(i){
  gc()
  g1g2now <- coloradosps.g1g2_168[coloradosps.g1g2_168$scientificName %in%
                                c(g1g2namesall68$Taxon[i],g1g2namesall68$AcceptedName[i])&
                                !is.na(coloradosps.g1g2_168$decimalLatitude),]
  if(nrow(g1g2now)>0){
    g1g2namesall68$AcceptedName[i]
  }
})
  
notnulls68 <- which(!unlist(lapply(whichnull, is.null)))
which(unlist(lapply(whichnull, is.null))) # 39
EORpolys[[39]] # Oonopsis sp. 1, hasn't been published yet!


# See Herb to EORs_v4 for notnulls68
length(notnull1_68) # 52
length(distXsp_noEORgrid1_68) # 68
length(notnulls68) # 67
length(notnulls68[notnull1_68]) # 52

which(!unlist(lapply(distXsp_noEORgrid1_68, is.null)))

distXspall <- do.call(rbind, Map(cbind, SpNum = which(!unlist(lapply(distXsp_noEORgrid1_68, is.null))), 
                                   distXsp_noEORgrid1_68[which(!unlist(lapply(distXsp_noEORgrid1_68, is.null)))]))
unique(distXspall[,c("SpNum","Species")])

distdistribution <- distXspall52$Dist[distXspall52$Dist<300000]

df_dist <- distXspall

df_dist$EORYear <- as.numeric(substring(df_dist$EORdate,1,4))
hist(df_dist$EORYear)

plot(log(df_dist$Area+1),log(df_dist$Dist+1),
     xlab=expression(paste("Species Range log(meters" ^2,")" )),
     ylab=expression(paste("log (meters)")))
abline(lm(log(Dist+1)~log(Area+1), data=df_dist))
length(unique(df_dist$SpNum))
max(df_dist$Dist) # 411023.6 meters

atleast12168 <- c()
for(l in notnull1_68){  # notnull1_68
  atleast12168[match(l,notnull1_68)] <- if(nrow(distXsp_noEORgrid1_68[[l]])>11){   # match(l,notnull1_68)
    l } else {
      NA
    }
}
atleast12168 <- atleast12168[!is.na(atleast12168)] 
```

Table 1 - species in niche models, put a column indiciating which are SDM
```{r}

out <- Rmisc::summarySE(df_dist, "Dist", "SpNum")

tab1 <- do.call(rbind, lapply(split(df_dist, df_dist$SpNum), function(x){
  data.frame(SpNum1 = unique(x$SpNum), Species = x$Species[1], DistAvg = summarySE(x, "Dist"), 
             Area = unique(x$Area), minYearDiff = min(x$Yeardiff), maxYearDiff = max(x$Yeardiff),
             SpNum= unique(x$SpNum), YearDiffAvg = summarySE(x, "Yeardiff"),
             MinEORyear = min(x$EORYear), MaxEORyear = max(x$EORYear), 
             MinYear = min(x$year), MaxYear = max(x$year), AvgLat = summarySE(x, "decimalLatitude"))
})) 



tab1$SDM[tab1$SpNum %in% atleast12168] <- "x"

write.csv(tab1, "Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Images/table1.csv")

library(Taxonstand)
TPLtbl1 <- TPL(tab1$Species)
write.csv(TPLtbl1, "Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Images/TPLtbl1.csv")

```

# Still need 29, 40; delete 65 and 61 once the correct ones go 62 and 66!!
c(29,40,62,66)


```{r}
jpeg("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Images/Figure2.jpg",
     width=225, height=105,units='mm', res=300)

layout(matrix(c(1,2), 1,2))

#default margins par(mar=c(5.1, 4.1, 4.1, 2.1))

par(mar=c(5.1, 4.1, 1.1, 0.1))
plot(log(df_dist$Area+1),
     log(df_dist$Dist+1),
     xlab=expression(paste("Species Range log(km" ^2,")")),
     ylab=expression(paste("log(km)")),
     pch=16, cex= .25) 
abline(lm((log(Dist+1))~ (log(Area+1)), data=df_dist))  # 
mtext("a)", side=3, line=0, adj=0)

par(mar=c(5.1, 2.1, 1.1, 0.1))
plot(jitter(df_dist$year),log(df_dist$Dist+1),
     xlab="Year",
     ylab="",
     pch=16, cex= .25)
abline(lm(log(Dist+1)~year, data=df_dist))
mtext("b)", side=3, line=0, adj=0)


dev.off()

```


The EOR AUC results from VisTrails, samesize runs from ForVisTrails.Rmd where I either thinned to match number of Herbarium records or where I bootstrapped and jittered to get matching points. Used a standalone Dan Warren thinning process to get equally spaced out but covering the most area (of information) for points.


Make maxent commands
for output prediction map, use ENMeval   
Using KDE at 95% in MAxent VisTrails which mostly makes it so that it appears there's less effort in areas with fewer points but might not be the case, just might be smaller popualtions so maybe just circles with a certain radius would be better. Maybe add some points along roads for circles. Could put points along roads and then first pick points along roads within a certain distance from presence points and then 'circles' around all those points ?  
follow https://rspatial.org/sdm/SDM.pdf while VisTrails isnt' working

Background points, in a 5km circle around presence points, 300 stratified across the circles. Not quite the KDE but if I combined other collections from same trip I think that would be a correct estimator, circles represent the posibility that folks hiked 3 miles on their collection trip, seems likely enough. 

```{r}

yearsofPolys <- lapply(atleast12168, function(i){
    polys <- l1G1G2and [l1G1G2and $GNAME %in% c(g1g2namesall68$AcceptedName[i],
                                        g1g2namesall68$Taxon[i]),] 
    proj4string(polys) <- CRS("+proj=utm +zone=13 ellps=NAD83 +ellps=WGS84") # CRS("+init=epsg:26913")
    # years <- format(as.Date(polys$LASTOBS, format="%Y-%m-%d"),"%Y")
    years <- as.numeric(substring(polys$LASTOBS, 1,4))
    if(length(years)>0){
    yearlater1980 <- c()
    for(l in 1:length(years)){
      yearlater1980[l] <- if(years[l]>1980){ 
        l } else {
          NA
        }
    }
    yearlater1980 <- yearlater1980[!is.na(yearlater1980)]  
    table(years[which(years>1980)])
    }
})
for(d in 1:length(yearsofPolys)) print(yearsofPolys[[d]])



```



## Herbarium without more error, 10 replicates   
```{r}

jar <- paste(system.file(package="dismo"), "/java/maxent.jar", sep='')
ncores=10
    
atleast12168

# notes in NicheSpaialIssuesManuscript_PCA from Merow et al. 2013 and ENMTools
# Set parameters for maxent niche models
argsformax <- make.args(RMvalues = 0, # regularization multiplier, seq(0.5, 4,0.5) to test multiple values?
                  fc = "LQH")  # features class combinations; fc = c("L", "LQ", "H", "LQH", "LQHP", "LQHPT"),)

# for(x in atleast12){
#         convertxy <- spTransform(distXsp_v3[[x]][[1]], CRS("+proj=longlat +datum=WGS84"))
#         bgpnts <- bgHerb[[match(x,notnull)]]
#         proj4string(bgpnts) <- CRS("+proj=utm +zone=13 ellps=NAD83 +ellps=WGS84")
#         bgpnts <- spTransform(bgpnts, CRS("+init=epsg:4326") )
#         
#   for(rep in 1:10){
#         convertxy$kfold <- kfold(convertxy, k=4) # to have 75:25%
#         xm <- maxent(x = rasterstack,p = convertxy[convertxy$kfold!=1,], a = bgpnts@coords) # , 
#                     # args=c("noautofeature","noproduct","nothreshold"))
#         write.csv(data.frame(convertxy@coords,convertxy@data),
#                   paste("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/RMaxEnt_herb_herbpluserror/presenceHerbnoerrorSp",x,"kfold",rep,".csv", sep=""))
#         save(xm, file= paste("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/RMaxEnt_herb_herbpluserror/maxentHerbnoerrorSp",x,"kfold",rep,".Rda", sep=""))        
# 
#         gc()
#         #register parallel computing backend
#         cl = parallel::makeCluster(ncores)
#         doParallel::registerDoParallel(cl,ncores)
#         #compute indices for data splitting
#         rows = 1:nrow(rasterstack)
#         split = sort(rows%%ncores)+1
#         outname = paste("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/RMaxEnt_herb_herbpluserror/PredictHerbnoerrorSp", x,"kfold",rep, sep="")
#         #perform the prediction on subsets of the predictor dataset
#         foreach(i=unique(split), .combine=c)%dopar%{
#           rows_sub = rows[split==i]
#           sub = raster::crop(rasterstack,raster::extent(rasterstack, min(rows_sub), max(rows_sub), 
#                                                         1, ncol(rasterstack)))
#           raster::predict(sub, xm, filename=paste(outname, i, sep="_"), overwrite=TRUE)
#         }
# 
#         e <- evaluate(convertxy[convertxy$kfold==1,], bgpnts, xm, rasterstack)
#         save(e, file= paste("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/RMaxEnt_herb_herbpluserror/evaluateHerbnoerrorSp",x,"kfold",rep,".Rda", sep=""))
# 
#         rm(xm)
#         gc()
#         stopCluster(cl)
#   }
#         e
# }
# 

```

Still missing 29 with error, 40 no error
```{r}
# from 1-60
notnull
atleast12

# from testing 1-68
atleast12168

pathstart6266 <- "Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/RMaxEnt_herb_herbpluserror/" 
spatialpointsdataframe_herb <- distXsp_noEORgrid1_68
hist(distdistribution) # ca line 300

#29 worked 40 can't get all the way through!
maxentrun(whichones = c(62,66), numberofReps = 10, 
          spatialpointsdataframe_herb = spatialpointsdataframe_herb,
          maxentarguments = FALSE,  filenames = "noerror", 
          pathstart = pathstart6266)

# c(3,62,66)
# c(29,40,62,66)
maxentrun(whichones = 29, numberofReps = 10, 
          spatialpointsdataframe_herb = spatialpointsdataframe_herb,
          maxentarguments = FALSE,  filenames = "WITHerror", error = TRUE,
          pathstart = pathstart6266, distdistribution = distdistribution)

# Errors for 66, maybe error put them too far away?? No, for some reason sometimes x,y sometimes lat,lon
maxentrun(whichones = 66, numberofReps = 10, 
          spatialpointsdataframe_herb = spatialpointsdataframe_herb,
          maxentarguments = FALSE,  filenames = "WITHerror", error = TRUE,
          pathstart = pathstart6266, distdistribution = distdistribution)

# (whichones, pathstart, patternmatch, rasternames) where rasternames is what the written raster will be called beyond the Sp # from whichones




stitchtogether(c(29,40,62,66), pathstart6266, "PredictHerbnoerrorSp", "Herbnoerrorkfold")
stitchtogether(3, pathstart6266, "PredictHerbnoerrorSp", "Herbnoerrorkfold")
stitchtogether(c(3,62,66), pathstart6266, "PredictHerbWITHerrorSp", "HerbWITHerrorkfold")
stitchtogether(66, pathstart6266, "PredictHerbWITHerrorSp", "HerbWITHerrorkfold")
stitchtogether(3, pathstart6266, "PredictHerbWITHerrorSp", "HerbWITHerrorkfold")

# Accounted for the missing on in the function, can do the ones I really want 
# 31? atleast12168[c(17:18,21:24,28)]  
atleast12168[28]
# 40 won't work
forhistofaverage(whichones = c(62,66), pathstart = pathstart6266, patternmatch = "Herbnoerrorkfold")
forhistofsd(whichones = c(29,40,62,66), pathstart = pathstart6266, patternmatch = "Herbnoerrorkfold")


forhistofaverage(whichones = c(40,62,66), pathstart = pathstart6266, patternmatch = "HerbWITHerrorkfold")
forhistofsd(whichones = c(29,40,62,66), pathstart = pathstart6266, patternmatch = "HerbWITHerrorkfold")

maxentrun(whichones = atleast12168[c(17:18,21:24,28)], numberofReps = 10, 
          spatialpointsdataframe_herb = spatialpointsdataframe_herb,
          maxentarguments = FALSE,  filenames = "noerror", 
          pathstart = pathstart6266)

# atleast12168[c(17:18,21:24,28)]
maxentrun(whichones = c(29,40,62,66), numberofReps = 10, 
          spatialpointsdataframe_herb = spatialpointsdataframe_herb,
          maxentarguments = FALSE,  filenames = "WITHerror", error = TRUE,
          pathstart = pathstart6266, distdistribution = distdistribution)

```

```{r}
### come back here, run these!!! 
maxentrun(whichones = atleast12168[c(22:24,28)] , numberofReps = 10, 
          # atleast12168[c(17:18,21:24,28)]; 40
          spatialpointsdataframe_herb = spatialpointsdataframe_herb,
          maxentarguments = FALSE,  filenames = "noerror", 
          pathstart = pathstart6266)

# 43 had errors
# atleast12168[c(17:18,21:24,28)]
maxentrun(whichones =  atleast12168[c(24,28)] , numberofReps = 10, 
          spatialpointsdataframe_herb = spatialpointsdataframe_herb,
          maxentarguments = FALSE,  filenames = "WITHerror", error = TRUE,
          pathstart = pathstart6266, distdistribution = distdistribution[distdistribution<150000])

maxentrun(whichones =  40, numberofReps = 10, 
          spatialpointsdataframe_herb = spatialpointsdataframe_herb,
          maxentarguments = FALSE,  filenames = "WITHerror", error = TRUE,
          pathstart = pathstart6266, distdistribution = distdistribution[distdistribution<150000])


hist(distdistribution[distdistribution<200000])

stitchtogether(atleast12168[c(17:18,21:24,28)], pathstart6266, "PredictHerbnoerrorSp", "Herbnoerrorkfold")
stitchtogether(atleast12168[c(17:18,21:24,28)], pathstart6266, "PredictHerbWITHerrorSp", "HerbWITHerrorkfold")

forhistofaverage(whichones = atleast12168[c(17:18,21:24,28)], pathstart = pathstart6266, patternmatch = "Herbnoerrorkfold")
forhistofsd(whichones = atleast12168[c(17:18,21:24,28)], pathstart = pathstart6266, patternmatch = "Herbnoerrorkfold")

# atleast12168[c(17:18,21:24,28)]
forhistofaverage(whichones = 3, pathstart = pathstart6266, patternmatch = "HerbWITHerrorkfold")
forhistofsd(whichones = 3, pathstart = pathstart6266, patternmatch = "HerbWITHerrorkfold")
###############
```



```{r}


###Testing maxent parameters
i<-1
k<-3
convertxy$kfold <- kfold(convertxy, k=4) # to have 75:25%
xm <- maxent(x = rasterstack,p = convertxy[convertxy$kfold!=1,], a = bgpnts@coords,args=c("noautofeature","noproduct","nothreshold"))
        save(xm, file= paste("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/RMaxEnt_herb_herbpluserror/maxentHerbnoerrorSp",x,"kfold",rep,".Rda", sep=""))        

        gc()
        #register parallel computing backend
        cl = parallel::makeCluster(ncores)
        doParallel::registerDoParallel(cl,ncores)
        #compute indices for data splitting
        rows = 1:nrow(rasterstack)
        split = sort(rows%%ncores)+1
        outname = paste("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/RMaxEnt_herb_herbpluserror/PredictHerbnoerrorSp", x,"kfold",rep, sep="")
        #perform the prediction on subsets of the predictor dataset
        foreach(i=unique(split), .combine=c)%dopar%{
          rows_sub = rows[split==i]
          sub = raster::crop(rasterstack,raster::extent(rasterstack, min(rows_sub), max(rows_sub), 
                                                        1, ncol(rasterstack)))
          raster::predict(sub, xm, filename=paste(outname, i, sep="_"), overwrite=TRUE)
        }
  gc()

    resultpath <- list.files(path = pathstart, 
                             pattern = paste("PredictHerbnoerrorSp",i,"kfold",k,"_",sep=""), 
                             full.names=TRUE)
    rastout <- lapply(resultpath, function(x){
      raster(x)
      })
    rastout$filename <- paste("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/RMaxEnt_herb_herbpluserror/ProbTiffSp",i,"Herbnoerrorkfold",k,".tif", sep="")
    rastout$overwrite <- TRUE
    m <- do.call(merge, rastout)
plot(m)
plot(mbackgroundnoargs)
# mbackgroundnoargs <- m
        rm(xm)
        gc()
        stopCluster(cl)
```


Stitch together the maps 
```{r}
pathstart <- ("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/RMaxEnt_herb_herbpluserror/")

  lapply(atleast12, function(i){
  gc()
  lapply(1:10, function(k){
    resultpath <- list.files(path = pathstart, 
                             pattern = paste("PredictHerbnoerrorSp",i,"kfold",k,"_",sep=""), 
                             full.names=TRUE)
    rastout <- lapply(resultpath, function(x){
      raster(x)
      })
    rastout$filename <- paste("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/RMaxEnt_herb_herbpluserror/ProbTiffSp",i,"Herbnoerrorkfold",k,".tif", sep="")
    rastout$overwrite <- TRUE
    m <- do.call(merge, rastout)
  })
})
```

MAxent and stitch together rasters as functions
```{r}

for(x in atleast12){
        convertxy <- spTransform(distXsp_v3[[x]][[1]], CRS("+proj=longlat +datum=WGS84"))
        bgpnts <- bgHerb[[match(x,notnull)]]
        proj4string(bgpnts) <- CRS("+proj=utm +zone=13 ellps=NAD83 +ellps=WGS84")
        bgpnts <- spTransform(bgpnts, CRS("+init=epsg:4326") )
        
  for(rep in 1:10){
        convertxy$kfold <- kfold(convertxy, k=4) # to have 75:25%
        xm <- maxent(x = rasterstack,p = convertxy[convertxy$kfold!=1,], a = bgpnts@coords) # , 
                    # args=c("noautofeature","noproduct","nothreshold"))
        write.csv(data.frame(convertxy@coords,convertxy@data),
                  paste("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/RMaxEnt_herb_herbpluserror/presenceHerbnoerrorSp",x,"kfold",rep,".csv", sep=""))
        save(xm, file= paste("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/RMaxEnt_herb_herbpluserror/maxentHerbnoerrorSp",x,"kfold",rep,".Rda", sep=""))        

        gc()
        #register parallel computing backend
        cl = parallel::makeCluster(ncores)
        doParallel::registerDoParallel(cl,ncores)
        #compute indices for data splitting
        rows = 1:nrow(rasterstack)
        split = sort(rows%%ncores)+1
        outname = paste("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/RMaxEnt_herb_herbpluserror/PredictHerbnoerrorSp", x,"kfold",rep, sep="")
        #perform the prediction on subsets of the predictor dataset
        foreach(i=unique(split), .combine=c)%dopar%{
          rows_sub = rows[split==i]
          sub = raster::crop(rasterstack,raster::extent(rasterstack, min(rows_sub), max(rows_sub), 
                                                        1, ncol(rasterstack)))
          raster::predict(sub, xm, filename=paste(outname, i, sep="_"), overwrite=TRUE)
        }

        e <- evaluate(convertxy[convertxy$kfold==1,], bgpnts, xm, rasterstack)
        save(e, file= paste("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/RMaxEnt_herb_herbpluserror/evaluateHerbnoerrorSp",x,"kfold",rep,".Rda", sep=""))

        rm(xm)
        gc()
        stopCluster(cl)
  }
        e
}
################################# for loop works but lapply and function do not work


pathstart <- "Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/RMaxEnt_herb_herbpluserror/"
stitchtogether(atleast12[1:21], pathstart,"PredictHerbWITHerrorSp", "HerbWITHerrorkfold")
```


# Use the function to run maxent with or without arguments   
Don't know how to make the function act like the for loop where lapply didnt' send info out of function and similarly the for loop within the function doesn't send the information out
```{r}

pathstart2 <- "Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/RMaxEnt_witharguments/" 
spatialpointsdataframe_herb <- distXsp_v3

maxentrun(whichones = atleast12[1], numberofReps = 2, spatialpointsdataframe_herb = spatialpointsdataframe_herb,
          maxentarguments = TRUE,  filenames = "noFooerror", 
          pathstart = pathstart2)

```



Take the overlap of all EOR polygons and the rasters to get the histograms of probability   
    1. Do I average over all the replicates? Then what do I do with the SD? That could also be extracted to see how much uncertainty falls within the spatial extent of the known populations. - that is another measure of precision.    
    2. I should get the EORs from Utah, Wyoming, Arizona?, and New Mexico   
    
Accuracy
```{r}
load("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/g1g2names.Rda")
library(maptools)
l1eor <- readShapePoly("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/L1shp")
l1G1G2 <- disaggregate(l1eor[l1eor$GRANK %in% c("G1","G2"),])
l1G1G2$PolyID <- do.call(rbind, lapply(split(l1G1G2,l1G1G2$OBJECTID),
                                       function(x){
              x$POlyID <- paste(x$EO_ID, LETTERS[seq(from=1, to= nrow(x))],sep="")
              x
              }))
plot(l1G1G2)



pathstartwithnoerror <- "Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/RMaxEnt_herb_herbpluserror/"
patternmatchno <- "Herbnoerrorkfold"
patternmatchwith <- "HerbWITHerrorkfold"

noerroravg <- forhistofaverage(pathstartwithnoerror, patternmatchno)
witherroravg <- forhistofaverage(pathstartwithnoerror, patternmatchwith)

# Extract takes the raster value only when the center of the raster is covered by the polygon. I've got lots of polygons that do not cover the center of the larger raster cells. Use weights for partialy covered cells and argument small for getting values for small polygons. 
# get map, make lists of 10 each for each replicate
accuracyhists <- function(pathstart, probmapnames, whichones, reps = 1:10){
  lapply(whichones, function(i){
  lrhist <- lapply(reps, function(k){
    r <- raster(paste(pathstart,"ProbTiffSp",i,probmapnames,k,".tif", sep=""))
    
    polys <- l1G1G2[l1G1G2$GNAME %in% c(g1g2names$AcceptedName[i],
                                    g1g2names$Taxon[i]),] 
    # Only want ones after 1980
    years <- as.numeric(substring(polys$LASTOBS,1,4))
    yearlater1980 <- c()
    for(l in 1:length(years)){
      yearlater1980[l] <- if(years[l]>1980){
                                      l } else {
                                        NA
                                      }
    }
    yearlater1980 <- yearlater1980[!is.na(yearlater1980)]
    polys <- polys[yearlater1980,]
    
    proj4string(polys) <- CRS("+proj=utm +zone=13 ellps=NAD83 +ellps=WGS84") # CRS("+init=epsg:26913")
    
    f <- spTransform(polys, CRS("+proj=longlat +datum=WGS84"))   
    fr <- extract(r, f, small = TRUE, weights = TRUE)
    # Want the value [,1] for each polygon but use the weight [,2] to get the pro-rated cells 
    out <- do.call(rbind,fr)
    out
    })
  lrhist
  })
}


accuracyhistsOneperSp <- function(pathstart, whichones =  atleast12[atleast12!=8], SDorAVG = "Avg", WithorNo = "WITH"){
  lapply(whichones, function(i){
    resultpath <- list.files(path = pathstart, 
                         pattern = paste(SDorAVG,"TiffSp",i,"Herb",WithorNo, sep=""), 
                         full.names=TRUE)
    
    r <- raster(resultpath)
    polys <- l1G1G2and[l1G1G2and$GNAME %in% c(g1g2namesall68$AcceptedName[i],
                                              g1g2namesall68$Taxon[i]),] 

    # Only want ones after 1980
    years <- as.numeric(substring(polys$LASTOBS,1,4))
    yearlater1980 <- c()
    for(l in 1:length(years)){
      yearlater1980[l] <- if(years[l]>1980){
                                      l } else {
                                        NA
                                      }
    }
    yearlater1980 <- yearlater1980[!is.na(yearlater1980)]
    polys <- polys[yearlater1980,]
    
    proj4string(polys) <- CRS("+proj=utm +zone=13 ellps=NAD83 +ellps=WGS84") # CRS("+init=epsg:26913")
    
    f <- spTransform(polys, CRS("+proj=longlat +datum=WGS84"))   
    fr <- extract(r, f, small = TRUE, weights = TRUE)
    # Want the value [,1] for each polygon but use the weight [,2] to get the pro-rated cells 
    out <- do.call(rbind,fr)
    out
    })
}

pathstart1 <- "Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/RMaxEnt_herb_herbpluserror/"
# probmapnames1 <- "Herbnoerrorkfold"
# whichones1 <- atleast12[atleast12!=8]

noerrorhists <- accuracyhists(pathstart1, probmapnames1, whichones1)
hist(noerrorhists[[1]][[2]])

witherrorhists <- accuracyhists("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/RMaxEnt_herb_herbpluserror/", "HerbWITHerrorkfold", whichones1)

# reallistofTiffs <- atleast12[atleast12!=3]
reallistofTiffs <- atleast12168[!atleast12168 %in% c(29,40)]
# reallistofTiffs <- reallistofTiffs[!reallistofTiffs %in% c(31,32,40)]

#Averaged and SD of probability maps second function
noerrorhistsofavg <- accuracyhistsOneperSp(whichones = reallistofTiffs, pathstart1, WithorNo = "no")
witherrorhistsofavg <- accuracyhistsOneperSp(whichones = reallistofTiffs,pathstart1)


# kstest are they different or the same,
# then 
# lognormal distribution the median and geometric mean are identical. 
# Oh, the log(mn) =  log(m) + log(n) where log(m/n) =  log(m) - log(n) and log(m^2) = n * log(m)

# Check sum of logs first.. 
# plots same or different 
# AICc has wrong parameters

save(noerrorhistsofavg, file = "Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/RMaxEnt_herb_herbpluserror/noerrorhistsofavg.Rdata")

save(witherrorhistsofavg, file = "Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/RMaxEnt_herb_herbpluserror/witherrorhistsofavg.Rdata")
```


```{r}
ks.outs <- lapply(1:length(noerrorhistsofavg), function(i){
  ks.test(log(data.frame(noerrorhistsofavg[[i]])$value)+
             log(data.frame(noerrorhistsofavg[[i]])$weight),
          log(data.frame(witherrorhistsofavg[[i]])$value)+
             log(data.frame(witherrorhistsofavg[[i]])$weight))
})

length(notnulls68) # 67

sapply(ks.outs, '[',2)

whichsigndiff <- sapply(1:length(noerrorhistsofavg), function(s){
  if(ks.outs[[s]]$p.value < 0.05){
    1
  } else {
    NA
  }
})



whichsigndiff
which(!unlist(lapply(whichsigndiff, is.na)))
length(which(!unlist(lapply(whichsigndiff, is.na)))) # 19
length(ks.outs) # 33

# install.packages("latticeExtra")
# library(latticeExtra)
# Default margins
par(mar=c(5.1, 4.1, 4.1, 2.1))

# https://chemicalstatistician.wordpress.com/2013/06/24/exploratory-data-analysis-conceptual-foundations-of-empirical-cumulative-distribution-functions/  
```

```{r}

jpeg("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Images/Figure4_accuracy.jpg",
     width=500, height=600,units='mm', res=300)

layout(matrix(c(1:20),nrow=5, byrow = TRUE))

par(mar=c(5.1,4.1,4.1,0.1))
for(i in which(!unlist(lapply(whichsigndiff, is.na)))){
  as_is <- ecdf(log(data.frame(noerrorhistsofavg[[i]])$value)+
             log(data.frame(noerrorhistsofavg[[i]])$weight))
  additional  <- ecdf(log(data.frame(witherrorhistsofavg[[i]])$value)+
             log(data.frame(witherrorhistsofavg[[i]])$weight))
  plot(as_is, col = "blue", main=paste(g1g2namesall68$AcceptedName[reallistofTiffs[i]]))  
  lines(additional, col = rgb(0,1,0,0.5))
}


dev.off()
```

# try ks plot kernel smoothing, smothing optimized for display
# plot ecdf, the sum of logs 
?ecdf

Compare those with and not with different accuracy curves for distances and sample size and area  
```{r}

# The SpNum that were compared, that had SDM for both as-is and with error
reallistofTiffs

# of reallistofTiffs the ones that are significantly different 
whichsigndiff


df_dist
df_dist$AccuracyDifferent <- "No"
df_dist$AccuracyDifferent[df_dist$SpNum %in% reallistofTiffs[which(!unlist(lapply(whichsigndiff, is.na)))]] <- "Yes"

N_after1980<-sapply(split(df_dist, df_dist$SpNum), function(x){
  nrow(x[x$year>1980,])
})

SampleSize <- data.frame(SpNum = as.numeric(names(N_after1980)), N_SDM = N_after1980)

df_dist[df_dist$SpNum >60,] # What da, I have ErBr as 63 and Lygodesmia grandiflora as 67 not 62 and 66. Ugg, but otherwise correct

df_dist2 <- merge(df_dist)
summary(lm(AccuracyDifferent ~ Area, data=df_dist$))


```



Compare distributions that are not the same using performR from Silas Tittes
```{r}
#install
devtools::install_github("silastittes/performr", local = FALSE)

#load other libraries used below
library(performr)
library(tidyverse)
library(ggridges)
theme_set(theme_minimal())

```

# run my own Bayesian analysis to look at the differeces in distirbutions by comparing moments (where \mu = (alpha/(alpha+beta))) and variance \sigma^2 = (alpha*beta)/((alpha+beta)^2(alpha+beta+1))     
<https://www4.stat.ncsu.edu/~reich/st590/code/BetaBinomJAGS>   
```{r}
#install.packages("R2jags")
library(R2jags)
library(rjags)
library(coda)

dat <- data.frame(x = log(data.frame(noerrorhistsofavg[[1]])$value), 
                  y = log(data.frame(noerrorhistsofavg[[1]])$value)+
                    log(data.frame(noerrorhistsofavg[[1]])$weight) )



```

# Precision
```{r}
# pathstart <- "Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/RMaxEnt_herb_herbpluserror/"
precisionplots <- lapply(c(reallistofTiffs,61,65), function(i){
                        resultpath <- list.files(path = pathstart, 
                                                 pattern = paste("AvgTiffSp",i,"Herbnoerror",sep=""), 
                                                 full.names=TRUE)
                        rastout <- raster(resultpath)
                        values(rastout)
                        })

# takes a long time, what if I subsample the points, does it go faster without losing information?
plot(ecdf(precisionplots[[1]]))

# it is 43,740,000 long
plot(ecdf(sample(precisionplots[[1]], 100000)))

# data:  precisionplots[[1]] and sample(precisionplots[[1]], 1e+06)
# D = 0.00072771, p-value = 0.6785
# alternative hypothesis: two-sided
ks.test(precisionplots[[1]],
        sample(precisionplots[[1]], 1000000))

# data:  precisionplots[[1]] and sample(precisionplots[[1]], 1e+05)
# D = 0.0021953, p-value = 0.7221
# alternative hypothesis: two-sided
ks.test(precisionplots[[1]],
        sample(precisionplots[[1]], 100000))

for(i in 1:length(precisionplots)){
  plot(ecdf(sample(precisionplots[[i]],1000000)), 
       main = g1g2namesall68$AcceptedName[c(reallistofTiffs,61,65)[i]])
}

```




#  Habitat specificity    
Principal components analysis of predictor variables    
PCA for how different - 
calculate the size of the ellipse <https://stackoverflow.com/questions/38782051/how-to-calculate-the-area-of-ellipse-drawn-by-ggplot2>
```{r}
habsp_EOR <- do.call(rbind, lapply(atleast12, function(i){
  polys <- l1G1G2[l1G1G2$GNAME %in% c(g1g2names$AcceptedName[i],
                                        g1g2names$Taxon[i]),]
  proj4string(polys) <- CRS("+proj=utm +zone=13 ellps=NAD83 +ellps=WGS84")
  EORpnts <- do.call(rbind, lapply(1:length(polys), function(f){
    # area <- sapply(slot(polys,"polygons"), slot, "area")[f]
    # round(sqrt(area)/165,0) # meters along a 'side'; bioclim is at about 165 meters
      outline <- polys@polygons[[f]]@Polygons[[1]]@coords # Use these and grid within
      grid <- makegrid(polys, cellsize = 165) # cellsize in map units!
      names(grid) <- c("x","y")
      gridout <- grid[splancs::inout(grid,outline), ]
      if(nrow(gridout)==0){
        gridout <- data.frame(x=outline[,1],y=outline[,2])
      }
      gridout
      }))
  coordinates(EORpnts) <- ~x+y
  proj4string(EORpnts) <- CRS("+proj=utm +zone=13 ellps=NAD83 +ellps=WGS84")
  EORpnts <- spTransform(EORpnts, CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))
  gridextract <- extract(rasterstack,EORpnts)
    out <- data.frame(SpeciesNum=i, Species=g1g2names$AcceptedName[i], EORpnts@coords, gridextract)
    out
    }))


prPCAEOR <- princomp(habsp_EOR[,-c(1:4)])
summary(prPCAEOR)
loadings(prPCAEOR)

allsphabsp_EOR <- data.frame(habsp_EOR[,1:2],prPCAEOR$scores)

# for one run of one species
ggplot(allsphabsp_EOR, aes(Comp.1,Comp.2, colour=as.factor(SpeciesNum)))+
  geom_point(alpha=0.1)+
  stat_ellipse(segments=201)+ #default is to draw 51 line segments to make the ellipse
  theme_bw()+
  xlab(expression(atop("PC1 (97% variance explained)", "Elevation = 0.981; Bio12 = 0.193")))+
  ylab(expression(atop("PC2 (2% variance explained)","Aspect = -0.990; Bio12 = -0.139")))
  

# link to all the data about from table 1
habsp
specificitytotab1 <- merge(allsphabsp_EOR, tab1, by.x = "SpeciesNum", by.y = "SpNum1")


ggplot()


table(allsphabsp_EOR$SpeciesNum) # 8 has one point, 51 has 2

# for all species
habspecificity_EOR <- do.call(rbind,lapply(atleast12[!atleast12 %in% c(8, 51)], function(x){
  # get and average across the replicates
  p <- ggplot(allsphabsp_EOR[allsphabsp_EOR$SpeciesNum==x,],
              aes(Comp.1,Comp.2))+
          geom_point()+
          stat_ellipse(segments=201) #default is to draw 51 line segments to make the ellipse
  # get ellipse coordinates
  pb <- ggplot_build(p)
  table(pb$data[[2]]$group)
  el <- pb$data[[2]][c("x","y")]
  
  # Center of ellipse
  ctr <- MASS::cov.trob(el)$center
  
  # Distance to center from each point on ellipse
  dist2center <- sqrt(rowSums((t(t(el)-ctr))^2))
  
  # Area of ellipse from semi-major and semi-minor axes which are largest and smallest of dist to center
  habitatspecificity <- pi*min(dist2center)*max(dist2center)
  hsout <- data.frame(SpeciesNum = x, habspec = habitatspecificity)
  hsout
  }))

```




# Need to have both in same PCA!!! 
```{r}
pathstarterror <- "Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/RMaxEnt_herb_herbpluserror/"
maxones <- "maxentHerbnoerror"

# But these treat each separetly and I need them together in one getting the ellipse model!
habspsame <- habitatSpecificity(whichones = atleast12, pathstart = pathstarterror,
                                whichmaxentpoints = maxones, replicates = 10)

habspwitherror <- habitatSpecificity(whichones = atleast12, pathstart = pathstarterror,
                                whichmaxentpoints = "maxentHerbWITHerror", replicates = 10)
#####

pathstarterrornoerror <- "Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/RMaxEnt_herb_herbpluserror/"

### both at once in one PCA
# HABitat SPecificity with NO error
 habspno <- do.call(rbind,lapply(atleast12, function(x){
   outno <- do.call(rbind,lapply(1:10, function(rep){
      load(paste(pathstarterrornoerror,"maxentHerbnoerror","Sp", x,"kfold",rep, ".Rda",sep=""))
      outinner <- data.frame(ErrorType = "no", SpeciesNum = x, kfold = rep, xm@presence)
      outinner
      }))
   outno
   }))
 
 habspwith <- do.call(rbind,lapply(atleast12, function(x){
  outno <- do.call(rbind,lapply(1:10, function(rep){
    load(paste(pathstarterrornoerror,"maxentHerbWITHerror","Sp", x,"kfold",rep, ".Rda",sep=""))
    outinner <- data.frame(ErrorType = "yes", SpeciesNum = x, kfold = rep, xm@presence)
    outinner
    }))
  outno
  }))
 
 names(habsp_EOR)
 names(habsp)
  
habsp <- rbind(habspno,habspwith)
table(habsp$ErrorType)
habsp_andEOR <- rbind(habsp, 
                      data.frame(ErrorType = "EOR", SpeciesNum = habsp_EOR[,1], 
                                 kfold = 1, habsp_EOR[,c(5:9)]))

table(habsp_andEOR$ErrorType)

# prPCA <- princomp(habsp[,-c(1:3)])
# allsphabsp <- data.frame(habsp[,1:3],prPCA$scores)

prPCA <- princomp(habsp_andEOR[,-c(1:3)])
allsphabsp <- data.frame(habsp_andEOR[,1:3],prPCA$scores)

for(i in atleast12){
  print(paste(i,": ", nrow(allsphabsp[allsphabsp$SpeciesNum==i&allsphabsp$ErrorType=="EOR",])))
}

###BACK HERE TO FIX
# for all species and EOR
habspecificity <- do.call(rbind,lapply(atleast12[atleast12!=8], function(x){
  if(nrow(allsphabsp[allsphabsp$SpeciesNum==i&allsphabsp$ErrorType=="EOR",])>2){
        peor <- ggplot(allsphabsp[allsphabsp$SpeciesNum==x&allsphabsp$ErrorType=="EOR",], 
                      aes(Comp.1,Comp.2))+
            geom_point()+
            stat_ellipse(segments=201) #default is to draw 51 line segments to make the ellipse
          # get ellipse coordinates
          pbeor <- ggplot_build(peor)
          table(pbeor$data[[2]]$group)
          
          eleor <- pbeor$data[[2]][c("x","y")]
          
          # Center of ellipse
          ctreor <- MASS::cov.trob(eleor)$center
          
          # Distance to center from each point on ellipse
          dist2centereor <- sqrt(rowSums((t(t(eleor)-ctreor))^2))
          
          # Area of ellipse from semi-major and semi-minor axes which are largest and smallest of dist to center
          habitatspecificityeor <- pi*min(dist2centereor)*max(dist2centereor)
          hsout_eor <- data.frame(ErrorType = "EOR", SpeciesNum = x, 
                              kfold = 1, habspec = habitatspecificityeor)
  }
    
    
  ellipseouterror <- do.call(rbind, lapply(c("no","yes"), function(errortype){
    # get and average across the replicates
    ellipseout <- do.call(rbind,lapply(1:10, function(rep){
      if(nrow(allsphabsp[allsphabsp$SpeciesNum==x&allsphabsp$kfold==rep&allsphabsp$ErrorType==errortype,]>2)){
        p <- ggplot(allsphabsp[allsphabsp$SpeciesNum==x&allsphabsp$kfold==rep&allsphabsp$ErrorType==errortype,],
                    aes(Comp.1,Comp.2))+
          geom_point()+
          stat_ellipse(segments=201) #default is to draw 51 line segments to make the ellipse
        # get ellipse coordinates
        pb <- ggplot_build(p)
        table(pb$data[[2]]$group)
        
        el <- pb$data[[2]][c("x","y")]
        
        # Center of ellipse
        ctr <- MASS::cov.trob(el)$center
        
        # Distance to center from each point on ellipse
        dist2center <- sqrt(rowSums((t(t(el)-ctr))^2))
        
        # Area of ellipse from semi-major and semi-minor axes which are largest and smallest of dist to center
        habitatspecificity <- pi*min(dist2center)*max(dist2center)
        hsout <- data.frame(ErrorType = errortype, SpeciesNum = x, kfold = rep, habspec = habitatspecificity)
        hsout
      }
        }))
      ellipseout
      }))
    allout <- rbind(ellipseouterror,hsout_eor)
    allout
}))

  

```

Plot habitat specificity
```{r}
habsp <- do.call(rbind,lapply(atleast12, function(x){
  out <- do.call(rbind,lapply(1:10, function(rep){
    load(paste("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/RMaxEnt_herb_herbpluserror/maxentHerbnoerrorSp", x,"kfold",rep, ".Rda",sep=""))
    outinner <- data.frame(SpeciesNum = x, kfold = rep, xm@presence)
    outinner
    }))
  out
  }))
# Would need to add a column for species and another for replicate and then ignore those columns for the princomp of all species and replicates at a time
prPCA <- princomp(habsp[,-c(1:2)])
summary(prPCA)
loadings(prPCA)

allsphabsp <- data.frame(habsp[,1:2],prPCA$scores)

# for one run of one species
ggplot(allsphabsp, aes(Comp.1,Comp.2, colour=as.factor(SpeciesNum)))+
  geom_point()+
  stat_ellipse(segments=201)+ #default is to draw 51 line segments to make the ellipse
  theme_bw()+
  xlab(expression(atop("PC1 (97% variance explained)", "Elevation = 0.974; Bio12 = 0.227")))+
  ylab(expression(atop("PC2 (2% variance explained)","Aspect = -0.999")))
  

summaryhabspec <- summarySE(habspecificity, measurevar = "habspec", groupvars = c("SpeciesNum"))


sumhabspec <- merge(summaryhabspec, tab1, by.x = "SpeciesNum", by.y = "SpNum1")
habspnoerror <- merge(habspsame, tab1, by.x = "SpeciesNum", by.y = "SpNum1")
habspWITHerror <- merge(habspwitherror, tab1, by.x = "SpeciesNum", by.y = "SpNum1")

plot(summaryhabspec$SpeciesNum,summaryhabspec$habspec, pch=16, cex=2)
points(habspsame$SpeciesNum, habspsame$habspec, col="red")

boxplot(habspnoerror$habspec ~ round(habspnoerror$Area/(1000^2),0), 
        col=rgb(0,0,0,.15), las=2, outline=F, 
        at = (1:length(unique(habspnoerror$Area)))-0.5) #c(,rgb(1,1,1,.15)),


ggplot(habspnoerror, aes(as.factor(round(Area/(1000^2),0)), habspec, colour=as.factor(SpeciesNum)))+
  geom_jitter()+
  geom_boxplot(position=position_dodge(width=-0.5))+
  theme_bw()+
  theme(legend.position = "none")+
  xlab(expression("Area km"^2))+
  geom_boxplot()

identical(names(habspnoerror),names(habspWITHerror))
habspnoerror$ErrorType <- "no"
habspWITHerror$ErrorType <- "yes"
toplothabsp <- rbind(habspnoerror, habspWITHerror)

habspall <- merge(habspecificity, tab1, by.x = "SpeciesNum", by.y = "SpNum1")

jpeg("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Images/Figure2.jpg",
     width=150, height=100,units='mm', res=300)

ggplot(habspall, aes(Area/(1000^2), habspec/1000, shape=ErrorType, colour=ErrorType))+
  geom_point(position = position_dodge(width=1))+
  stat_smooth(method="lm", se=FALSE)+
  theme_bw()+
  scale_colour_grey()+
  theme(legend.position = "none")+
  xlab(expression("Area km"^2))+
  ylab("Habitat specificity (size of 95% CI)")

dev.off()

# ggplot(toplothabsp, aes( habspec, DistAvg.Dist,shape=ErrorType, colour=ErrorType))+
#   geom_point()+
#   stat_smooth(method="lm")+
#   theme_bw()
#   # theme(legend.position = "none")+
#   # xlab(expression("Area km"^2))

```


Distance from all specimens to all EORs regardless of year      
See distXsp_noEORgrid <- in "Herb to EORs_v4.Rmd"
```{r}
load("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/distXsp_noEORgrid.Rda")

allspecies <- mapply(cbind, distXsp_noEORgrid, "SpeciesNum" = 1:60, SIMPLIFY = FALSE)
allspecies <- do.call(rbind, allspecies[notnull])
gc()

```






