---
title: "R Notebook"
output: html_notebook
---

I think the outputs that don't have _EOR or _herb are the ones that used Herb to train and EOR to test. can see what the AUC patterns were for those. 
Making new herb train and EOR test ones with VisTrails and same size EOR numbers
```{r}
rm(list=ls())

library(ggplot2)
library(rgeos)
library(raster)
library(foreach)
library(parallel)
library(doParallel)

library(dismo)
library(rgeos)



# Instead, load:
load("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/distXsp_v3.Rda")

notnull <- which(!unlist(lapply(mapply('[[', distXsp_v3, 1), is.null)))
morethan12 <- c()
for(l in notnull){
  morethan12[match(l,notnull)] <- if(nrow(distXsp_v3[[l]][[1]]@coords)>12){ 
    l } else {
      NA
    }
}
morethan12 <- morethan12[!is.na(morethan12)]  

template <- raster("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/VisTrails/PARC_COplus_slope25/COplus_slope50.tif")


# coplus50int.tif
coElev_res <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/ElevationResampled.tif")
# aspect50int.tif
coAspect_res <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/AspectResampled.tif")

# COplus_ruggedInt50.tif
coRugged_res <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/RuggedResampled.tif")

bio1_res <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/Bio1Resampled.tif")
bio12_res <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/Bio12Resampled.tif")


rasterstack <- stack(coElev_res,coAspect_res,coRugged_res,bio1_res,bio12_res)

# pathstart <- "Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/RMaxEnt/Arabis crandallii_EOR/"
# 
# resultpath <- list.files(path = pathstart, pattern = "maxentResults.csv") # , full.names = TRUE)[[1]]

# I don't think I want these, went through one by one but had issues. 
    # maxouts <- list.dirs(path = "Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/RMaxEnt",full.names = TRUE, recursive = TRUE)
    # 
    # maxoutsubset <- maxouts[!grepl("plots", maxouts)]
    # maxoutsubset <- maxoutsubset[!grepl("_EOR", maxoutsubset)]
    # maxoutsubset <- maxoutsubset[!grepl("_herb", maxoutsubset)]
    # 
    # sapply(strsplit(maxoutsubset,"/"), function(x){
    #   x[7]
    # })

# Where is the information about EOR size and distances to match to the names? 

      # AUCresults <- do.call(rbind,lapply(maxoutsubset, function(x){
      #   read.csv(paste(x,"/maxentResults.csv", sep=""))
      # }))
      # 
      # names(AUCresults)[grep("AUC",names(AUCresults))]
      # 
      # jpeg("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Images/AUCtraintest1.jpg",
      #      width=200, height=175,units='mm', res=300)
      # hist(AUCresults$Training.AUC, breaks = 10,
      #      main="",
      #      xlab="AUC")
      # dev.off()
```

Get all the files that I'm making with VisTrails, then extract the information from the sets I want. Named similar runs the same way. 
```{r}
maxouts2 <- list.dirs(path = "Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/VisTrails/EORvHERB_2019/EORvHERB_2019",full.names = TRUE, recursive = TRUE)


#EORsamesize matches _Sp with the same number
# These all have k-fold of 5 so keeping out 20% each time to test but for some, too few points, so should compare to subsampling with replacement/bootstrapping 
EORsame <- maxouts2[grepl("_EORsamesize", maxouts2)]
speciesEOR <- unique(unlist(lapply(regmatches(EORsame, gregexpr("[[:digit:]]+", EORsame)),"[[", 3)))

# Must be an easier way but only want the first folder, don't want to dig in any deeper
EORsamesub <- EORsame[!grepl("Merged", EORsame)]
EORsamesub <- EORsamesub[!grepl("plots", EORsamesub)]
EORsamesub <- EORsamesub[!grepl("cvSplit", EORsamesub)]
EORsamesub <- EORsamesub[!grepl("Expanded", EORsamesub)]
EORsamesub <- EORsamesub[!grepl("responseCurves", EORsamesub)]

# The 9th layer is the folder for that Maxent output
unique(sapply(strsplit(EORsamesub,"/"), function(x){
  x[9]
}))

# Which species number is each? 
unique(sapply(strsplit(EORsame,"Sp"), function(x){
  as.numeric(gsub("([0-9]+).*$", "\\1", x[2]))
}))

# Split at the Sp, take that second string that starts with the species number and extract the first number
sapply(strsplit(EORsamesub, "Sp"), function(x){
  as.numeric(gsub("([0-9]+).*$", "\\1", x[2]))
})

# Where is the information about EOR size and distances to match to the names? 
AUCresults2 <- do.call(rbind,lapply(EORsamesub, function(x){
  resultfile <- read.csv(paste(x,"/maxentResults.csv", sep=""))
  resultfile$Species <- as.numeric(resultfile$Species)
  y <- strsplit(x,"Sp")
  resultfile$Species <- as.numeric(gsub("([0-9]+).*$", "\\1", y[[1]][2]))
  resultfile
}))

hist(AUCresults2$Training.AUC)

jpeg("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Images/AUC_EOR.jpg",
     width=200, height=175,units='mm', res=300)
hist(AUCresults2$Training.AUC, breaks = 10,
     main="",
     xlab="AUC")
dev.off()
```

Now get these rasters and see where the EOR differs from the Herb
```{r}
# Use the resampled one instead
# coElev <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/coplus50int.tif")
# coAspect <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/aspect50int.tif")
# coSlope <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/COplus_slope50.tif")
# coRugged <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/COplus_ruggedInt50.tif")

# x<-EORsamesub[[1]]
# raster1 <- raster(paste(x,"/Maxent_prob_map.tif", sep=""))

# resample(coAspect, raster1, method="bilinear", filename = "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/AspectResampled.tif")
# 
# resample(coSlope, raster1, method="bilinear", filename = "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/SlopeResampled.tif")
# 
# resample(coRugged, raster1, method="bilinear", filename = "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/RuggedResampled.tif")

# Get raster1 to match coElev, no go coElev to raster1
# resample(coElev, raster1, method="bilinear", filename = "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/ElevationResampled.tif")

# resample(coElev, raster1, method="bilinear", filename = "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/ElevationResampled.tif")

coElev_res <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/ElevationResampled.tif")

coAspect_res <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/AspectResampled.tif")

coSlope_res <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/SlopeResampled.tif")

coRugged_res <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/RuggedResampled.tif")

# Function to calculate the linear model of one raster to another
# 1 would be the intercept I guess, while y are the values for y (from the resulting probability map)
# Don't quite get it
          # X <- cbind(1, y)
          # invXtX <- solve(t(X) %*% X) %*% t(X)
          # quickfun <- function(i) (invXtX %*% i)
          # # raster::calc calculates values for a new raster from another
          # m <- calc(s, quickfun) 
          # names(m) <- c('intercept', 'slope')
# r <- calc(s, fun)
# y <- values(raster1)
# X1 <- values(coElev)

# Could also ignore the not likely but that cuts off lots... maybe just the really near 0?
raster1sm <- raster1
raster1sm[raster1sm<0.05] <- NA

# But this is basically just what maxent does, right? but with a prior that is flat or 'maximizes entropy'?
s <- stack(raster1, coElev_res, coAspect_res, coSlope_res, coRugged_res)
# Make it a little smaller
v <- sampleRegular(s, 100000)  
v <- data.frame(na.omit(v))
variables <- names(v)
m <- lm(v[,1] ~ v[,2] + v[,3] + v[,4] + v[,5])
summary(m)
plot(v[,5],v[,1],
     xlab = names(v)[5],
     ylab = names(v)[1])
# I think I want to compare the EOR to climate and Herb and see if the slopes are different
# What do I really want to see? Maybe AIC if the model is different for Herb vs. EOR? That would be similar to loadings fo ra PCA? 
library(ks)
library(MASS)
?kde2d
plot(v[,1],v[,2])
contour(kde2d(v[,1], v[,2], lims = c(min(v[,1]), max(v[,1]),
                                     min(v[,2]), max(v[,2]))), 
        levels= c(min(v[,1]), max(v[,1]),
                                     min(v[,2]), max(v[,2])))

# X <- cbind(1, values(s))
# X <- cbind(1, as.matrix(v))
# invXtX <- solve(t(X) %*% X) %*% t(X)
# quickfun <- function(i) (invXtX %*% i)
# m <- quickfun(as.matrix(v))
# summary(m)

EORrasters <- lapply(EORsamesub, function(x){
  raster1 <- raster(paste(x,"/Maxent_prob_map.tif", sep=""))
  s <- stack(raster1, coElev_res, coAspect_res, coSlope_res, coRugged_res)
  v <- sampleRegular(s, 100000)  
  v <- data.frame(na.omit(v))
  variables <- names(v)
  m <- lm(v[,1] ~ v[,2])
  summary(m)
  plot(v[,2],v[,1])
  abline(m)
})

```



Now the herbarium specimens!
```{r}
# Compare to the Herb results
Herbsame <- maxouts2[grepl("Maxent_Sp", maxouts2)]
speciesHerb <- unique(unlist(lapply(regmatches(EORsame, gregexpr("[[:digit:]]+", EORsame)),"[[", 3)))
match(speciesHerb,speciesEOR)

# Must be an easier way but only want the first folder, don't want to dig in any deeper
Herbsamesub <- Herbsame[!grepl("Merged", Herbsame)]
Herbsamesub <- Herbsamesub[!grepl("plots", Herbsamesub)]
Herbsamesub <- Herbsamesub[!grepl("cvSplit", Herbsamesub)]
Herbsamesub <- Herbsamesub[!grepl("Expanded", Herbsamesub)]
Herbsamesub <- Herbsamesub[!grepl("responseCurves", Herbsamesub)]

# read.csv(paste(Herbsamesub[[5]],"/maxentResults.csv", sep=""))

AUClist <- lapply(Herbsamesub, function(x){
  resultfile <- read.csv(paste(x,"/maxentResults.csv", sep=""))
  resultfile$Species <- as.numeric(resultfile$Species)
  y <- strsplit(x,"Sp")
  resultfile$Species <- as.numeric(gsub("([0-9]+).*$", "\\1", y[[1]][2]))
  resultfile
})


# Go back and see if can run those without threshold and such so they're the same. 
# don't want 5, 7, 10
AUCresults3 <- do.call(rbind,lapply(c(1:4,6,8:9,11:14), function(x){
  AUClist[[x]]
}))


# AUCresults3 <- do.call(rbind,lapply(Herbsamesub, function(x){
#   read.csv(paste(x,"/maxentResults.csv", sep=""))
# }))

hist(AUCresults3$Training.AUC)

jpeg("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Images/AUC_Herb.jpg",
     width=200, height=175,units='mm', res=300)
hist(AUCresults3$Training.AUC, breaks = 10,
     main="",
     xlab="AUC")
dev.off()
# TestEOR_ only want one per species, manually go in and delete extras that had mishaps

```

If i can get more, add 'em up
```{r}



jpeg("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Images/AUC_EORandHerb.jpg",
     width=400, height=175,units='mm', res=300)

layout(matrix(c(1,2), 1,2, byrow=FALSE),
             width=c(6,6), height=c(1))
hist(AUCresults2$Training.AUC, breaks = 10,
     main="EOR",
     xlab="AUC")

hist(AUCresults3$Training.AUC, breaks = 10,
     main="Herbarium",
     xlab="AUC")

dev.off()


hist(AUCresults2$Training.AUC, breaks=15, col=rgb(1,0,0,0.5),
     main="Red is EOR; Blue is Herb")  
hist(AUCresults3$Training.AUC, add=TRUE,breaks=15, col=rgb(0,0,1,0.5))

```

# Rounds of VisTrails with elevation: coplus50int; aspect: aspect50int; ruggedness: COplus_ruggedInt50; annual temp Bio1 and annual precip Bio12 

Name the EOR ones as EORSp1... and compare   
<https://gis.stackexchange.com/questions/29118/how-to-find-the-average-raster-value-of-an-area-defined-by-a-shapefile-using-r> 
```{r}
dirs.vistrails <- list.dirs(path = "Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/VisTrails/EORvHERB_2019/EORvHERB_2019",full.names = TRUE, recursive = TRUE) 

dirs.vistrails[grep("_Sp",dirs.vistrails)]


```


# Maxent 
# VisTrails stopped working, error in making the probability maps so trying here
```{r}
# load("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/pca_data.Rda")


# linkspeciesnumber <- do.call(rbind,lapply(1:length(pca_data), function(x){
#   data.frame(SpNum = x, Species = unique(pca_data[[x]]$GNAME[pca_data[[x]]$pa==3]),
#              Dist = pca_data[[x]]$dist[pca_data[[x]]$pa==2], # pa==2 are herb
#              SpeciesArea = unique(pca_data[[x]]$area[pca_data[[x]]$pa==1]))
# }))

x <- mapply('[[', distXsp_v3, 1)
Herbcoords <- x[!unlist(lapply(x, is.null))]
notnull <- which(!unlist(lapply(x, is.null)))



linkspeciesnumber <- do.call(rbind,lapply(notnull, function(x){
  data.frame(Species = distXsp_v3[[x]][[1]]@data$scientificName[1], Dist = distXsp_v3[[x]][[2]],
             Area = sum(distXsp_v3[[x]][[3]]), SpNum = x)
}))

ggplot(linkspeciesnumber, aes(Species, log(Dist)))+
  geom_boxplot(outlier.shape=NA) + #avoid plotting outliers twice
  geom_jitter(position=position_jitter(width=.1, height=0))

ggplot(linkspeciesnumber, aes(Area, log(Dist)))+
  geom_point()+
  stat_smooth(method="lm")+
  ylab("Distance (log(meters))")+
  theme_bw()


# the EOR AUC results from VisTrails, samesize runs from ForVisTrails.Rmd where I either thinned to match number of Herbarium records or where I bootstrapped and jittered to get matching points. Used a standalone Dan Warren thinning process to get equally spaced out but covering the most area (of information) for points. 
# 2 is EORs, .x; 3 is Herbs, .y
EORmaxentdata <- merge(linkspeciesnumber, AUCresults2[,c(1:2,5:7)], by.x = "SpNum", by.y = "Species")
EORxHarbymaxentdata <- merge(EORmaxentdata, AUCresults3[,c(1:2,5:7)], by.x = "SpNum", by.y = "Species")

avgDist <- aggregate(EORmaxentdata$Dist, list(EORmaxentdata$SpNum), mean) 
sdDist <- aggregate(EORmaxentdata$Dist, list(EORmaxentdata$SpNum), sd) 
EORsummarymaxent <- merge(unique(linkspeciesnumber), AUCresults2, by.x = "SpNum", by.y = "Species")

summary(lm(Training.AUC ~ Dist,data = EORmaxentdata))

ggplot(EORmaxentdata, aes(Area, Training.AUC))+
  geom_point()+
  stat_smooth(method="lm")+
  theme_bw()

ggplot(EORmaxentdata, aes(log(Dist), Training.AUC))+
  geom_point()+
  stat_smooth(method="lm")+
  theme_bw()

ggplot(EORmaxentdata, aes(as.factor(Training.AUC), log(Dist)))+
  geom_boxplot(outlier.shape=NA) + #avoid plotting outliers twice
  geom_jitter(position=position_jitter(width=.1, height=0))

ggplot(EORmaxentdata, aes(as.factor(Training.AUC), Area))+
  geom_boxplot(outlier.shape=NA) + #avoid plotting outliers twice
  geom_jitter(position=position_jitter(width=.1, height=0))

EORHERBarea <- unique(EORxHarbymaxentdata[,-3])

# Red is EOR, Blue is Herb

plot(log(EORHERBarea$Area), EORHERBarea$Training.AUC.x, col="red",pch=16,
     ylim=c(0.55,.95))
points(log(EORHERBarea$Area), EORHERBarea$Training.AUC.y, col="blue", pch=4)


jpeg("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Images/AUCxSampleSizeandArea.jpg",
     width=250, height=175,units='mm', res=300)
plot(EORHERBarea$X.Training.samples.x, EORHERBarea$Training.AUC.x, col="red",pch=16,
     ylim=c(0.55,.95), cex=round(log(EORHERBarea$Area)-11,0),
     xlab="Training sample size",
     ylab="AUC")
points(EORHERBarea$X.Training.samples.x, EORHERBarea$Training.AUC.y, col="blue", pch=16,
       cex = round(log(EORHERBarea$Area)-11,0))
dev.off()

```


Want to add vegetation types, layer too. should work now that I've got all the other layers a better size

Make maxent commands
for output prediction map, use ENMeval   
Using KDE at 95% in MAxent VisTrails which mostly makes it so that it appears there's less effort in areas with fewer points but might not be the case, just might be smaller popualtions so maybe just circles with a certain radius would be better. Maybe add some points along roads for circles. Could put points along roads and then first pick points along roads within a certain distance from presence points and then 'circles' around all those points ?  
follow https://rspatial.org/sdm/SDM.pdf while VisTrails isnt' working
```{r}
# tested in Herb to EORs_v4

# library(adehabitatHR)
# library(ks)

#Ah ha! so many errors for kernel size, so switch to circles
# bgHerbKDE <- lapply(notnull, function(x){
#   xy <- data.frame(distXsp_v3[[x]][[1]]@coords)
#   names(xy) <- c("X","Y")
#   xy <- SpatialPoints(xy)
#   ud <- kernelUD(xy, extent=0.5, grid=150)
#   ver <- getverticeshr(ud, 80)
#   # just put random points for background points witin the kernelUD 
#   spsample(ver, 300, "stratified")
#   })

x <- mapply('[[', distXsp_v3, 1)
Herbcoords <- x[!unlist(lapply(x, is.null))]
notnull <- which(!unlist(lapply(x, is.null)))

xEOR <- mapply('[[', distXsp_v3, 4)
EORcoords <- xEOR[!unlist(lapply(xEOR, is.null))]
notnullEOR <- which(!unlist(lapply(xEOR, is.null)))

setdiff(notnull,notnullEOR) # 8 is naughty, who is 8?
distXsp_v3[[8]][[1]]@data$scientificName # Boechera glareosa, where did you go?! Boechera gunnisoniana? 

class(distXsp_v3[[1]][[1]])
#Background points, in a 5km circle around presence points, 300 stratified across the circles. Not quite the KDE but if I combined other collections from same trip I think that would be a correct estimator, circles represent the posibility that folks hiked 3 miles on their collection trip, seems likely enough. 

bgHerb <- lapply(notnull, function(x){
  circles <- circles(distXsp_v3[[x]][[1]], d = 5000) #Should be 5km around
  polygns <- polygons(circles)
  bg <- spsample(polygns, 300, "stratified") # one single random location in each 'cell' 
  # do I want to get just one per cell? I'm gonna skip cellFromXY(mask, bg) # where mask should be my one used to set the size and clip the others in vistrails
  bg
  })
save(bgHerb, file= "Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/bgHerb.Rda")

# I know which they are, in order they are notnullEOR
bgEOR <- lapply(notnullEOR, function(x){
  df <- distXsp_v3[[x]][[4]]
  coordinates(df) <- ~x+y
  proj4string(df) <- CRS("+proj=utm +zone=13 ellps=NAD83 +ellps=WGS84")
  circles <- circles(df, d = 5000) #Should be 5km around
  polygns <- polygons(circles)
  bg <- spsample(polygns, 300, "stratified") # one single random location in each 'cell' 
  # do I want to get just one per cell? I'm gonna skip cellFromXY(mask, bg) # where mask should be my one used to set the size and clip the others in vistrails
  bg
  })
save(bgEOR, file= "Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/bgEOR.Rda")


```

Extract values from rasters
```{r}

template <- raster("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/VisTrails/PARC_COplus_slope25/COplus_slope50.tif")


# coplus50int.tif
coElev_res <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/ElevationResampled.tif")
# aspect50int.tif
coAspect_res <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/AspectResampled.tif")
# coSlope_res <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/SlopeResampled.tif")
# COplus_ruggedInt50.tif
coRugged_res <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/RuggedResampled.tif")
# bio1 <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/BioClim/20190122_WorldClimv2/wc2.0_bio_30s_01.tif")
# bio12 <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/BioClim/20190122_WorldClimv2/wc2.0_bio_30s_12.tif")

#before resampled to the output Maxent probability map based on template (should be able to do template but it doesn't work...)
resample(bio1, coElev_res, method="bilinear", filename = "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/Bio1Resampled.tif", overwrite=TRUE)
resample(bio12, coElev_res, method="bilinear", filename = "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/Bio12Resampled.tif", overwrite=TRUE)



bio1_res <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/Bio1Resampled.tif")
bio12_res <- raster("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/NationalMapUSDA_DEM_aroundcolorado/Bio12Resampled.tif")


rasterstack <- stack(coElev_res,coAspect_res,coRugged_res,bio1_res,bio12_res)

presvalsHerb <- lapply(notnull, function(x){
  convertxy <- spTransform(distXsp_v3[[x]][[1]], CRS("+proj=longlat +datum=WGS84"))
  dfp <- data.frame(pb = 1, extract(rasterstack, convertxy))
  back <- bgHerb[[match(x,notnull)]]
  proj4string(back) <- CRS("+proj=utm +zone=13 +datum=NAD83 +units=m +no_defs +ellps=GRS80 +towgs84=0,0,0")
  convertback <-spTransform(back, CRS("+proj=longlat +datum=WGS84"))
  dfb <- data.frame(pb = 0, extract(rasterstack, convertback))
  rbind(dfp,dfb)
})

table(presvalsHerb[[4]]$pb)

presvalsEOR <- lapply(notnullEOR, function(x){
  df <- distXsp_v3[[x]][[4]]
  coordinates(df) <- ~x+y
  proj4string(df) <- CRS("+proj=utm +zone=13 ellps=NAD83 +ellps=WGS84")
  convertEOR <- spTransform(df, CRS("+proj=longlat +datum=WGS84"))
  dfp <- data.frame(pb = 1, extract(rasterstack, convertEOR))
  back <- bgEOR[[match(x,notnullEOR)]]
  proj4string(back) <- CRS("+proj=utm +zone=13 +datum=NAD83 +units=m +no_defs +ellps=GRS80 +towgs84=0,0,0")
  convertback <-spTransform(back, CRS("+proj=longlat +datum=WGS84"))
  dfb <- data.frame(pb = 0, extract(rasterstack, convertback))
  rbind(dfp,dfb)
})


#Still not same number of records! how, why, why!?!?
for(i in notnull){
  print(paste(distXsp_v3[[i]][[1]]@data$scientificName[1]))
  print(table(presvalsHerb[[match(i,notnull)]]$pb))
  print(table(presvalsEOR[[match(i,notnullEOR )]]$pb))
}

#Could sample again to get same number of records for presence and absence for each
jar <- paste(system.file(package="dismo"), "/java/maxent.jar", sep='')

x<-1
convertxy <- spTransform(distXsp_v3[[x]][[1]], CRS("+proj=longlat +datum=WGS84"))
convertxy$kfold <- kfold(convertxy, k=4) # to have 75:25%
xm <- maxent(rasterstack,convertxy[convertxy$kfold!=1,])
bgpnts <- bgHerb[[1]]
proj4string(bgpnts) <- CRS("+proj=utm +zone=13 ellps=NAD83 +ellps=WGS84")
bgpnts <- spTransform(bgpnts, CRS("+init=epsg:4326") )
## Loading required namespace: rJava
# plot(xm)
# response(xm)

 # evaluate(pres_test, backg_test, bc, pred_nf)
nrow(convertxy[convertxy$kfold==1,])
nrow(convertxy[convertxy$kfold!=1,])

e <- evaluate(convertxy[convertxy$kfold==1,], bgpnts, xm, rasterstack)
ext <- extent(template)
gc()
predict(rasterstack, xm, ext=ext, progress='', filename = "Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/Predict1.tif", )
par(mfrow=c(1,2))
plot(px, main='Maxent, raw values')
plot(wrld_simpl, add=TRUE, border='dark grey')


library(foreach)
library(parallel)
library(doParallel)
#register parallel computing backend
ncores=10
cl = parallel::makeCluster(ncores)
doParallel::registerDoParallel(cl,ncores)

#compute indices for data splitting
rows = 1:nrow(rasterstack)
split = sort(rows%%ncores)+1

outname = "Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/Predict1"

#perform the prediction on subsets of the predictor dataset
prediction = foreach(i=unique(split), .combine=c)%dopar%{
  rows_sub = rows[split==i]
  sub = raster::crop(rasterstack,raster::extent(rasterstack, min(rows_sub), max(rows_sub), 
                                               1, ncol(rasterstack)))
  raster::predict(sub, xm, filename=paste(outname, i, sep="_"))
}

gc()
ext
template <- raster(ext)
projection(template) <- '+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs'
writeRaster(template, file="Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/Predictall", format="GTiff")

# get all these files we made
pathstart <- "Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/"

resultpath <- list.files(path = pathstart, pattern = "Predict1" , full.names = TRUE) #[[1]]

x <- lapply(resultpath, function(x){
  raster(x)
})
# x <- list(r1, r2)
x$filename <- "Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/PredictAll1.tif"
x$overwrite <- TRUE
m <- do.call(merge, x)

plot(m)

```
## Herbarium specimens
```{r}


#Could sample again to get same number of records for presence and absence for each
jar <- paste(system.file(package="dismo"), "/java/maxent.jar", sep='')

ncores=10
cl = parallel::makeCluster(ncores)
ext <- extent(bio1_res)

    
  
for(x in morethan12[-1]){
  set.seed(1234)
    convertxy <- spTransform(distXsp_v3[[x]][[1]], CRS("+proj=longlat +datum=WGS84"))
    convertxy$kfold <- kfold(convertxy, k=4) # to have 75:25%
    xm <- maxent(rasterstack,convertxy[convertxy$kfold!=1,])
    save(xm, file= paste("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/maxentSp",x,".Rda", sep=""))
   
    gc()
    
    #register parallel computing backend
    cl = parallel::makeCluster(ncores)
    doParallel::registerDoParallel(cl,ncores)
    #compute indices for data splitting
    rows = 1:nrow(rasterstack)
    split = sort(rows%%ncores)+1
    outname = paste("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/Predict", x, sep="")
    #perform the prediction on subsets of the predictor dataset
    prediction <- foreach(i=unique(split), .combine=c)%dopar%{
      rows_sub = rows[split==i]
      sub = raster::crop(rasterstack,raster::extent(rasterstack, min(rows_sub), max(rows_sub), 
                                                    1, ncol(rasterstack)))
      raster::predict(sub, xm, filename=paste(outname, i, sep="_"), overwrite=TRUE)
    }
    rm(xm)
    gc()
    stopCluster(cl)
    # template <- raster(ext)
    # projection(template) <- '+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs'
    # writeRaster(template, file=paste("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/Predictall",x,sep=""), format="GTiff")
    # # get all these files we made
    # pathstart <- "Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/"
    # resultpath <- list.files(path = pathstart, pattern = paste("Predict",x,sep="") , full.names = TRUE) #[[1]]
    # xout <- lapply(resultpath, function(y){
    #   raster(y)
    #   })
    # 
    # xout$filename <- paste("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/Predictall",x,".tif", sep="")
    # xout$overwrite <- TRUE
    # m <- do.call(merge, xout)
  }





```
EOR 
```{r}

ncores=10
cl = parallel::makeCluster(ncores)
ext <- extent(bio1_res)


for(x in morethan12[morethan12!=8]){
  set.seed(1234)
  sampleto <- nrow(distXsp_v3[[x]][[1]]@coords)
  df <- distXsp_v3[[x]][[4]]
  df <- df[sample(nrow(df), sampleto), ]
  coordinates(df) <- ~x+y
  proj4string(df) <- CRS("+proj=utm +zone=13 ellps=NAD83 +ellps=WGS84")
  convertEOR <- spTransform(df, CRS("+proj=longlat +datum=WGS84"))
  convertEOR$kfold <- kfold(convertEOR, k=4) # to have 75:25%
  
  xm <- maxent(rasterstack,convertxy[convertxy$kfold!=1,])
  save(xm, file= paste("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/maxentEORSp",x,".Rda", sep=""))
  gc()
  
  #register parallel computing backend
    cl = parallel::makeCluster(ncores)
    doParallel::registerDoParallel(cl,ncores)
    #compute indices for data splitting
    rows = 1:nrow(rasterstack)
    split = sort(rows%%ncores)+1
    outname = paste("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/PredictEOR", x, sep="")
    #perform the prediction on subsets of the predictor dataset
    prediction <- foreach(i=unique(split), .combine=c)%dopar%{
      rows_sub = rows[split==i]
      sub = raster::crop(rasterstack,raster::extent(rasterstack, min(rows_sub), max(rows_sub), 
                                                    1, ncol(rasterstack)))
      raster::predict(sub, xm, filename=paste(outname, i, sep="_"), overwrite=TRUE)
    }
    rm(xm)
    gc()
    stopCluster(cl)
   
  }





```


# Now grab all the maxent models and get the predict and evalute data from them   
Crap, set seed not in there so the background group might be different than it was first, crap crap
I think each maxent model took about 45 minutes 
```{r}

notnull
morethan12

lapply(morethan12, function(x){
    set.seed(1234)
    convertxy <- spTransform(distXsp_v3[[x]][[1]], CRS("+proj=longlat +datum=WGS84"))
    convertxy$kfold <- kfold(convertxy, k=4) # to have 75:25%
    # xm <- maxent(rasterstack,convertxy[convertxy$kfold!=1,])
    load(paste("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/maxentSp",x,".Rda", sep=""))
    gc()
    bgpnts <- bgHerb[[match(x,notnull)]]
    proj4string(bgpnts) <- CRS("+proj=utm +zone=13 ellps=NAD83 +ellps=WGS84")
    bgpnts <- spTransform(bgpnts, CRS("+init=epsg:4326") )
    e <- evaluate(convertxy[convertxy$kfold==1,], bgpnts, xm, rasterstack)
    save(e, file= paste("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/evaluateSp",x,".Rda", sep=""))
    e
    })

    
```

```{r}

evaluations <- do.call(rbind,lapply(morethan12, function(x) {
  load(paste("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/evaluateSp",x,".Rda", sep=""))
  out <- data.frame(SpNum = x, AUC = e@auc)
  out
})
)

hist(evaluations$AUC)
```

# make all the rasters from the 10 strips
```{r}

ext <- extent(template)

for(x in morethan12){
    template <- raster(ext)
    projection(template) <- '+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs'
    writeRaster(template, file=paste("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/PredictallSp",x,sep=""), format="GTiff", overwrite=TRUE)
    # get all these files we made
    pathstart <- "Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/"
    resultpath <- list.files(path = pathstart, pattern = paste("Predict",x,sep="") , full.names = TRUE) #[[1]]
    xout <- lapply(resultpath, function(y){
      raster(y)
      })
    
    xout$filename <- paste("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/PredictallSp",x,".tif", sep="")
    xout$overwrite <- TRUE
    m <- do.call(merge, xout)
}
```

Examine the maps
```{r}
sp1 <- raster("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/PredictallSp1.tif")

plot(sp1, main=distXsp_v3[[1]][[1]]$scientificName[1])
```




PCA for how different
```{r}



```
