---
title: "Environmental Distance_Herb to EOR"
author: "Michelle DePrenger-Levin"
date: "October 11, 2018"
output: html_document
---

Not doing genetic clustering so probably don't need Thibaut's package...
```{r}
library(devtools)
# install_github("thibautjombart/adegenet")
# install.packages("metricsgraphics")
# library("adegenet")
# library(ade4)
# library(pegas)
# library(strataG)

library(sp)
library(ggplot2)
library(dplyr)
library(htmltools)
library(htmlwidgets)
library(metricsgraphics)
library(RColorBrewer)
library(randomForest)
library(maptools)
library(Taxonstand)
require(AICcmodavg)
require(lme4)
library(lattice)


# load("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/distXsp.Rda")
# load("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/distXsp1.Rda")


load("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/pca_data.Rda")
load("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/pca_data2.Rda")
load("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/stackmini.Rda")
load("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Data/g1g2names.Rda")

l1eor <- readShapePoly("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/L1shp")
#Each polygon might be made of multiple disconnected polygons. Need to separate and label
l1G1G2 <- disaggregate(l1eor[l1eor$GRANK %in% c("G1","G2"),])
l1G1G2$PolyID <- do.call(rbind, lapply(split(l1G1G2,l1G1G2$OBJECTID),
                                       function(x){
              x$POlyID <- paste(x$EO_ID, LETTERS[seq(from=1, to= nrow(x))],sep="")
              x
              }))

# Could take all the synonyms and names of l1G1G2 to make the list of names to pull from SEINet big list
# namesg1g2 <- table(l1G1G2$GNAME)
#length(namesg1g2[namesg1g2 > 0]) #60 species
# g1g2names <- TPL(names(namesg1g2[namesg1g2 > 0]))

length(table(g1g2names$Family))
g1g2names[g1g2names$Family=="",] #Add the Family for Gutierrezia  and Oonopsis

```

Reference stackmini to have same environmental variables in PCA as in the maxent models
```{r}

animatepca_data <- Map(data.frame, pca_data2, SpeciesID=as.list(seq(1,length(pca_data2),1)))

pca_all <- lapply(animatepca_data, function(x){
  x.pca <- princomp(~WinterPrecip+SummerPrecip+WinterMinTemp+WinterMaxTemp+FallMaxTemp, 
                    data = x[,c("WinterPrecip","SummerPrecip","WinterMinTemp","WinterMaxTemp","FallMaxTemp")], 
                    na.action = na.omit, cor=TRUE)
  # x.load <- loadings(x.pca)
  x.pc1 <- predict(x.pca)
  x.pc <- data.frame(x.pc1,x[!is.na(x$WinterMaxTemp),]) #x$pa where 1=EOR presence, 3=EOR background; 2=Herb presence, 4=herb background
  x.PoV <- x.pca$sdev^2/sum(x.pca$sdev^2) #Percent of variance explained by the components
  out <- data.frame(x.pc,PoV1=x.PoV[1],PoV2=x.PoV[2])
  out
})



anim_all <- do.call(rbind, pca_all)

animatePCA <- list()
for(i in 1:46){
  animatePCA[[i]] <- ggplot(anim_all[anim_all$SpeciesID == i,], aes(Comp.1, Comp.2, 
                                 colour=as.factor(pa), 
                                 shape=as.factor(pa),
                                 frame=SpeciesID))+
    ggtitle(anim_all$GNAME[anim_all$SpeciesID == i][1])+
    geom_point()+
                    stat_ellipse()+
                    theme_bw()+
                    xlab(paste("PC1 (",round(anim_all$PoV1[anim_all$SpeciesID == i][1],2)*100,"%)",sep=""))+
                    ylab(paste("PC2 (",round(anim_all$PoV2[anim_all$SpeciesID == i][2],2)*100,"%)",sep=""))+
                    theme(plot.title=element_text(hjust=0))+
                    theme(legend.justification=c(0,1), legend.position=c(0,1),
                          legend.background = element_rect(fill=alpha('white', 0.1)))+
                    #labs(color = "Presence/Background", shape = "Presence/Background")+
    scale_shape_manual(name = "Presence/Background", 
                       labels=c("EOR presence",
                              "Herbarium presence",
                              "EOR background",
                              "Herbarium background"),
                       values=c(16,17,1,2)) +               
    scale_colour_manual(name = "Presence/Background",
                       labels=c("EOR presence",
                              "Herbarium presence",
                              "EOR background",
                              "Herbarium background"),
                       values=c("indianred4","hotpink3","lightslategray","lightsteelblue3")) #("darkolivegreen4","green4","lightgrey","ivory4"))
}

for(i in 1:46){
  print(animatePCA[[i]])
}

```



##CRASHING! skip to cluster analysis for now
Which herbarium speciens fall outside the EOR 95% CI for environmental space?
```{r}
Testone <- 4
# Build the plot first
p <- animatePCA
x <- pca_all[[Testone]] #pca_all is the list that goes into making the plots so should match
any(table(x$pa)<3)

for(Testone in 1:5){
    x <- pca_all[[Testone]] #pca_all is the list that goes into making the plots so should match
  
  # Extract components for the first, the EOR ellipse
    build <- ggplot_build(p[[Testone]])$data
    points <- build[[1]] # [build[[1]]$group==1,] # want to check all points can subset later
    ell <- build[[2]][build[[1]]$group==1,]  # just want one ellipse of the four made
    
    # Find which points are inside the ellipse, and add this to the data
    dat <- data.frame(points[1:2], 
                      in.ell = as.logical(point.in.polygon(points$x, points$y, ell$x, ell$y)))
    
    x[which(dat$in.ell),]
    x[which(!dat$in.ell & x$pa==2),]
    nrow(x[x$pa==2,]);nrow(ell)
    
    hist(x$dist[which(!dat$in.ell & x$pa==2)]/1000, breaks=10)
}
x <- pca_all[[Testone]] #pca_all is the list that goes into making the plots so should match

any(table(x$pa)<3)

# Extract components for the first, the EOR ellipse
  build <- ggplot_build(p[[Testone]])$data
  points <- build[[1]] # [build[[1]]$group==1,] # want to check all points can subset later
  ell <- build[[2]][build[[1]]$group==1,]  # just want one ellipse of the four made
  
  # Find which points are inside the ellipse, and add this to the data
  dat <- data.frame(points[1:2], 
                    in.ell = as.logical(point.in.polygon(points$x, points$y, ell$x, ell$y)))
  
  x[which(dat$in.ell),]
  x[which(!dat$in.ell & x$pa==2),]
  nrow(x[x$pa==2,]);nrow(ell)
  
  hist(x$dist[which(!dat$in.ell & x$pa==2)]/1000, breaks=10)
```

  
#AAHHH crashing, don't do yet!  
```{r}  
outsideEORellipse <- mapply(function(plt, x){
  # Extract components for the first, the EOR ellipse
  if(any(table(x$pa)<3)){
    NA
  } else {
    build <- ggplot_build(plt)$data
    points <- build[[1]] #the EOR presence group
    ell <- build[[2]][build[[1]]$group==1,]    #the EOR presence group
    
    # Find which points are inside the ellipse, and add this to the data
    dat <- data.frame(points[1:2], 
                      in.ell = as.logical(point.in.polygon(points$x, points$y, ell$x, ell$y)))
    #Points that are herbarium specimens outside of the EOR ellipse
    x[which(!dat$in.ell & x$pa==2),]
  }
}, plt = p, x = pca_all)


```

https://rpubs.com/gabrielmartos/ClusterAnalysis  
https://nishanthu.github.io/articles/ClusteringUsingRandomForest.html  

```{r}
x<-pca_all[[1]]

#Do two groups, in and out? ie. presence and background
km.cluster <- kmeans(x[,c("WinterPrecip","SummerPrecip","WinterMinTemp","WinterMaxTemp","FallMaxTemp")], centers = 2, iter.max = 20, nstart = 10)

x$kmeans.cluster <- km.cluster$cluster
table(x$pa, km.cluster$cluster)
table(x$pa, x$kmeans.cluster)
#seems like the two presences eor(1) and herb(2) are mostly in group 1 and a little in 2, background are in group 3

mjs_plot(x, x=Comp.1, y=Comp.2) %>%
  mjs_point(color_accessor=kmeans.cluster) %>%
  mjs_labs(x="principal comp 1", y="principal comp 2")

```


# clustering of k-means - partition points into k groups such that the sum of squares from points to assigned cluster centers is minimized Hartigan and Wong (1979)
Should be able to use Jombart et al. 2010 Discriminant Analysis of Principal Components find.clusters() to see if the herbarium specimens are clustering differently from EORs. For now just run several random starts in kmeans()    
```{r}
kmeansEORHerb <- lapply(pca_all, function(x){
  km.cluster <- kmeans(x[,c("WinterPrecip","SummerPrecip","WinterMinTemp","WinterMaxTemp","FallMaxTemp")],
                       centers = 2, iter.max = 20, nstart = 100)

  x$kmeans.cluster <- km.cluster$cluster
  x
  })

#is the first cluster always the most EORs and Herb presences? Nope, need the one with
for(i in 1:length(kmeansEORHerb)){
  tabl <- table(kmeansEORHerb[[i]]$pa, kmeansEORHerb[[i]]$kmeans.cluster)
 print(tabl)
  if(tabl[1,1]>tabl[1,2]){
    print("1")
  } else {
    print("2")
  }
  
}

#want linear model of in group cluster 1 or 2 as determined by what
#Want all, not just the outliers 
kmall <- do.call(rbind,kmeansEORHerb)

distoutliers <- lapply(kmeansEORHerb, function(x){
  tbl <- table(x$pa,x$kmeans.cluster)
  #Want the outliers so after determining which cluster most EOR presence points went, then pull the other group for the presence group 2 $pa==2 are the Herb points
  if(tbl[1,1]>tbl[1,2]){
    x[x$pa==2 & x$kmeans.cluster==2,]
  } else {
    x[x$pa==2 & x$kmeans.cluster==1,]
  }
})

distout <- do.call(rbind, distoutliers)
is.factor(distout$kmeans.cluster) #needs to be categorical
distout$kmeans.cluster <- factor(distout$kmeans.cluster)


kmall$kmeans.cluster <- factor(kmall$kmeans.cluster)
contrasts(kmall$kmeans.cluster) #so in cluster 2 will get a 1 and in cluster 1 will get a 0; cluster2==1, cluster1==0

table(as.factor(distout$kmeans.cluster))
str(distout$Lastobs)
kmall$Lastobs <- as.Date(kmall$Lastobs, format="%Y-%m-%d")
distout$Lastobs <- as.Date(distout$Lastobs, format="%Y-%m-%d")
kmall$Year <- format(as.Date(kmall$Lastobs, format="%Y-%m-%d"),"%Y")
kmall[!is.na(kmall$Year),]
str(kmall$Year)
kmall$Year <- as.numeric(kmall$Year)

#Model fitting? Should I use some train and testing data? 
model <- glm(factor(kmeans.cluster) ~ Year + area, family=binomial(link="logit"),data= kmall[kmall$pa==2,])
summary(model)

#Make a new column that figures out which is the presence factor for kmeans.cluster, 1 or 2 depending on which has more EOR points, label it 0 for out, 1 for in
InorOut <- do.call(rbind,lapply(kmeansEORHerb, function(x){
  x$InOut <- 0
  tbl <- table(x$pa,x$kmeans.cluster)
  #Want the outliers so after determining which cluster most EOR presence points went, then pull the other group for the presence group 2 $pa==2 are the Herb points
  if(tbl[1,1]>tbl[1,2]){
    x$InOut[x$kmeans.cluster==1] <- 1
    x
  } else {
    x$InOut[x$kmeans.cluster==2] <- 1
    x
  }
}))

InorOut$kmeans.cluster <- factor(InorOut$kmeans.cluster)
InorOut$Lastobs <- as.Date(InorOut$Lastobs, format="%Y-%m-%d")
InorOut$Year <- format(as.Date(InorOut$Lastobs, format="%Y-%m-%d"),"%Y")
InorOut$Year <- as.numeric(InorOut$Year)
```
Now just look at the in or out of a cluster for the herbarium specimens; $pa==2
```{r}
f1 <- glm(InOut ~ as.factor(SpeciesID)*Year + area, family=binomial(link="logit"), data= InorOut[InorOut$pa==2,])
f2 <- glm(InOut ~ Year + area, family=binomial(link="logit"), data= InorOut[InorOut$pa==2,])
f3 <- glm(InOut ~ Year, family=binomial(link="logit"), data= InorOut[InorOut$pa==2,])
f4 <- glm(InOut ~ area, family=binomial(link="logit"), data= InorOut[InorOut$pa==2,]) #x is the area of the species known range
f5 <- glm(InOut ~ 1, family=binomial(link="logit"), data= InorOut[InorOut$pa==2,])
f6 <- glm(InOut ~ Year*area, family=binomial(link="logit"), data= InorOut[InorOut$pa==2,]) #year given the range
f7 <- glm(InOut ~ as.factor(SpeciesID)*Year, family=binomial(link="logit"), data= InorOut[InorOut$pa==2,])


(lmresults <- aictab(list(f1,f2,f3,f4,f5,f6,f7),
       modnames=as.character(unlist(lapply(list(f1,f2,f3,f4,f5,f6,f7),formula)))))

evidence(aictab(cand.set = list(f1,f2,f3,f4,f5,f6,f7),
                modnames = as.character(unlist(lapply(list(f1,f2,f3,f4,f5,f6,f7),formula)))))

sapply(1:length(lmresults$Delta_AICc), function(i){
  exp(-0.5*lmresults$Delta_AICc[i])/sum(exp(-0.5*lmresults$Delta_AICc))
})

(er1 <- exp(0.5*lmresults$Delta_AICc[2]))
(er2 <- exp(0.5*lmresults$Delta_AICc[3]))



```
This is different than before so different distances for what puts the presence in or out of the EOR in cluster
```{r}
f1 <- glm(InOut ~ as.factor(SpeciesID)*Year + area + dist, family=binomial(link="logit"), data= InorOut[InorOut$pa==2,])
f2 <- glm(InOut ~ Year * area * dist, family=binomial(link="logit"), data= InorOut[InorOut$pa==2,])
f3 <- glm(InOut ~ Year*dist, family=binomial(link="logit"), data= InorOut[InorOut$pa==2,])
f4 <- glm(InOut ~ area, family=binomial(link="logit"), data= InorOut[InorOut$pa==2,])
f5 <- glm(InOut ~ 1, family=binomial(link="logit"), data= InorOut[InorOut$pa==2,])
f6 <- glm(InOut ~ Year*area, family=binomial(link="logit"), data= InorOut[InorOut$pa==2,]) #year given the range
f7 <- glm(InOut ~ as.factor(SpeciesID)*Year, family=binomial(link="logit"), data= InorOut[InorOut$pa==2,])
f8 <- glm(InOut ~ dist, family=binomial(link="logit"), data= InorOut[InorOut$pa==2,]) 


(lmresults <- aictab(list(f1,f2,f3,f4,f5,f6,f7,f8),
       modnames=as.character(unlist(lapply(list(f1,f2,f3,f4,f5,f6,f7,f8),formula)))))

evidence(aictab(cand.set = list(f1,f2,f3,f4,f5,f6,f7,f8),
                modnames = as.character(unlist(lapply(list(f1,f2,f3,f4,f5,f6,f7,f8),formula)))))

sapply(1:length(lmresults$Delta_AICc), function(i){
  exp(-0.5*lmresults$Delta_AICc[i])/sum(exp(-0.5*lmresults$Delta_AICc))
})

(er1 <- exp(0.5*lmresults$Delta_AICc[2]))
(er2 <- exp(0.5*lmresults$Delta_AICc[3]))

table(InorOut$Year[InorOut$pa==2])
str(InorOut$Year)
```

```{r}
datapost1980 <- InorOut[InorOut$pa==2 & InorOut$Year>1980,]

f1 <- glm(InOut ~ as.factor(SpeciesID)*Year + area + dist, family=binomial(link="logit"), data= datapost1980)
f2 <- glm(InOut ~ Year * area * dist, family=binomial(link="logit"), data= datapost1980)
f3 <- glm(InOut ~ Year*dist, family=binomial(link="logit"), data= datapost1980)
f4 <- glm(InOut ~ area, family=binomial(link="logit"), data= datapost1980)
f5 <- glm(InOut ~ 1, family=binomial(link="logit"), data= datapost1980)
f6 <- glm(InOut ~ Year*area, family=binomial(link="logit"), datapost1980) #year given the range
f7 <- glm(InOut ~ as.factor(SpeciesID)*Year, family=binomial(link="logit"), data= datapost1980)
f8 <- glm(InOut ~ dist, family=binomial(link="logit"), data= datapost1980) 


(lmresults <- aictab(list(f1,f2,f3,f4,f5,f6,f7,f8),
       modnames=as.character(unlist(lapply(list(f1,f2,f3,f4,f5,f6,f7,f8),formula)))))

evidence(aictab(cand.set = list(f1,f2,f3,f4,f5,f6,f7,f8),
                modnames = as.character(unlist(lapply(list(f1,f2,f3,f4,f5,f6,f7,f8),formula)))))

sapply(1:length(lmresults$Delta_AICc), function(i){
  exp(-0.5*lmresults$Delta_AICc[i])/sum(exp(-0.5*lmresults$Delta_AICc))
})

(er1 <- exp(0.5*lmresults$Delta_AICc[2]))
(er2 <- exp(0.5*lmresults$Delta_AICc[3]))

table(datapost1980$Year)

```

```{r}
jpeg("Q:/Research/Projects/alpine-phenology/Uncertainty in niche modeling/Images/HerbPA.jpg",
     width=450, height=125,units='mm', res=300)

ggplot(InorOut[InorOut$pa==2,], aes(Year,InOut,colour=GNAME))+
  stat_smooth(method="loess", formula=y~x,
              alpha=0.2, size=2, aes(fill=GNAME)) +  # area/(1000^2)
   geom_point()+ #position=position_jitter(height=0.003, width=0)
  # facet_wrap(~SpeciesID)+
  theme_bw()+
  facet_wrap(~GNAME, ncol = 18)+
  ylab("Herbarium presence out (0) or in (1)")+
  xlab("Year")+
  theme(legend.position="none")
  # scale_colour_gradient2(name= expression(paste("Species range (km"^"2",")")),
  #                        low='red',mid='violet',high='blue')

dev.off()

str(as.factor(InorOut$InOut))

```


```{r}
ggplot(InorOut[InorOut$pa==2,], aes(factor(InOut), Year, colour= area/(1000^2)))+
  geom_boxplot() +
  theme_bw()+
  xlab("Herbarium presence out (0) or in (1)")+
  geom_jitter(position=position_jitter(0.25))+
  scale_colour_gradient2(name= expression(paste("Species range (km"^"2",")")),
                         low='red',mid='violet',high='blue')

ggplot(InorOut[InorOut$pa==2,], aes(factor(InOut), area/(1000^2), colour= as.numeric(Year)))+
  geom_boxplot() +
  theme_bw()+
  xlab("Herbarium presence out (0) or in (1)")+
  geom_jitter(position=position_jitter(0.25))+
  scale_colour_gradient2(name= "Year",
                         low='green',mid='violet',high='blue')

ggplot(InorOut[InorOut$pa==2,], aes(factor(InOut), Year, colour= area/(1000^2)))+
  geom_boxplot() +
  facet_wrap(~SpeciesID)+
  theme_bw()+
  xlab("Herbarium presence out (0) or in (1)")+
  geom_jitter(position=position_jitter(0.25))+
  scale_colour_gradient2(name= "Species Range (km2)",
                         low='green',mid='violet',high='blue')

boxplot(Year ~ InOut, data = InorOut[InorOut$pa==2,])
stripchart(Year ~ InOut, vertical = TRUE, data = InorOut[InorOut$pa==2,], 
    method = "jitter", add = TRUE, pch = 20, col = 'blue')

#Oh crap but for each species the presence group and the background group differ

boxplot(log(dist/1000) ~ (InOut), data = InorOut[InorOut$pa==2,],
        xlab = "Cluster", ylab="Geographic distance log(km) to EOR")
stripchart(log(dist/1000) ~ InOut, vertical = TRUE, data = InorOut[InorOut$pa==2,], 
    method = "jitter", add = TRUE, pch = 20, col = 'blue')



boxplot(log(dist/1000) ~ (InOut), data = InorOut[InorOut$pa==2,],
        xlab = "Cluster", ylab="Geographic distance log(km) to EOR")
stripchart(log(dist/1000) ~ InOut, vertical = TRUE, data = InorOut[InorOut$pa==2,], 
    method = "jitter", add = TRUE, pch = 20, col = 'blue')

xyplot(InOut ~ Year|as.factor(SpeciesID), InorOut[InorOut$pa==2,],
       # maybe g:'reference grid' 'point' and r: linear 'regression', a:panel.average for interaction plots
       type = c("p","spline"),
       #index = function(x,y) coef(lm(y~x))[[1]],
       xlab = "Colleciton Year",
       ylab = "Clustered with presence (1) or background (0)")

```

anim_all <-- animatepca_data which is pca_data2 with a species id identifier... Could add that identifier to HAD_all2 to merge them.   
HAD_all2 is the herbarium data from distXsp, the area of EORs from distXsp, and the distance from herbarium points from distXsp to their closest EOR, and the _all2 is all species merged with  
```{r}
rf.fit <- randomForest(x = x[,c("WinterPrecip","SummerPrecip","WinterMinTemp","WinterMaxTemp","FallMaxTemp")], 
                       y = NULL, ntree = 10000, proximity = TRUE, oob.prox = TRUE)
hclust.rf <- hclust(as.dist(1-rf.fit$proximity), method = "ward.D2")
rf.cluster = cutree(hclust.rf, k=4)
x$rf.clusters <- rf.cluster
table(rf.cluster, x$pa) # so the 1:4 along the top are the herb and eor presence and backgrounds

#How far away are the herbarium collections that didn't fall into the 1st or second cluster (the present clusters)?
x$dist[x$rf.clusters==1]

x[x$rf.clusters==1&x$pa==2,] #2 are herbarium that have distances

```

```{r}
mjs_plot(x, x=Comp.1, y=Comp.2) %>%
  mjs_point(color_accessor=rf.clusters) %>%
  mjs_labs(x="principal comp 1", y="principal comp 2")

```


Are the points beyond the 95% CI of the cluster (all the presence points) herbarium specimens that are also far away, so what's the distribution of distance form EOR information   
```{r}
a <- mean(x$Comp.1[x$pa<3])
s <- sd(x$Comp.1[x$pa<3])
n <- length(x$Comp.1)
error <- qnorm(0.975)*s/sqrt(n)


a2 <- mean(x$Comp.2[x$pa<3])
s2 <- sd(x$Comp.2[x$pa<3])
n2 <- length(x$Comp.2)
error2 <- qnorm(0.975)*s2/sqrt(n2)


x[which(x$Comp.1>(a+error)&(x$Comp.2>(a2+error2))&x$pa <3|
        x$Comp.1<(a-error)&x$Comp.2<(a2-error2)&x$pa<3),]

hist(x$Comp.1[])
abline(v=a)

?pointsToEl
```

Find points inside and outside the ellipse
```{r}

# Build the plot first
p <- ggplot(faithful, aes(waiting, eruptions)) +
  geom_point() +
  stat_ellipse()

# Extract components
build <- ggplot_build(p)$data
points <- build[[1]]
ell <- build[[2]]

# Find which points are inside the ellipse, and add this to the data
dat <- data.frame(points[1:2], 
                  in.ell = as.logical(point.in.polygon(points$x, points$y, ell$x, ell$y)))

# Plot the result
ggplot(dat, aes(x, y)) +
  geom_point(aes(col = in.ell)) +
  stat_ellipse()
```

