---
title: "Herbarium Bias CNHP"
author: "Michelle DePrenger-Levin"
date: "December 4, 2017"
output: html_document
---

```{r}
rm(list=ls())

library(ggplot2)
library(rgeos)
library(sp)
library(spdep)
library(rgdal)
library(maptools)
library(raster)
library(Taxonstand)
library(mapproj)
library(maps)

# exact match multisub
multisub <- function(pattern,replacement,x){
  result <- x
  for(i in 1:length(pattern)){
    result <- gsub(paste("^",pattern[i],"$"), paste("^",replacement[i],"$"),result)
  }
  result
}
```

Mo Ewing:    
     1) split all polygons (from CNHP) with ArcGIS 'explosion' tool    
     2) Copy the Feature_id field into the Parcel-num field and annotate each parcel: "a", "b" etc    
     3) Import G1G2 Seinet Vouchers into ArcMap      
     4) Create linking field: "Parcel_num"          
     5) Go through each parcel, linking the Eos by duplicating the Parcel_num.    
     6) determin the closest 'parcel' to each herabarium point using ArcMap 'near' tool Analyst Tools/Proximity/Near   *I assume to the nearest line of the polygon, nearest spot of the polygon, not to the center of the polygon          
     7) Add Map Method informaiton     
            i. "GPS" = indication that GPS was used, just GPS coordinates present    
            ii. "tr" = township/range info only recorded to township and range     
            iii. "trs" = township/Range info recorded to section, or indication that tr conversion tool used       
            iv. "trqs" = Township/Range info recorded to quarter-section       
            v. "Locality" = no verbatim data recorded, only written description recorded    
                     or "Terrain nav" (terrain navigator used)    
                     or "Nat. Geog. TOPO"    
                     or "Mapstedi"     
                     or "digital mape"     
                     or "geoLocate"     
                     or "GoogleEarth"      



# Mo worked on this list to add map method and distance
```{r}
seinet <- read.csv("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/SEINET G1 G2 Vouchers/seinet_colorado_with_coordinates_ALLG1G2.csv")

seinetMo <- TPL(unique(seinet$scientificName))
sum(table(seinetMo$Taxonomic.status)) #93 names
#Mo has 2 names not in TPL, 9 synonyms, 7 unresolved
seinetMo[seinetMo$Taxonomic.status != "Accepted",]
seinet[grep("rotundifo", seinet$scientificName),]

length(table(seinet$scientificName[!is.na(seinet$decimalLatitude)])) #93 species with locaiton information

eors <- read.csv("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/Distance Tables/Vouchers_distance.csv")

length(unique(eors$scientific)) #53
eors.tax <- TPL(unique(eors$scientific))
table(eors.tax$Taxonomic.status)
# 5 syn, 2 unresolved, 1 not in

eors[eors$institutio=="USU",]
eors[eors$institutio=="USUUB",]
eors$institutio[eors$institutio=="USU"] <- "USUUB"

```


#Michelle DePrenger-Levin      
In ArcMap do this: <https://blogs.esri.com/esri/arcgis/2010/09/16/nearbygroup/>
repeat distance analysis in R          
SEINet collections from Colorado - downloaded 12/6/2017 from SEINet        
Only G1G2 EORs (to limit range of species)   
Change names according to TPL as a best effort to match up.    
Only this first chunck if you want to redo naming changes- otherwise SKIP TO NEXT        
```{r}
coloradosps <- read.csv("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/SEINET G1 G2 Vouchers/occurrences.csv")

#colonames <- TPL(unique(coloradosps$scientificName))
#write.csv(colonames, "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/AllcoloSEINet_TPL.csv")
colonames <- read.csv("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/AllcoloSEINet_TPL.csv")

sum(table(colonames$Taxonomic.status))
# 1929 not in The Plant List, 6 misapplied??, lots of synonyms. 
#             Accepted Misapplied    Synonym Unresolved 
#      1929       9149          6       6083        682
colonames[colonames$Taxonomic.status == "Misapplied",]

head(colonames[colonames$Taxonomic.status == "Misapplied",])

fixcolonames <- colonames[colonames$Taxonomic.status == "Synonym",]

coloradosps$genus <- multisub(fixcolonames$Genus, fixcolonames$New.Genus, coloradosps$genus) 
coloradosps$specificEpithet <- multisub(fixcolonames$Species, 
                                   fixcolonames$New.Species, coloradosps$specificEpithet) 

write.csv(coloradosps, "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/coloradosps_synonymsTPL.csv")
```

```{r}
coloradosps <- read.csv("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/coloradosps_synonymsTPL.csv")


l1eor <- readShapePoly("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/L1shp")
```

#Skip this because want split by species, not labeled and match by label. 
```{r}
#Each polygon might be made of multiple disconnected polygons. Need to separate and label
l1G1G2 <- disaggregate(l1eor[l1eor$GRANK %in% c("G1","G2"),])
l1G1G2$PolyID <- do.call(rbind, lapply(split(l1G1G2,l1G1G2$OBJECTID),
                                       function(x){
              x$POlyID <- paste(x$EO_ID, LETTERS[seq(from=1, to= nrow(x))],sep="")
              x
              }))
```

```{r}
#The Plant List
sp43 <- TPL(intersect(speciesG1G2, species))
table(sp43$Taxonomic.status)
syn <- sp43[sp43$Taxonomic.status == "Synonym",]

seinet$genus <- multisub(syn$Genus, syn$New.Genus, seinet$genus) 
seinet$specificEpithet <- multisub(syn$Species, syn$New.Species, seinet$specificEpithet) 

species <- unique(seinet$scientificName)
sort(intersect(species, speciesG1G2))
```

Download all Colorado collections from SEINet (coloradosps) and limit to matching G1 G2 global names by CNHP. Will be missing synonyms
```{r}
#Without synonym corrections
coloG1G2 <- coloradosps[coloradosps$scientificName %in% 
                          unique(l1eor$GNAME[l1eor$GRANK %in% c("G1","G2")]),]

#write.csv(coloG1G2, "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/occurrencesG1G2_nosynonyms.csv")

nrow(coloG1G2) #2125
length(unique(coloG1G2$scientificName)) #59
length(species)
```


```{r}
nrow(seinet[!is.na(seinet$decimalLatitude),]) #1875 points

#Get all points, transform to UTM
pts <- SpatialPoints(seinet[,c("decimalLongitude","decimalLatitude")])
proj4string(pts) <- CRS("+proj=longlat +datum=WGS84")

#Project data into a planar coordinate system (UTM zone 13)
UTMpts <- spTransform(pts, CRS("+proj=utm +zone=13 ellps=WGS84"))

head(pts)
head(UTMpts)

plot(l1eor[l1eor$GRANK %in% c("G1","G2"),])
points(UTMpts)
names(l1eor)

plot(l1eor[l1eor$GNAME == "Astragalus microcymbus",])

#Does not read th .prj file so need to tell it that they're in UTM
proj4string(l1eor) <- "+proj=utm +zone=13 ellps=WGS84"

#might be tooooo big!
#PLSS <- readOGR(dsn="Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/Colorado/PLSS Geo Coordinates_2008")



```


Try to do an apply and split by species, don't really care which EO, just which mapped occurance of a species is closest     
```{r}
#59 species

x<- unique(coloG1G2$scientificName)[1]
nms <- intersect(unique(seinet$scientificName),unique(l1eor$GNAME)) #67 names

mindistxsp <- lapply(nms, function(x){
  pts.sp <- SpatialPoints(seinet[seinet$scientificName == x,
                                 c("decimalLongitude","decimalLatitude")])
  proj4string(pts.sp) <- CRS("+proj=longlat +datum=WGS84")
  UTMpts.sp <- spTransform(pts.sp,CRS("+proj=utm +zone=13 ellps=WGS84"))
  
  l1.sp <- disaggregate(l1eor[l1eor$GNAME == x,])
  
  distout <- apply(gDistance(UTMpts.sp, l1.sp, byid=TRUE),2,min)
  out <- data.frame(distout,seinet[seinet$scientificName == x,])
  list(out,UTMpts.sp,l1.sp)
})


head(mindistxsp[[1]])

mndistsp <- do.call(rbind,lapply(mindistxsp,'[[',1))

#This is in UTM!!! Want it! WANT!
colomap<-  shapefile("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/Colorado/CO_Counties/counties.shp")

for(i in 1:length(mindistxsp)){
  plot(mindistxsp[[i]][[3]],
     xlim=c(142063.3, 740922.7), ylim=c(4094495, 4543961), 
     border="red",cex=5,lwd=3)
  points(mindistxsp[[i]][[2]],col="blue")
  points(gCentroid(colomap, byid=TRUE),col="green",pch=5)
  plot(colomap,add=TRUE)
  }

```

# Have mndistsp is the minimum distances from each point (from SEINet gather all collections from Colorado; match names that are G1G2 with names in SEINet; will be missing any synonyms)   
Unlike eors2 where Mo added a label for method used or accuracy of mapping, need to add to mndistsp     
I don't think we can tell how much infomration there really was without looking at the specimens.   
```{r}
table(mndistsp$year) #1871 to 2017
#Find degree symbol
head(mndistsp[grep(intToUtf8(176),mndistsp$locality),])
mndistsp[grep("[T]{1}[0-9]{2}[NS]{1}",mndistsp$locality),] #Lists TRS but also other UTM or lat long

table(mndistsp$verbatimCoordinates[grep()])


mndistsp[mndistsp$year == 2017 & !is.na(mndistsp$year),] #Eriogonum coloradense

```

```{r}
ggplot(mndistsp[mndistsp$distout<1000000,], aes(scientificName,distout))+
  geom_boxplot()+ 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


```




# How do the maping techniques change by time, herbarium, or collector (amount of informaiton given)?   
```{r}
eors2 <- eors[eors$Year.2>0,]


methodxyear <- data.frame(table(eors2$Map.Method, eors2$Year.2))
methodxyear$Var2 <- as.numeric(as.character(methodxyear$Var2))

eors2$Map.Method <- factor(eors2$Map.Method, levels = c("Locality","GPS","trqs",
                                                        "trs","tr"))

ggplot(methodxyear, aes(Var2, Freq, colour=Var1))+
  geom_point()+
  stat_smooth(se=FALSE)+
  theme_bw()+
  xlab("Year")+
  ylab("Herbarium collections")+
  labs(color='Georeference Method')+
  theme(legend.justification=c(0,1), legend.position=c(0,1),
        legend.background = element_blank()) 
```


```{r}
ggplot(eors[eors$Year.2>0,], aes( Year.2,NEAR_DIST,colour=institutio))+
  geom_point()+
  facet_wrap(~institutio)+
  theme_bw()+
  stat_smooth(method="lm", se=FALSE)

ggplot(eors[eors$Year.2>0,], aes( Year.2,NEAR_DIST))+
  geom_point()+
  theme_bw()+
  stat_smooth(method="lm", se=FALSE)

summary(lm(NEAR_DIST~Year.2, data=eors[eors$Year.2>0,]))
```

How did the type of mapping method change over time?
```{r}

table(eors$Year.2)



```

Which method has the shortest distance?   
```{r}


ggplot(eors2, aes(x=Map.Method))+
  geom_bar(aes(fill=institutio))


ggplot(eors2, aes(Year.2,NEAR_DIST, colour=institutio))+
  geom_point()+
  theme_bw()

ggplot(eors2, aes(x=Year.2, fill=Map.Method))+
  geom_density(alpha=0.5)+
  theme_bw()
#+  facet_wrap(~Map.Method)

ggplot(eors2, aes(x=Year.2, fill=Map.Method))+
  geom_histogram(alpha=0.25)+
  theme_bw()


```

Maybe regional collections are much better?   
FLD = Fort Lewis College    
DES = Dessert Arizona   
MESA = Colorado Mesa University   
MISSA = Miss State Univ   
NMC = New Mexico State University    
SJNM = San Juan College, New Mexico   
UCR = University of California Riverside    
USUUB = Utah STate University Uintah Basin (probably is USU)   
UVSC = Utah Valley University	U.S.A. Utah. Orem.      
```{r}

local <- c("COLO","CS","DBG","FLD","MESA","RM","RMBL")
Adjacent <- c("ASC","ASU","BRY","DES","NMC","RM","SJNM","UNM","USUUB","UVSC")
far <- c("CM","F","MISSA","MO","NY","RENO","UCR")

eors2$InstNear[eors2$institutio %in% local] <- "Local"
eors2$InstNear[eors2$institutio %in% Adjacent] <- "Near"
eors2$InstNear[eors2$institutio %in% far] <- "Far"

eors2$InstNear <- factor(eors2$InstNear, levels = c("Far","Near","Local"))


ggplot(eors, aes(x=institutio))+
  geom_bar(aes(fill=Map.Method))


ggplot(eors2, aes(x=InstNear))+
  geom_bar(aes(fill=Map.Method))

ggplot(eors2, aes(Year.2,NEAR_DIST, colour=InstNear))+
  geom_point()+
  theme_bw()+
  stat_smooth(method="lm", se=FALSE)


ggplot(eors2, aes(NEAR_DIST))+
  geom_density(aes(fill=InstNear))
```

Eastern Illinois Univeristy (Charleston)     
FLD Fort Lewis College     
MARY University of Maryland    
MNA Museum of Northern Arizona     
MWI Western Illinois Univeristy   
NAVA Navajo Nation Department of FIsh and Wildlife     
SDSU San Diego State Univ    
SEINet: error in reporting??    
```{r}
local <- c("COLO","CS","DBG","FLD","MESA","RM","RMBL")
Adjacent <- c("ARIZ","ASC","ASU","BRY","DES","MNA","NAVA","NMC","RM","SJNM","UNM","USUUB","UVSC")
far <- c("CHIC","CM","EIU","F","MARY","MISSA","MO","MWI","NY","RENO","SDSU","UCR","SEINet")

table(mndistsp$institutionCode)
mndistsp[mndistsp$]

eors2$InstNear[eors2$institutio %in% local] <- "Local"
eors2$InstNear[eors2$institutio %in% Adjacent] <- "Near"
eors2$InstNear[eors2$institutio %in% far] <- "Far"

eors2$InstNear <- factor(eors2$InstNear, levels = c("Far","Near","Local"))



```


Would assume that Locality method would be best for local institutions, worse for far
```{r}


ggplot(eors2, aes(Map.Method, NEAR_DIST, colour=Map.Method))+
  geom_boxplot()+
  facet_wrap(~InstNear)+
  theme_bw()+
  xlab("Mapping method")+
  ylab("Distance to closest mapped occurence")
```

Assuming CNHP has accurate and complete data, then Distance to a known mapped location 
```{r}
table(eors2$Map.Method, eors2$InstNear)

length(table(eors2$scientific)) #53

#species are the names collected from SEINet
match(sort(unique(eors2$scientific)),sort(species))

InEOR <- do.call(rbind,lapply(split(eors2, eors2$scientific), function(x){
  out <- x[x$NEAR_DIST ==0,]
  out
})
)

head(InEOR)
table(InEOR$recordedBy)

# $georeferen is the only place to know who did the georeferencing, that would be very helpful to know
```


The Weber example of an out of state person mapping a collection to Gray's Peak from Chicago Lake that should be Mount Evans.     
- <https://conps.org/wp-content/uploads/2015/05/Mount_Evans_Summit_Lake_4-24-2011.pdf>    
```{r}


```






