---
title: "Herbarium Bias CNHP"
author: "Michelle DePrenger-Levin"
date: "December 4, 2017"
output: html_document
---

Rick says don't need quotes for table or column names

MySQL query

#There are only rows in the synonym_links table when the names are synonyms, if it's not in synonym_links, it's the accepted name; taxonomic_units is the main table, if it's a synonym it's linked by synonym_links
SELECT `synonym_links`.*, `taxonomic_units`.* , `taxon_authors_lkp`.*, `geographic_div`.*
FROM `taxonomic_units` 
LEFT JOIN `synonym_links` ON `synonym_links`.`tsn` = `taxonomic_units`.`tsn` 
LEFT JOIN `taxon_authors_lkp` ON `taxon_authors_lkp`.`taxon_author_id` = `taxonomic_units`.`taxon_author_id` 
LEFT JOIN `geographic_div` ON `geographic_div`.`tsn` = `taxonomic_units`.`tsn`
WHERE `taxonomic_units`.`kingdom_id` = 3 AND `taxonomic_units`.`rank_id` = 220 AND `geographic_div`.`geographic_value` IN ('North America')



#The authorities are giving me trouble right now; can do NOT IN (0,3,9999) or whatever
SELECT `synonym_links`.*, `taxonomic_units`.* , `taxon_authors_lkp`.*
FROM `taxonomic_units` 
JOIN `synonym_links` ON `synonym_links`.`tsn` = `taxonomic_units`.`tsn` 
JOIN `taxon_authors_lkp` ON `taxon_authors_lkp`.`taxon_author_id` = `taxonomic_units`.`taxon_author_id` 
WHERE `taxonomic_units`.`kingdom_id` = 3 AND `taxonomic_units`.`rank_id` = 220 AND `taxonomic_units`.`unaccept_reason` IS NULL AND `taxonomic_units`.`taxon_author_id` NOT IN (0) AND `taxonomic_units`.`name_usage` NOT IN ('invalid') AND `taxonomic_units`.`n_usage` IN ('accepted')



#the taxon_authors_lkp.taxon_author_id isn't the tsn, it's the taxonomic_units.taxon_author_id!!!
SELECT `synonym_links`.*, `taxon_authors_lkp`.*, `taxonomic_units`.* 
FROM `taxonomic_units` 
LEFT JOIN `synonym_links` ON `synonym_links`.`tsn` = `taxonomic_units`.`tsn`
LEFT JOIN `taxon_authors_lkp` ON `taxon_authors_lkp`.`taxon_author_id` = `taxonomic_units`.`taxon_author_id` 
WHERE `taxonomic_units`.`kingdom_id` = 3 AND `taxonomic_units`.`rank_id` = 220 




```{r}
rm(list=ls())

library(ggplot2)
library(rgeos)
library(sp)
library(spdep)
library(rgdal)
library(maptools)
library(raster)
library(Taxonstand)
library(mapproj)
library(maps)

# exact match multisub
multisub <- function(pattern,replacement,x){
  result <- x
  for(i in 1:length(pattern)){
    result <- gsub(paste("^",pattern[i],"$"), paste("^",replacement[i],"$"),result)
  }
  result
}


lseq <- function(from=1, to=100000, length.out=6) {
  # logarithmic spaced sequence
  # blatantly stolen from library("emdbook"), because need only this
  exp(seq(log(from), log(to), length.out = length.out))
}
```

Mo Ewing:    
     1) split all polygons (from CNHP) with ArcGIS 'explosion' tool    
     2) Copy the Feature_id field into the Parcel-num field and annotate each parcel: "a", "b" etc    
     3) Import G1G2 Seinet Vouchers into ArcMap      
     4) Create linking field: "Parcel_num"          
     5) Go through each parcel, linking the Eos by duplicating the Parcel_num.    
     6) determin the closest 'parcel' to each herabarium point using ArcMap 'near' tool Analyst Tools/Proximity/Near   *I assume to the nearest line of the polygon, nearest spot of the polygon, not to the center of the polygon          
     7) Add Map Method informaiton     
            i. "GPS" = indication that GPS was used, just GPS coordinates present    
            ii. "tr" = township/range info only recorded to township and range     
            iii. "trs" = township/Range info recorded to section, or indication that tr conversion tool used       
            iv. "trqs" = Township/Range info recorded to quarter-section       
            v. "Locality" = no verbatim data recorded, only written description recorded    
                     or "Terrain nav" (terrain navigator used)    
                     or "Nat. Geog. TOPO"    
                     or "Mapstedi"     
                     or "digital mape"     
                     or "geoLocate"     
                     or "GoogleEarth"      



# Mo worked on this list to add map method and distance
```{r}
seinet <- read.csv("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/SEINET G1 G2 Vouchers/seinet_colorado_with_coordinates_ALLG1G2.csv")

seinetMo <- TPL(unique(seinet$scientificName))
sum(table(seinetMo$Taxonomic.status)) #93 names
#Mo has 2 names not in TPL, 9 synonyms, 7 unresolved
seinetMo[seinetMo$Taxonomic.status != "Accepted",]
seinet[grep("rotundifo", seinet$scientificName),]

length(table(seinet$scientificName[!is.na(seinet$decimalLatitude)])) #93 species with locaiton information

eors <- read.csv("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/Distance Tables/Vouchers_distance.csv")

length(unique(eors$scientific)) #53
eors.tax <- TPL(unique(eors$scientific))
table(eors.tax$Taxonomic.status)
# 5 syn, 2 unresolved, 1 not in

#eors[eors$institutio=="USU",]
#eors[eors$institutio=="USUUB",]
eors$institutio[eors$institutio=="USU"] <- "USUUB"

```


#Michelle DePrenger-Levin      
In ArcMap do this: <https://blogs.esri.com/esri/arcgis/2010/09/16/nearbygroup/>
repeat distance analysis in R          
SEINet collections from Colorado - downloaded 12/6/2017 from SEINet        
Only G1G2 EORs (to limit range of species)   
Change names according to TPL as a best effort to match up.    
Only this first chunck if you want to redo naming changes- otherwise SKIP TO NEXT        
```{r}
coloradosps <- read.csv("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/SEINET G1 G2 Vouchers/occurrences.csv")

#colonames <- TPL(unique(coloradosps$scientificName))
#write.csv(colonames, "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/AllcoloSEINet_TPL.csv")
colonames <- read.csv("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/AllcoloSEINet_TPL.csv")

sum(table(colonames$Taxonomic.status))
# 1929 not in The Plant List, 6 misapplied??, lots of synonyms. 
#             Accepted Misapplied    Synonym Unresolved 
#      1929       9149          6       6083        682
colonames[colonames$Taxonomic.status == "Misapplied",]

head(colonames[colonames$Taxonomic.status == "Misapplied",])

fixcolonames <- colonames[colonames$Taxonomic.status == "Synonym",]

coloradosps$genus <- multisub(fixcolonames$Genus, fixcolonames$New.Genus, coloradosps$genus) 
coloradosps$specificEpithet <- multisub(fixcolonames$Species, 
                                   fixcolonames$New.Species, coloradosps$specificEpithet) 

write.csv(coloradosps, "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/coloradosps_synonymsTPL.csv")
```

Skip to here to keep name changes from above
```{r}
coloradosps <- read.csv("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/coloradosps_synonymsTPL.csv")

coloradosps$institutionCode[coloradosps$institutionCode == "USU"] <- "USUUB"

l1eor <- readShapePoly("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/L1shp")

#Does not read th .prj file so need to tell it that they're in UTM
proj4string(l1eor) <- "+proj=utm +zone=13 ellps=WGS84"
```


#Skip this because want split by species, not labeled and match by label. 
```{r}
#Each polygon might be made of multiple disconnected polygons. Need to separate and label
l1G1G2 <- disaggregate(l1eor[l1eor$GRANK %in% c("G1","G2"),])
l1G1G2$PolyID <- do.call(rbind, lapply(split(l1G1G2,l1G1G2$OBJECTID),
                                       function(x){
              x$POlyID <- paste(x$EO_ID, LETTERS[seq(from=1, to= nrow(x))],sep="")
              x
              }))
```

#Skip, don't seem to link any synonyms
```{r}
#The Plant List
sp43 <- TPL(intersect(speciesG1G2, species))
table(sp43$Taxonomic.status)
syn <- sp43[sp43$Taxonomic.status == "Synonym",]

seinet$genus <- multisub(syn$Genus, syn$New.Genus, seinet$genus) 
seinet$specificEpithet <- multisub(syn$Species, syn$New.Species, seinet$specificEpithet) 

species <- unique(seinet$scientificName)
sort(intersect(species, speciesG1G2))
```





Download all Colorado collections from SEINet (coloradosps) and limit to matching G1 G2 global names by CNHP. Will be missing synonyms
```{r}
#Without synonym corrections
#coloG1G2 <- coloradosps[coloradosps$scientificName %in% 
#                          unique(l1eor$GNAME[l1eor$GRANK %in% c("G1","G2")]),]

#write.csv(coloG1G2, "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/occurrencesG1G2_nosynonyms.csv")

coloG1G2 <- read.csv("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/occurrencesG1G2_nosynonyms.csv")

nrow(coloG1G2) #2125
length(unique(coloG1G2$scientificName)) #59
```

#Can skip this too
```{r}
nrow(seinet[!is.na(seinet$decimalLatitude),]) #1875 points

#Get all points, transform to UTM
pts <- SpatialPoints(seinet[,c("decimalLongitude","decimalLatitude")])
proj4string(pts) <- CRS("+proj=longlat +datum=WGS84")

#Project data into a planar coordinate system (UTM zone 13)
UTMpts <- spTransform(pts, CRS("+proj=utm +zone=13 ellps=WGS84"))

head(pts)
head(UTMpts)

plot(l1eor[l1eor$GRANK %in% c("G1","G2"),])
points(UTMpts)
names(l1eor)

plot(l1eor[l1eor$GNAME == "Astragalus microcymbus",])

#Does not read th .prj file so need to tell it that they're in UTM
proj4string(l1eor) <- "+proj=utm +zone=13 ellps=WGS84"

#might be tooooo big!
#PLSS <- readOGR(dsn="Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/Colorado/PLSS Geo Coordinates_2008")

PLSS <- shapefile("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/Colorado/PLSS Geo Coordinates_2008/gcdb_mm_2008_polygon.shp")

```


Try to do an apply and split by species, don't really care which EO, just which mapped occurance of a species is closest     
```{r}
#59 species
nms <- intersect(unique(seinet$scientificName),unique(l1eor$GNAME)) #67 names

mindistxsp <- lapply(nms, function(x){
  pts.sp <- SpatialPoints(seinet[seinet$scientificName == x,
                                 c("decimalLongitude","decimalLatitude")])
  proj4string(pts.sp) <- CRS("+proj=longlat +datum=WGS84")
  UTMpts.sp <- spTransform(pts.sp,CRS("+proj=utm +zone=13 ellps=WGS84"))
  
  l1.sp <- disaggregate(l1eor[l1eor$GNAME == x,])
  
  distout <- apply(gDistance(UTMpts.sp, l1.sp, byid=TRUE),2,min)
  out <- data.frame(distout,seinet[seinet$scientificName == x,])
  list(out,UTMpts.sp,l1.sp)
})

mndistsp <- do.call(rbind,lapply(mindistxsp,'[[',1))

#Need UTM layer to match points
colomap<-  shapefile("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/Colorado/CO_Counties/counties.shp")

for(i in 1:length(mindistxsp)){
  plot(mindistxsp[[i]][[3]],
     xlim=c(142063.3, 740922.7), ylim=c(4094495, 4543961), 
     border="red",cex=5,lwd=3)
  points(mindistxsp[[i]][[2]],col="blue")
  points(gCentroid(colomap, byid=TRUE),col="green",pch=5)
  plot(colomap,add=TRUE)
  }

```

#A number of these points do line up, is this different by species or by herbarium or by year? 
```{r}
cntycenters <- coordinates(gCentroid(colomap, byid=TRUE))
str(gCentroid(colomap, byid=TRUE)@coords)

#Fortifying a map makes the data to draw the outline
#Bet I want to disaggregate first then fortify??
colomap.fort.pieces <- fortify(disaggregate(colomap))

colomap.fort <- fortify(colomap)
countylist <- colomap@data$COUNTY

#"coordinates" extracts the centroids
centroids.colo <- as.data.frame(coordinates(colomap))
data.frame(countylist,centroids.colo)


#Polygons from CNHP
mindistxsp[[1]][[3]]@data #These are polygons split, disaggregate
mindistxsp[[1]][[2]]@coords #now converted to UTM
# all polygons
mappolys <- do.call(rbind, lapply(1:length(mindistxsp), function(x){
  out.fort <- data.frame(fortify(mindistxsp[[x]][[3]]),nms[x])
  out.fort
}))


#all points
mapdata <- do.call(rbind,lapply(1:length(mindistxsp), function(x){
  out <- data.frame(mindistxsp[[x]][[2]]@coords,nms[x])
  names(out) <- c("Easting","Northing","Species")
  out
}))

extent(colomap)
#class       : Extent 
#xmin        : 139982.4 
#xmax        : 763240.8 
#ymin        : 4094070 
#ymax        : 4546718 

mapdata2 <- mapdata[mapdata$Northing<4546718&
                      mapdata$Northing>4094070&
                      mapdata$Easting>139982.4&
                      mapdata$Easting<763240.8,]
# Easting   Northing  Species
# -4429510  6010981	  Draba grayana
# 282427	  3040916	  Astragalus iodopetalus
# -458653.4	4218252	  Packera mancosana

```



not working all together     
<https://stackoverflow.com/questions/22038640/labeling-center-of-map-polygons-in-r-ggplot> 
```{r}
ggplot()+
  geom_point(data=mapdata2, aes(Easting,Northing))+
  theme_bw()+
  geom_polygon(data=colomap, aes(x=long,y=lat,fill=as.factor(order)),
               alpha=0.25)+
  theme(legend.position = "none")

#  geom_polygon(data=colomap.fort.pieces, aes(x=long,y=lat,group=id), alpha=0,colour=order)+
#  coord_map()

```

mndistsp is the distance from each herbarium point to the closest CHNP polygon of that speices; a dataframe of original data and distances.            
class(mindistxsp[[1]][[2]]) #SpatialPoints     


# 1) are they in the county they say they are in?
```{r}
#map seinet points to match colo counties UTM projection 
pts.all <- SpatialPoints(seinet[,c("decimalLongitude","decimalLatitude")])
proj4string(pts.all) <- CRS("+proj=longlat +datum=WGS84")
pts.all <- spTransform(pts.all,CRS(" +proj=utm +zone=13 +datum=NAD83 +units=m +no_defs +ellps=GRS80 +towgs84=0,0,0 "))

incounty.colomap <- over(pts.all, colomap)
#to do comparison later, find anywhere that correct county is mentioned
seinet$county <- as.character(seinet$county)

seinet <- data.frame(seinet,incounty.colomap)

head(seinet[,c("county","COUNTY","decimalLongitude","decimalLatitude","scientificName","G1G2_Speci")],20)

#now when is the county that the point is in match the county name in the $county
for(i in 1:nrow(seinet)){
  print(seinet$pointincounty[i])
}

head(seinet[is.na(seinet$pointincounty),])
#change all factor columns to character
i <- sapply(seinet, is.factor)
head(seinet[,i]) #recordNumber might be the only one needing to be numeric
seinet[,i] <- lapply(seinet[,i], as.character)

#pulls out matches and when the point wasn't in any county
correctcounty <- lapply(1:nrow(seinet),function(x){
  r<-seinet[x,]
  if(!is.na(r$COUNTY)){
    matches <- grep(r$COUNTY,r[-89],ignore.case=TRUE,value=TRUE)
    if(length(matches)>0){
      out <- list(seinet[x,],matches)
      out
    }
  }})

#subset out nulls
#matchout <- correctcounty[!vapply(correctcounty, is.null, logical(1))]
length(matchout) #1793


correctcounty[[1]][1:20]

matched <- do.call(rbind,lapply(1:nrow(seinet), function(row){
   r<-seinet[row,]
  if(!is.na(r$COUNTY)){
    matches <- grep(r$COUNTY,r[-89],ignore.case=TRUE,value=TRUE)
    if(length(matches)>0){
      r
    }
  }}
  ))

head(matched[,c("county","COUNTY")],40)
nrow(matched)/ #1793 out of the 1875 rows of seinet
nrow(seinet[!is.na(seinet$decimalLatitude),]) #1875 have a point somewhere

notmatched <- do.call(rbind,lapply(1:nrow(seinet), function(row){
   r<-seinet[row,]
  if(!is.na(r$COUNTY)){
    matches <- grep(r$COUNTY,r[-89],ignore.case=TRUE,value=TRUE)
    if(length(matches)==0){ #Want rows where no other column has correct county info
      r
    }
  }}
  ))

head(notmatched[,c("county","COUNTY")],40)
nrow(notmatched)

```
  1  51  52  53  85  86 145 153 187 188 483 629 655 
 NA   4   4   4   4   4  34   4  28  28  27  27  27 

              county 
1           Hinsdale             
51  Archuleta County             
52         Archuleta             
53         Archuleta             
85  Archuleta County             
86         Archuleta             
145     La Plata Co.             
153 Archuleta County             
187         Hinsdale             
188         Hinsdale             
483         Gunnison  

```{r}
#Fair to remove points mapped outside of Colorado when supposed to be in Colorado
distNout <- mndistsp[mndistsp$decimalLatitude<41&
                       mndistsp$decimalLatitude>37&
                       mndistsp$decimalLongitude<=-102&
                       mndistsp$decimalLongitude>=-109,]

#More are mapped but more are mapped farther away in later years
ggplot(distNout, aes(year,distout,colour=genus))+
  geom_point()+
  theme_bw()
```

```{r}
distNout$DistanceGroup <- cut(distNout$distout,
                              breaks=c(-0.001,lseq(1,(max(distNout$distout))+0.001)),
                              labels= c(1:6))
  

# Who has no order? 
table(distNout$order)
distNout[distNout$order == "",]

#From Plants USDA
#Add factor levels
levels(distNout$order)
levels(distNout$order) <- c(levels(distNout$order),"Capparales")

distNout$order[distNout$order == ""] <- c("Asterales","Capparales")

# As you get older, more distant (should be function of infomration avaliable) not consistant among Orders. 
ggplot(distNout, aes(DistanceGroup,year,colour=order))+
  geom_boxplot()+
  theme_bw()

ggplot(distNout, aes(y=distout,x=institutionCode,fill=institutionCode))+
  geom_boxplot()+
  theme_bw()

names(distNout)
```


```{r}
local <- c("COLO","CS","DBG","FLD","MESA","RM","RMBL","SEINet")
Adjacent <- c("ARIZ","ASC","ASU","BRY","DES","MNA","NAVA","NMC","RM","SJNM",
              "UNM","USUUB","USU","UVSC","UT") #USU doesn't exist, should be USUUB
far <- c("CHIC","CM","EIU","F","MARY","MISSA","MO","MWI","NY","RENO","SDSU",
         "UCR")

distNout$InstNear[distNout$institutio %in% local] <- "Local"
distNout$InstNear[distNout$institutio %in% Adjacent] <- "Near"
distNout$InstNear[distNout$institutio %in% far] <- "Far"

distNout[is.na(distNout$InstNear),]

ggplot(distNout, aes(y=distout,x=InstNear))+ #fill=institutionCode
  geom_boxplot()+
  theme_bw()

anova(lm(distout~InstNear, data=distNout))
```

##START HERE!
```{r}
#locality


#verbatimCoordinates
# That have Township
nrow(distNout[grep("[T]{1}[0-9]{2}[NS]{1}",distNout$verbatimCoordinates),]) #462
head(distNout[grep("[T]{1}[0-9]{2}[NS]{1}",distNout$verbatimCoordinates),])
#Township and Range and Section
nrow(distNout[grep("[S]{1}[0-9]*",distNout$verbatimCoordinates),]) #780
head(distNout[grep("[S]{1}[0-9]*",distNout$verbatimCoordinates),])
table(distNout$locality[grep("[S]{1}[0-9]*",distNout$verbatimCoordinates)])
head(distNout$verbatimCoordinates[grep("[S]{1}[0-9]*",distNout$verbatimCoordinates)])

#georeferenceRemarks: ie. "UTM used is gnis UTM for ..."


# how they did it from georeferenceSources ie. "GPS Reading"


#georeferenceProtocol i.e. "Mapstedi"


# some say who did the georeferencing in georeferencedBy
ggplot(distNout[distNout$georeferencedBy != "",], 
       aes(y=distout,x=georeferencedBy))+
  geom_boxplot()+
  theme_bw()

distNout$georeferencedBy[distNout$distout>50000] #Neil Snow, Ryan Allen, Michael Yost, David Newbury

sort(table(distNout$georeferencedBy[distNout$distout==0])) #Ryan Allen, Leslie Johnson, Todd Fielder, Zorio, Stephanie D., Willimams, Charles F., Amanda Cook, Allison Smith, Robert L. Johnson, Rick Levy,  C.F. Williams , Charles F. Williams, David Newbury, Fertig, Leslie Johnson

```






# Have mndistsp is the minimum distances from each point (from SEINet gather all collections from Colorado; match names that are G1G2 with names in SEINet; will be missing any synonyms)   
Unlike eors2 where Mo added a label for method used or accuracy of mapping, need to add to mndistsp     
I don't think we can tell how much infomration there really was without looking at the specimens.   
```{r}
table(mndistsp$year) #1871 to 2017
#Find degree symbol
head(mndistsp[grep(intToUtf8(176),mndistsp$locality),])
mndistsp[grep("[T]{1}[0-9]{2}[NS]{1}",mndistsp$locality),] #Lists TRS but also other UTM or lat long

table(mndistsp$verbatimCoordinates[grep()])


mndistsp[mndistsp$year == 2017 & !is.na(mndistsp$year),] #Eriogonum coloradense

```

```{r}
ggplot(mndistsp[mndistsp$distout<1000000,], aes(scientificName,distout))+
  geom_boxplot()+ 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


```




# How do the maping techniques change by time, herbarium, or collector (amount of informaiton given)?   
```{r}
eors2 <- eors[eors$Year.2>0,]


methodxyear <- data.frame(table(eors2$Map.Method, eors2$Year.2))
methodxyear$Var2 <- as.numeric(as.character(methodxyear$Var2))

eors2$Map.Method <- factor(eors2$Map.Method, levels = c("Locality","GPS","trqs",
                                                        "trs","tr"))

ggplot(methodxyear, aes(Var2, Freq, colour=Var1))+
  geom_point()+
  stat_smooth(se=FALSE)+
  theme_bw()+
  xlab("Year")+
  ylab("Herbarium collections")+
  labs(color='Georeference Method')+
  theme(legend.justification=c(0,1), legend.position=c(0,1),
        legend.background = element_blank()) 
```


```{r}
ggplot(eors[eors$Year.2>0,], aes( Year.2,NEAR_DIST,colour=institutio))+
  geom_point()+
  facet_wrap(~institutio)+
  theme_bw()+
  stat_smooth(method="lm", se=FALSE)

ggplot(eors[eors$Year.2>0,], aes( Year.2,NEAR_DIST))+
  geom_point()+
  theme_bw()+
  stat_smooth(method="lm", se=FALSE)

summary(lm(NEAR_DIST~Year.2, data=eors[eors$Year.2>0,]))
```

How did the type of mapping method change over time?
```{r}

table(eors$Year.2)



```

Which method has the shortest distance?   
```{r}


ggplot(eors2, aes(x=Map.Method))+
  geom_bar(aes(fill=institutio))


ggplot(eors2, aes(Year.2,NEAR_DIST, colour=institutio))+
  geom_point()+
  theme_bw()

ggplot(eors2, aes(x=Year.2, fill=Map.Method))+
  geom_density(alpha=0.5)+
  theme_bw()
#+  facet_wrap(~Map.Method)

ggplot(eors2, aes(x=Year.2, fill=Map.Method))+
  geom_histogram(alpha=0.25)+
  theme_bw()


```

Maybe regional collections are much better?   
FLD = Fort Lewis College    
DES = Dessert Arizona   
MESA = Colorado Mesa University   
MISSA = Miss State Univ   
NMC = New Mexico State University    
SJNM = San Juan College, New Mexico   
UCR = University of California Riverside    
USUUB = Utah STate University Uintah Basin (probably is USU)   
UVSC = Utah Valley University	U.S.A. Utah. Orem.      
```{r}

local <- c("COLO","CS","DBG","FLD","MESA","RM","RMBL")
Adjacent <- c("ASC","ASU","BRY","DES","NMC","RM","SJNM","UNM","USUUB","UVSC")
far <- c("CM","F","MISSA","MO","NY","RENO","UCR")

eors2$InstNear[eors2$institutio %in% local] <- "Local"
eors2$InstNear[eors2$institutio %in% Adjacent] <- "Near"
eors2$InstNear[eors2$institutio %in% far] <- "Far"

eors2$InstNear <- factor(eors2$InstNear, levels = c("Far","Near","Local"))


ggplot(eors, aes(x=institutio))+
  geom_bar(aes(fill=Map.Method))


ggplot(eors2, aes(x=InstNear))+
  geom_bar(aes(fill=Map.Method))

ggplot(eors2, aes(Year.2,NEAR_DIST, colour=InstNear))+
  geom_point()+
  theme_bw()+
  stat_smooth(method="lm", se=FALSE)


ggplot(eors2, aes(NEAR_DIST))+
  geom_density(aes(fill=InstNear))
```

Eastern Illinois Univeristy (Charleston)     
FLD Fort Lewis College     
MARY University of Maryland    
MNA Museum of Northern Arizona     
MWI Western Illinois Univeristy   
NAVA Navajo Nation Department of FIsh and Wildlife     
SDSU San Diego State Univ    
SEINet: error in reporting??    
```{r}
local <- c("COLO","CS","DBG","FLD","MESA","RM","RMBL","SEINet")
Adjacent <- c("ARIZ","ASC","ASU","BRY","DES","MNA","NAVA","NMC","RM","SJNM","UNM","USUUB","UVSC",
              "UT")
far <- c("CHIC","CM","EIU","F","MARY","MISSA","MO","MWI","NY","RENO","SDSU","UCR")

noquote(paste(local,sep=","))

table(mndistsp$institutionCode)
mndistsp[mndistsp$institutionCode=="SEINet",] #"Human observation", Shelley Silva
mndistsp[mndistsp$institutio=="USU",]
mndistsp[mndistsp$institutio=="USUUB",]
mndistsp$institutio[mndistsp$institutio=="USU"] <- "USUUB"

mndistsp$InstNear[mndistsp$institutio %in% local] <- "Local"
mndistsp$InstNear[mndistsp$institutio %in% Adjacent] <- "Near"
mndistsp$InstNear[mndistsp$institutio %in% far] <- "Far"

mndistsp$InstNear <- factor(mndistsp$InstNear, levels = c("Far","Near","Local"))



```


```{r}
ggplot(mndistsp, aes(InstNear,log(distout+1),fill=DistanceGroup))+
  geom_boxplot()+
  theme_bw()

ggplot(mndistsp, aes(InstNear, log(distout+1)))+
  geom_boxplot()
anova(lm(log(distout+1)~InstNear, data=mndistsp))


ggplot(mndistsp, aes(InstNear, log(distout+1), colour=year))+
  geom_point()
summary(lm(log(distout+1)~InstNear*year, data=mndistsp))


summary(lm(log(distout+1)~year, data=mndistsp))
```


Would assume that Locality method would be best for local institutions, worse for far
```{r}


ggplot(eors2, aes(Map.Method, NEAR_DIST, colour=Map.Method))+
  geom_boxplot()+
  facet_wrap(~InstNear)+
  theme_bw()+
  xlab("Mapping method")+
  ylab("Distance to closest mapped occurence")
```

Assuming CNHP has accurate and complete data, then Distance to a known mapped location 
```{r}
table(eors2$Map.Method, eors2$InstNear)

length(table(eors2$scientific)) #53

#species are the names collected from SEINet
match(sort(unique(eors2$scientific)),sort(species))

InEOR <- do.call(rbind,lapply(split(eors2, eors2$scientific), function(x){
  out <- x[x$NEAR_DIST ==0,]
  out
})
)

head(InEOR)
table(InEOR$recordedBy)

# $georeferen is the only place to know who did the georeferencing, that would be very helpful to know
```


The Weber example of an out of state person mapping a collection to Gray's Peak from Chicago Lake that should be Mount Evans.     
- <https://conps.org/wp-content/uploads/2015/05/Mount_Evans_Summit_Lake_4-24-2011.pdf>    
```{r}


```






