---
title: "Herbarium Bias CNHP"
author: "Michelle DePrenger-Levin"
date: "December 4, 2017"
output: html_document
---
Rick says don't need quotes for table or column names

MySQL query

#There are only rows in the synonym_links table when the names are synonyms, if it's not in synonym_links, it's the accepted name; taxonomic_units is the main table, if it's a synonym it's linked by synonym_links
SELECT `synonym_links`.*, `taxonomic_units`.* , `taxon_authors_lkp`.*, `geographic_div`.*
FROM `taxonomic_units` 
LEFT JOIN `synonym_links` ON `synonym_links`.`tsn` = `taxonomic_units`.`tsn` 
LEFT JOIN `taxon_authors_lkp` ON `taxon_authors_lkp`.`taxon_author_id` = `taxonomic_units`.`taxon_author_id` 
LEFT JOIN `geographic_div` ON `geographic_div`.`tsn` = `taxonomic_units`.`tsn`
WHERE `taxonomic_units`.`kingdom_id` = 3 AND `taxonomic_units`.`rank_id` = 220 AND `geographic_div`.`geographic_value` IN ('North America')



#The authorities are giving me trouble right now; can do NOT IN (0,3,9999) or whatever
SELECT `synonym_links`.*, `taxonomic_units`.* , `taxon_authors_lkp`.*
FROM `taxonomic_units` 
JOIN `synonym_links` ON `synonym_links`.`tsn` = `taxonomic_units`.`tsn` 
JOIN `taxon_authors_lkp` ON `taxon_authors_lkp`.`taxon_author_id` = `taxonomic_units`.`taxon_author_id` 
WHERE `taxonomic_units`.`kingdom_id` = 3 AND `taxonomic_units`.`rank_id` = 220 AND `taxonomic_units`.`unaccept_reason` IS NULL AND `taxonomic_units`.`taxon_author_id` NOT IN (0) AND `taxonomic_units`.`name_usage` NOT IN ('invalid') AND `taxonomic_units`.`n_usage` IN ('accepted')



#the taxon_authors_lkp.taxon_author_id isn't the tsn, it's the taxonomic_units.taxon_author_id!!!
SELECT `synonym_links`.*, `taxon_authors_lkp`.*, `taxonomic_units`.* 
FROM `taxonomic_units` 
LEFT JOIN `synonym_links` ON `synonym_links`.`tsn` = `taxonomic_units`.`tsn`
LEFT JOIN `taxon_authors_lkp` ON `taxon_authors_lkp`.`taxon_author_id` = `taxonomic_units`.`taxon_author_id` 
WHERE `taxonomic_units`.`kingdom_id` = 3 AND `taxonomic_units`.`rank_id` = 220 



SELECT `synonym_links`.*, `taxon_authors_lkp`.*, `taxonomic_units`.* 
FROM `taxonomic_units` 
LEFT JOIN `synonym_links` ON `synonym_links`.`tsn` = `taxonomic_units`.`tsn`
LEFT JOIN `taxon_authors_lkp` ON `taxon_authors_lkp`.`taxon_author_id` = `taxonomic_units`.`tsn` 
WHERE `taxonomic_units`.`kingdom_id` = 3 AND `taxonomic_units`.`rank_id` = 220 



cd /p/hackathon/Simulations/

```{r}
rm(list=ls())

library(ggplot2)
library(rgeos)
library(sp)
library(spdep)
library(rgdal)
library(maptools)
library(raster)
library(Taxonstand)
library(RCurl)
library(taxize)

# exact match multisub
multisub <- function(pattern,replacement,x){
  result <- x
  for(i in 1:length(pattern)){
    result <- gsub(paste("^",pattern[i],"$"), paste("^",replacement[i],"$"),result)
  }
  result
}

```

Mo Ewing:    
     1) split all polygons (from CNHP) with ArcGIS 'explosion' tool    
     2) Copy the Feature_id field into the Parcel-num field and annotate each parcel: "a", "b" etc    
     3) Import G1G2 Seinet Vouchers into ArcMap      
     4) Create linking field: "Parcel_num"          
     5) Go through each parcel, linking the Eos by duplicating the Parcel_num.    
     6) determin the closest 'parcel' to each herabarium point using ArcMap 'near' tool Analyst Tools/Proximity/Near   *I assume to the nearest line of the polygon, nearest spot of the polygon, not to the center of the polygon          
     7) Add Map Method informaiton     
            i. "GPS" = indication that GPS was used, just GPS coordinates present    
            ii. "tr" = township/range info only recorded to township and range     
            iii. "trs" = township/Range info recorded to section, or indication that tr conversion tool used       
            iv. "trqs" = Township/Range info recorded to quarter-section       
            v. "Locality" = no verbatim data recorded, only written description recorded    
                     or "Terrain nav" (terrain navigator used)    
                     or "Nat. Geog. TOPO"    
                     or "Mapstedi"     
                     or "digital mape"     
                     or "geoLocate"     
                     or "GoogleEarth"      




```{r}
seinet_Ewingdistances <- read.csv("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/Distance Tables/Vouchers_distance.csv")

seinet <- read.csv("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/SEINET G1 G2 Vouchers/seinet_colorado_with_coordinates_ALLG1G2.csv") #I must have made this one sometime, need to check again that matched all names

#seinetMo <- TPL(unique(seinet$scientificName))
#write.csv(seinetMo,"Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/Distance Tables/TPLseinetMo.csv")
seinetMo <- read.csv("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/Distance Tables/TPLseinetMo.csv")
sum(table(seinetMo$Taxonomic.status)) #93 names
#Mo has 2 names not in TPL, 9 synonyms, 7 unresolved
seinetMo[seinetMo$Taxonomic.status != "Accepted",]
seinet[grep("rotundifo", seinet$scientificName),]

length(table(seinet$scientificName[!is.na(seinet$decimalLatitude)])) #93 species with locaiton information

eors <- read.csv("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/Distance Tables/Vouchers_distance.csv")

length(unique(eors$scientific)) #53 - Mo going through 1 by 1 getting SEINet data for G1G2 species ended up with 53
eors.tax <- TPL(unique(eors$scientific))
table(eors.tax$Taxonomic.status)
# 5 syn, 2 unresolved, 1 not in

eors[eors$institutio=="USU",]
eors[eors$institutio=="USUUB",]


eors$institutio[eors$institutio=="USU"] <- "USUUB"

#change all factor columns to character
i <- sapply(seinet, is.factor)
head(seinet[,i]) #recordNumber might be the only one needing to be numeric
seinet[,i] <- lapply(seinet[,i], as.character)
```


#Michelle DePrenger-Levin      
In ArcMap do this: <https://blogs.esri.com/esri/arcgis/2010/09/16/nearbygroup/>
repeat distance analysis in R          
"coloradosps" == SEINet collections from Colorado - downloaded 12/6/2017 from SEINet        
Only G1G2 EORs (to limit range of species)
"colonames" == TPL synonomy for all coloradosps names
```{r}
#coloradosps <- read.csv("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/SEINET G1 G2 Vouchers/occurrences.csv")

#save(coloradosps, file= "P:/hackathon/Simulations/coloradosps.Rda")
load("P:/hackathon/Simulations/coloradosps.Rda")

#colonames <- TPL(unique(coloradosps$scientificName))
#write.csv(colonames, "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/AllcoloSEINet_TPL.csv")
#colonames <- read.csv("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/AllcoloSEINet_TPL.csv")
#save(colonames,file="P:/hackathon/Simulations/colonames.Rda")
load("P:/hackathon/Simulations/colonames.Rda")

colonames <- colonames[-grep("Middle Park",colonames$Taxon),] #get rid of offset record
colonames <- colonames[!(colonames$Taxon %in% ""),] #get rid of row with no name

#merge taxon from colonames (with TPL categorization of name) with coloradosps$scientificName
coloradosps$scientificName <- as.character(coloradosps$scientificName)
colonames$Taxon <- as.character(colonames$Taxon)

coloradosps$Taxon <- paste(coloradosps$genus,coloradosps$specificEpithet,
                           coloradosps$taxonRank,coloradosps$infraspecificEpithet,
                           sep=" ")

#remove any leading or trailing white spaces around both names
coloradosps$Taxon <- gsub("\\s+$","",coloradosps$Taxon)
#colonames$Taxon <- gsub("\\s+","",colonames$Taxon) # There are 390 repeated names, why?
coloradospsnames <- merge(coloradosps,colonames,by="Taxon")


sum(table(colonames$Taxonomic.status))
# 1929 not in The Plant List, 6 misapplied??, lots of synonyms. 
# Some are just Genus. 
#             Accepted Misapplied    Synonym Unresolved 
#      1929       9149          6       6083        682
head(unique(colonames$Taxon[colonames$Taxonomic.status == ""]),30)
table(coloradospsnames$Taxonomic.status)


l1eor <- readShapePoly("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/L1shp")

#Each polygon might be made of multiple disconnected polygons. Need to separate and label
l1G1G2 <- disaggregate(l1eor[l1eor$GRANK %in% c("G1","G2"),])
l1G1G2$PolyID <- do.call(rbind, lapply(split(l1G1G2,l1G1G2$OBJECTID),
                                       function(x){
              x$POlyID <- paste(x$EO_ID, LETTERS[seq(from=1, to= nrow(x))],sep="")
              x
              }))

#How many names in 
length(unique(l1G1G2$GNAME)) #60

# Could take all the synonyms and names of l1G1G2 to make the list of names to pull from SEINet big list
namesg1g2 <- table(l1G1G2$GNAME)
length(namesg1g2[namesg1g2 > 0]) #60 species
g1g2names <- TPL(names(namesg1g2[namesg1g2 > 0]))
table(g1g2names$Taxonomic.status) #2 not in there, "Gutierrezia elegans" "Oonopsis sp. 1", 10 synonyms, 3 unresolved

#Are both synonyms and accepted in as GNAME?
notaccepted <- l1G1G2$GNAME[!is.na(match(l1G1G2$GNAME,
      g1g2names$Taxon[g1g2names$Taxonomic.status != "Accepted"]))]
unique(notaccepted)
g1g2names[g1g2names$Taxonomic.status != "Accepted",]

unique(l1G1G2$GNAME[grep("Boechera",l1G1G2$GNAME)])

#Are any accepted names also in the list?
l1G1G2$GNAME[l1G1G2$GNAME %in% 
               g1g2names$New.Genus[g1g2names$Taxonomic.status
                                   != "Accepted"]]

g1g2names[g1g2names$Taxonomic.status == "",] #Gutierrezia elegans; Arabis crandallii --> Boechera crandallii

g1g2names$Taxon
g1g2names$Taxon[g1g2names$Taxonomic.status == ""]
g1g2names$Taxon[g1g2names$Taxonomic.status == "Synonym"]

# All names to look for in SEINet, both the stated and the corrected by TPL
allnames <- c(g1g2names$Taxon, 
              paste(g1g2names$New.Genus,g1g2names$New.Species,sep=" "))
length(unique(allnames)) #71

```

```{r}
colorado <- get_map(location = c(lon = -105.6,lat = 39.35), 
                    maptype = "satellite", zoom = 6)
counties <- map_data("county")
co_county <- subset(counties, region == 'colorado')

#l1G1G2 are in UTM zone 13
l1G1G2@proj4string <- CRS("+proj=utm +zone=13 ellps=NAD83")

str(colorado)
str(co_county)

# Do these need to be projected? And to 
plot(l1G1G2)
```

Taxonomy 
```{r}
#1 change all G1G2 eor names to accepted - will need to deal with Gutierrezia elegans and Oonopsis sp. 1 by hand. 
#Level 1 EORs imported as l1G1G2 keeping only G1 G2 species
# l1G1G2$GNAME
g1g2names$toaccepted <- paste(g1g2names$New.Genus,
                    g1g2names$New.Species,
                    sep=" ")
l1G1G2$GNAME <- multisub(g1g2names$Taxon[g1g2names$Taxonomic.status=="Synonym"],
                         g1g2names$toaccepted[g1g2names$Taxonomic.status=="Synonym"],
                         l1G1G2$GNAME)


#2 change all SEINet names to accepted, go by TPL for simplicity, use coloradospsnames to keep old and who was wrong by herbarium
colonames$toaccepted <- paste(colonames$New.Genus,colonames$New.Species,
                              sep =" ")
# If I've already linked, I don't need to do a multisub, I can just take the name given in New 


```

```{r}
coloradospsnames$scientificName <-
  multisub(colonames$Taxon[colonames$Taxonomic.status=="Synonym"],
           colonames$toaccepted[colonames$Taxonomic.status=="Synonym"],
           coloradospsnames$scientificName)
write.csv(coloradospsnames, "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/SEINET G1 G2 Vouchers/occurrences_accepted.csv")
coloradospsnames_accpt <- read.csv("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_ArcMap_Projects/Link_EO_Herbarium_Records/SEINET G1 G2 Vouchers/occurrences_accepted.csv")

head(coloradospsnames_accpt)

#3 how many synonyms by institution or by year?
head(coloradosps) #has institution and now has accepted name, 
head(coloradospsnames_accpt) #with institution, yay!
head(colonames) # has taxonomic status
#get coloradosps again with synonyms
#poop, have names alone, not with institution....

table(coloradosps$institutionCode[coloradosps])



```

I think ignore this chunk
```{r}
#length(setdiff(allnames,species)) #3 are in G1G2 but not SEINet;  "Oonopsis sp. 1"      "Nuttallia rhizomata" "Oonopsis sp."  
#N. rhizomata --> Mentzelia rhizomata Reveal

#Need to link G1G2s to all names within SEINet
#Use TPL to make a list of names and there synonyms 
#how many specimens give an uncertainty amount? seinet is the list that Mo made; coloradosps is the list I pulled from SEINet; After fixing synonomy with TPL or ITIS will want to link l1G1G2 (the disaggregated CNHP EOR shapefiles) and coloradosps to limit to just G1/G2 species 
head(seinet)
table(seinet$basisOfRecord) #4 HumanObservation, 1871 PreservedSpecimen

head(seinet_Ewingdistances)
table(seinet_Ewingdistances$Map.Method)
#     GPS Locality       tr     trqs      trs 
#      63      495        6      114      126

head(coloG1G2)

```


All CNHP tracked species   
```{r}
#Each polygon might be made of multiple disconnected polygons. Need to separate and label
l1eor.disaggregate <- disaggregate(l1eor)

uniqueIDstring <- lapply(split(l1eor.disaggregate, l1eor.disaggregate$OBJECTID),
                         function(x){
                           x$POlyID <- paste(x$EO_ID, LETTERS[seq(from=1, to= nrow(x))],
                                             sep="")
                           x
                         })

polyid <- do.call(rbind, uniqueIDstring)
polyrast <- polyid
polyrast <- raster(polyrast)

```


```{r}
nrow(seinet[!is.na(seinet$decimalLatitude),]) #1875 points

#Get all points, transform to UTM
pts <- SpatialPoints(seinet[,c("decimalLongitude","decimalLatitude")])
proj4string(pts) <- CRS("+proj=longlat +datum=WGS84")

#Project data into a planar coordinate system (UTM zone 13)
UTMpts <- spTransform(pts, CRS("+proj=utm +zone=13 ellps=WGS84"))

head(pts)
head(UTMpts)

plot(l1eor[l1eor$GRANK %in% c("G1","G2"),])
points(UTMpts)
names(l1eor)

plot(l1eor[l1eor$GNAME == "Astragalus microcymbus",])

#Does not read th .prj file so need to tell it that they're in UTM
proj4string(l1eor) <- "+proj=utm +zone=13 ellps=WGS84"

#might be tooooo big!
#PLSS <- readOGR(dsn="Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/Colorado/PLSS Geo Coordinates_2008")



```


Try to do an apply and split by species, don't really care which EO, just which mapped occurance of a species is closest     
```{r}

mindistxsp <- lappply(unique(coloG1G2$scientificName), function(x){
  distout <- apply(gDistance(UTMpts[UTMpts]))
})

```


```{r}
#Distance from points to nearest polygon - this is to all polygons for G1:G5
mindist <- apply(gDistance(UTMpts, polyid, byid = TRUE),2,min)



## For each point, find name of nearest polygon

#name of the nearest polygon
minname <- apply()


```






# How do the maping techniques change by time, herbarium, or collector (amount of informaiton given)?   
```{r}
eors2 <- eors[eors$Year.2>0,]


methodxyear <- data.frame(table(eors2$Map.Method, eors2$Year.2))
methodxyear$Var2 <- as.numeric(as.character(methodxyear$Var2))

eors2$Map.Method <- factor(eors2$Map.Method, levels = c("Locality","GPS","trqs",
                                                        "trs","tr"))

ggplot(methodxyear, aes(Var2, Freq, colour=Var1))+
  geom_point()+
  stat_smooth(se=FALSE)+
  theme_bw()+
  xlab("Year")+
  ylab("Herbarium collections")+
  labs(color='Georeference Method')+
  theme(legend.justification=c(0,1), legend.position=c(0,1),
        legend.background = element_blank()) 
```


```{r}
ggplot(eors[eors$Year.2>0,], aes( Year.2,NEAR_DIST,colour=institutio))+
  geom_point()+
  facet_wrap(~institutio)+
  theme_bw()+
  stat_smooth(method="lm", se=FALSE)

ggplot(eors[eors$Year.2>0,], aes( Year.2,NEAR_DIST))+
  geom_point()+
  theme_bw()+
  stat_smooth(method="lm", se=FALSE)

summary(lm(NEAR_DIST~Year.2, data=eors[eors$Year.2>0,]))
```

How did the type of mapping method change over time?
```{r}

table(eors$Year.2)



```

Which method has the shortest distance?   
```{r}


ggplot(eors2, aes(x=Map.Method))+
  geom_bar(aes(fill=institutio))


ggplot(eors2, aes(Year.2,NEAR_DIST, colour=institutio))+
  geom_point()+
  theme_bw()

ggplot(eors2, aes(x=Year.2, fill=Map.Method))+
  geom_density(alpha=0.5)+
  theme_bw()
#+  facet_wrap(~Map.Method)

ggplot(eors2, aes(x=Year.2, fill=Map.Method))+
  geom_histogram(alpha=0.25)+
  theme_bw()


```

Maybe regional collections are much better?   
FLD = Fort Lewis College    
DES = Dessert Arizona   
MESA = Colorado Mesa University   
MISSA = Miss State Univ   
NMC = New Mexico State University    
SJNM = San Juan College, New Mexico   
UCR = University of California Riverside    
USUUB = Utah STate University Uintah Basin (probably is USU)   
UVSC = Utah Valley University	U.S.A. Utah. Orem.      
```{r}

local <- c("COLO","CS","DBG","FLD","MESA","RM","RMBL")
Adjacent <- c("ASC","ASU","BRY","DES","NMC","RM","SJNM","UNM","USUUB","UVSC")
far <- c("CM","F","MISSA","MO","NY","RENO","UCR")

eors2$InstNear[eors2$institutio %in% local] <- "Local"
eors2$InstNear[eors2$institutio %in% Adjacent] <- "Near"
eors2$InstNear[eors2$institutio %in% far] <- "Far"

eors2$InstNear <- factor(eors2$InstNear, levels = c("Far","Near","Local"))


ggplot(eors, aes(x=institutio))+
  geom_bar(aes(fill=Map.Method))


ggplot(eors2, aes(x=InstNear))+
  geom_bar(aes(fill=Map.Method))

ggplot(eors2, aes(Year.2,NEAR_DIST, colour=InstNear))+
  geom_point()+
  theme_bw()+
  stat_smooth(method="lm", se=FALSE)


ggplot(eors2, aes(NEAR_DIST))+
  geom_density(aes(fill=InstNear))
```

Would assume that Locality method would be best for local institutions, worse for far
```{r}


ggplot(eors2, aes(Map.Method, NEAR_DIST, colour=Map.Method))+
  geom_boxplot()+
  facet_wrap(~InstNear)+
  theme_bw()+
  xlab("Mapping method")+
  ylab("Distance to closest mapped occurence")
```

Assuming CNHP has accurate and complete data, then Distance to a known mapped location 
```{r}
table(eors2$Map.Method, eors2$InstNear)

length(table(eors2$scientific)) #53

#species are the names collected from SEINet
match(sort(unique(eors2$scientific)),sort(species))

InEOR <- do.call(rbind,lapply(split(eors2, eors2$scientific), function(x){
  out <- x[x$NEAR_DIST ==0,]
  out
})
)

head(InEOR)
table(InEOR$recordedBy)

# $georeferen is the only place to know who did the georeferencing, that would be very helpful to know
```


The Weber example of an out of state person mapping a collection to Gray's Peak from Chicago Lake that should be Mount Evans.     
- <https://conps.org/wp-content/uploads/2015/05/Mount_Evans_Summit_Lake_4-24-2011.pdf>    
```{r}


```


"coloradosps" == SEINet collections from Colorado - downloaded 12/6/2017 from SEINet      
"coloG1G2" == species matching names in L1 EORs but without any synonym corrections     
"seinet" == Mo worked on this, made distances from nearest EOR and looked at posted informaiton to figure out what might have been used to pick point on map        
      
ITISlist is synonmy pulled from ITIS into SQL    
TPL is pulling from synonym from the plant list   "colo"   
      "seinetMo"is synonymy with TPL for Mo's list of 53 species (that are G1/G2s)   
      "colonames" is synonym with TPL for all colorado specimens from SEINet   
```{r}
ITISlist <- read.csv("C:/Users/deprengm/Dropbox/Biological Collections_concerns&needs/ITIS_sqlsynonyms.csv")
length(unique(ITISlist$tsn[ITISlist$tsn != "NULL"])) #2528 are synonyms
length(unique(ITISlist$tsn_accepted[ITISlist$tsn_accepted != "NULL"])) #2277, 251 combined with synonmy
nrow(ITISlist[ITISlist$tsn != "NULL",]) #2530 are not synonyms 
nrow(ITISlist) #22570, huh. 

head(ITISlist[ITISlist$tsn != "NULL",])
ITISlist[ITISlist$tsn.1 == 897847,] #Want Equisetum giganteum
#Probably could have done this in SQL for the download.....
ITISsynonyms <- merge(ITISlist[,c("tsn","tsn_accepted","unit_name1","unit_name2")], 
                   ITISlist[,c("tsn.1","unit_name1","unit_name2")], by.x = "tsn_accepted", by.y = "tsn.1")

ITISsynonyms[ITISsynonyms$tsn_accepted == 897847,]
length(unique(ITISsynonyms$tsn_accepted)) #1777 --hummm... So these are the number of taxa (the accepted names) from ones that have some synonomy. There are still others that are just accepted. 
head(ITISsynonyms)

ITIScoloradosps <- coloradosps
ITIScoloradosps$genus
```




