Code snippits for sample design impact on SDM
Michelle DePrenger-Levin
2020-02-27

Partial dependence plots    
suggested packages pdp; edarf     
these seem to all be for random forest models. Can this be applied to a glm?    
```{r}
library(pdp)
library(edarf)
```
Plotting marginal effect of regression models    
This sounds more aligned with maxent output but how to tell?  
```{r}
library(sjPlot)
library(ggplot2)

```
```{r}
library(raster)

```

From the layers as-is (rasterstack3)
this is how I normalized layers for rasterstack_norm
```{r, eval=FALSE}
rasterstack_norm1 <- raster::stack("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/My Documents/UCDenver_phd/Dissertation/Chapter1/rasterstack_normalized.tif")

# rasterstack_norm <- rasterstack3
# # Normalize to 0-1
# for(i in 1:6){
#   rasterstack_norm[[i]] <- calc(rasterstack_norm[[1]], function(x){
#     (x-min(getValues(rasterstack_norm[[i]])))/(max(getValues(rasterstack_norm[[i]]))-min(getValues(rasterstack_norm[[i]])))
#   })
# }
```

Create my own glm for comparison with maxent
```{r}
# make a raster to hold response, suitability scores
predic_raster <- rasterstack_norm[[1]]

# lm to assign suitaiblity, but this is not 0-1
# predic_raster[] <- getValues(rasterstack_norm[[1]])*5 +
#                    getValues(rasterstack_norm[[2]])*2 + 
#                    getValues(rasterstack_norm[[3]])*3 + 
#                    getValues(rasterstack_norm[[4]])*4 + 
#                    getValues(rasterstack_norm[[5]])*6 +
#                    getValues(rasterstack_norm[[6]]*.01)
plot(predic_raster)
# this is not the same as y ~ exp(lm)/(1-exp(lm))

# trying exp(lm)/(1+exp(lm)) response is between 0 and 1
# lm1 <- formula(~getValues(rasterstack_norm[[1]]) + getValues(rasterstack_norm[[2]])*2 + getValues(rasterstack_norm[[3]])*3 + getValues(rasterstack_norm[[4]])*4 + getValues(rasterstack_norm[[5]])*5 + getValues(rasterstack_norm[[6]]))
lm1 <- formula(~getValues(rasterstack_norm[[1]])^2 + 
                 getValues(rasterstack_norm[[2]])+2 + 
                 getValues(rasterstack_norm[[3]])*2.5 - 
                 getValues(rasterstack_norm[[4]])*4 + 
                 getValues(rasterstack_norm[[5]])*5 + 
                 getValues(rasterstack_norm[[6]])*1)

hist(exp(as.formula(lm1[[2]]))/(1+exp(as.formula(lm1[[2]]))))

# Why this gives errors, it's the same length, should replace just fine. 
# > length(as.formula(lm1[[2]]))
# [1] 406808
# > length(1+exp(as.formula(lm1[[2]])))
# [1] 406808
# > length(predic_raster)
# [1] 406808
# setValues(x,values) values must be a vector - are there still NAs that are not treated as numeric?
# class(as.numeric(exp(as.formula(lm1[[2]]))/(1+exp(as.formula(lm1[[2]])))))
predic_raster <- setValues(predic_raster, as.numeric(exp(as.formula(lm1[[2]]))/(1+exp(as.formula(lm1[[2]]))))) # from a 'try-error' 
plot(predic_raster)

```
Try fancy response curves
```{r}
# make a raster to hold response, suitability scores
predic_raster <- rasterstack_norm[[1]]

# lm to assign suitaiblity, but this is not 0-1
# predic_raster[] <- getValues(rasterstack_norm[[1]])*5 +
#                    getValues(rasterstack_norm[[2]])*2 + 
#                    getValues(rasterstack_norm[[3]])*3 + 
#                    getValues(rasterstack_norm[[4]])*4 + 
#                    getValues(rasterstack_norm[[5]])*6 +
#                    getValues(rasterstack_norm[[6]]*.01)
plot(predic_raster)
# this is not the same as y ~ exp(lm)/(1-exp(lm))
```

```{r}
# trying exp(lm)/(1+exp(lm)) response is between 0 and 1
# lm1 <- formula(~getValues(rasterstack_norm[[1]]) + getValues(rasterstack_norm[[2]])*2 + getValues(rasterstack_norm[[3]])*3 + getValues(rasterstack_norm[[4]])*4 + getValues(rasterstack_norm[[5]])*5 + getValues(rasterstack_norm[[6]]))
beta1 <-1
alpha1 <- 1
alpha_1 <- 7
beta2 <-3
alpha_log2 <- 10
beta_log2 <- 10
lm1 <- formula(~getValues(rasterstack_norm[[1]])*5 + 
                 # lognormal
       (1/(getValues(rasterstack_norm[[2]])*sqrt(2*pi*beta1^2)))*(exp(-(log(getValues(rasterstack_norm[[2]]))-alpha1)^2)/(2*beta1^2)) + 
                 # beta 
                 (getValues(rasterstack_norm[[3]])^(alpha_1-1))*((1-getValues(rasterstack_norm[[3]]))^(beta2-1)) +
                 # lognormal
   (1/(getValues(rasterstack_norm[[4]])*sqrt(2*pi*beta_log2^2)))*(exp(-(log(getValues(rasterstack_norm[[4]]))-alpha_log2)^2)/(2*beta_log2^2)) +
             
                 getValues(rasterstack_norm[[5]])*2 + 
                 getValues(rasterstack_norm[[6]])*1)

hist(exp(as.formula(lm1[[2]]))/(1+exp(as.formula(lm1[[2]]))))

# Why this gives errors, it's the same length, should replace just fine. 
# > length(as.formula(lm1[[2]]))
# [1] 406808
# > length(1+exp(as.formula(lm1[[2]])))
# [1] 406808
# > length(predic_raster)
# [1] 406808
# setValues(x,values) values must be a vector - are there still NAs that are not treated as numeric?
# class(as.numeric(exp(as.formula(lm1[[2]]))/(1+exp(as.formula(lm1[[2]])))))
predic_raster <- setValues(predic_raster, as.numeric(exp(as.formula(lm1[[2]]))/(1+exp(as.formula(lm1[[2]]))))) # from a 'try-error' 
plot(predic_raster)

```


```{r}
lm2 <- formula(~getValues(rasterstack_norm[[1]])*0.5 + 
                 getValues(rasterstack_norm[[2]])*2 -
                 getValues(rasterstack_norm[[3]])*5 +
                 getValues(rasterstack_norm[[4]]) *
                 getValues(rasterstack_norm[[5]]) +  #interaction term 
                 getValues(rasterstack_norm[[4]])*1.5 + 
                 getValues(rasterstack_norm[[5]])*1.9 + 
                 getValues(rasterstack_norm[[6]])*1)

hist(exp(as.formula(lm2[[2]]))/(1+exp(as.formula(lm2[[2]]))))

predic_raster <- setValues(predic_raster, as.numeric(exp(as.formula(lm2[[2]]))/(1+exp(as.formula(lm2[[2]])))))
plot(predic_raster)


```




Or a slower way
```{r}
# same result, 1s and lots of NAs
predic_raster[] <- exp(getValues(rasterstack_norm[[1]])*1.5 + 
                   getValues(rasterstack_norm[[2]])*2 + 
                   getValues(rasterstack_norm[[3]])*3 + 
                   getValues(rasterstack_norm[[4]])*4 + 
                   getValues(rasterstack_norm[[5]])*5 + 
                   getValues(rasterstack_norm[[6]])*1)/
                (1+exp(getValues(rasterstack_norm[[1]])*1.5 + 
                   getValues(rasterstack_norm[[2]])*2 + 
                   getValues(rasterstack_norm[[3]])*3 + 
                   getValues(rasterstack_norm[[4]])*4 + 
                   getValues(rasterstack_norm[[5]])*5 + 
                   getValues(rasterstack_norm[[6]])*1))

plot(predic_raster)
```

Response curves?
```{r}
plot(rasterstack_norm[[1]][], predic_raster[],col="blue", cex=0.5, pch=16)

# Maxent outputs a response line with 'all other variables held constant'

modeldf <- data.frame(y=predic_raster[], x1=rasterstack_norm[[1]][],x2=rasterstack_norm[[2]][],x3=rasterstack_norm[[3]][],
                      x4=rasterstack_norm[[4]][],x5=rasterstack_norm[[5]][],x6=rasterstack_norm[[6]][])

lm <- lm(y ~ x1 +x2 + x3 + x4 + x5 + x6, data=modeldf)

theme_set(theme_sjplot())
sjPlot::plot_model(lm, type = "pred", terms = "x1") # plot predicted values (marginal effects) for specific model terms; warnings prediction from a rank-deficient fit may be misleading  


```




